---
title: "Wrangling & Analysis"
format: html
execute:
  cache: false       # disables caching
  eval: false        # optional: only enable when ready to run
---

## Background

These data come from a common garden drought experiment in Davis, California investigating the functional trait strategies of California grassland plants from five different grassland communities spanning California's latitudinal precipitation gradient. There 10-12 species per community or provenance, each provenance has 16 plots (8 drought, 8 control) across 10 shelters (5 drought, 5 control). In 2022-2023, we focused on collecting whole plant traits, including reproductive investment. This study focuses on how these traits relate to reproductive investment depending on native status, growth habit or provenance. This document will clean and wrangle the many datasets and images associated with this study.

## Traits & Objectives

-   Reproductive allocation
-   End of Season biomass
-   Relative Growth Rate
-   Photosynthetic Rate
-   Leaf Mass Area
-   Leaf Area
-   WUE Height
-   Leaf Nitrogen
-   Date of first flower
-   Mid season biomass partial
-   Root biomass partial
-   Root:shoot ratio partial
-   Root diameter partial
-   Specific root length partial
-   Root nitrogen partial

## To-do list
- Check for effects of drought treatment
- Check assumptions for models
- Fix Map orientation
- Licor outliers and mismatches
- Include seed mass in PCA
- Figure out if PCA is predicting something
- check for outliers in not RGR and non reproductive stuff

## Setup

```{r}
#| message: false
#| warning: false
#| echo: false

library(colorspace)
library(readr)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(lme4)
library(LeafArea)
#library(mapdata)
#library(maps)
#library(terra)
library(broom)
library(tidyverse)
#library(sf)
#library(geodata)
library(tigris)
library(ggnewscale)
#library(scatterpie)
library(lmtest)
library(car)
library(bbmle)
library(cropgrowdays)
library(emmeans)
library(here)
#install.packages("DHARMa")
library(DHARMa)
```

## CA Map and Plant Comp

Make a map of California with the inside of the boundary as a heat map of an aridity index. Then add five pie charts of plant composition.

Precip data from WorlClim

```{r}
#| eval: false
# ----- climate data -----
ppt <- worldclim_global(var = "prec", res = 2.5, path = "data/")
ppt_annual <- sum(ppt)

# ----- California boundary -----
ca <- states(cb = TRUE) |> subset(NAME == "California")
ca_vect <- vect(ca)

# ----- crop & mask -----
ppt_ca <- crop(ppt_annual, ca_vect)
ppt_ca <- mask(ppt_ca, ca_vect)

# ----- convert to df -----
ppt_df <- as.data.frame(ppt_ca, xy = TRUE)
names(ppt_df) <- c("x", "y", "ppt")

# ----- plot -----
ggplot(ppt_df) +
  geom_raster(aes(x = x, y = y, fill = ppt)) +
  geom_sf(data = ca, fill = NA, color = "black") +
  scale_fill_viridis_c(name = "Annual precip (mm)") +
  coord_sf() +
  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal()

rm(ppt_ca, ppt, ppt_annual)

# Put it all together now

## Need the pie chart data to have the following columns: ID, Lat, Long, Sp1 proportion, Sp2 proportion, etc.

# so we need to pivot wider the dataset and average across ten plots and add coordaintes for each site then use mapPies function
plant_comp <- read_csv(here::here("data/USDA_Site Composition.csv"))
str(plant_comp)

# First average across ten plots
comp_sum <- plant_comp %>%
  filter(!Species == "Grass") %>% 
  group_by(Site, Species) %>% 
  summarise(avg_cov = mean(Percent_Cover, na.rm = TRUE))

# get rid of bare ground, rock, and litter
comp_sum <- comp_sum %>%
  filter(!Species %in% c("Litter","Rock","Bare ground")) %>% 
  group_by(Site) %>% 
  arrange(desc(avg_cov), .by_group = TRUE)

PlList_2023 <- read_csv(here::here("data/USDA Drought Project 2023 Plant List.csv"))
pl_list <- PlList_2023 %>% 
  select(Site, Species)

comp_reduced <- left_join(pl_list, comp_sum, by=join_by(Site, Species))

comp_reduced$prop_cov <- comp_reduced$avg_cov/100

comp_reduced <- comp_reduced %>% 
  group_by(Site) %>% 
  mutate(prop_total = sum(prop_cov))

comp_reduced$prop_adj <- comp_reduced$prop_cov/comp_reduced$prop_total

comp_wide <- comp_reduced %>% 
  select(Site,
         Species,
         prop_adj) %>% 
  pivot_wider(names_from = Species,
              values_from = prop_adj)

## add lat and long columns
comp_wide$lat <- c(39.75141,37.37914,38.87508,34.70243,34.91771)
comp_wide$long <- c(-123.6302,-121.72841,-122.42114,-120.0545,-119.16882)
comp_wide$long_move <- comp_wide$long - 2
comp_wide$lat_move<- comp_wide$lat
comp_wide$lat_move[4] <- comp_wide$lat[4] + 1
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1
comp_wide$lat_move[1] <- comp_wide$lat[1] + 1
comp_wide$lat_move[2] <- comp_wide$lat[2] - 0.5
comp_wide$long_move[3] <- comp_wide$long[3] - 2.5
comp_wide$lat_move[3] <- comp_wide$lat[3] - 0.25
comp_wide$long_move[4] <- comp_wide$long[4] - 3.5
comp_wide$lat_move[4] <- comp_wide$lat[4] + 0.1
comp_wide$long_move[5] <- comp_wide$long[5] - 2.5
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1.5
#comp_wide$long_move[1] <- comp_wide$long[1] + 0.2
comp_wide

comp_wide[is.na(comp_wide)] <- 0
colors <- rainbow_hcl(37)


# ----- plot -----
ggplot() +
  geom_raster(data = ppt_df, aes(x = x, y = y, fill = ppt)) +
    scale_fill_viridis_c(name = "Annual precip (mm)") +
  new_scale_fill() +
 geom_segment(
  data = comp_wide,
  aes(x = long, y = lat, xend = long_move, yend = lat_move),
  color = "black", linewidth = 0.75
) +
   geom_scatterpie(
    data = comp_wide,
    aes(long_move, lat_move, r = 1),
    cols = c("Acmispon parviflorus"
             ,"Aira caryophyllea"
             ,"Bromus diandrus"
             ,"Bromus hordeaceus"
             ,"Clarkia purpurea"
             ,"Danthonia californica"
             ,"Daucus pusillus"
             ,"Eschscholzia californica"
             ,"Galium parisiense"
             ,"Hypochaeris glabra"
             ,"Madia gracilis"
             ,"Sanicula bipinnatifida"
             ,"Avena barbata"
             ,"Erodium botrys"
             ,"Geranium dissectum"
             ,"Lupinus bicolor"
             ,"Torilis arvensis"
             ,"Trifolium hirtum"
             ,"Vicia sativa"
             ,"Bromus madritensis"
             ,"Elymus caput-medusae"
             ,"Holocarpha virgata"
             ,"Micropus californicus"
             ,"Plantago erecta"
             ,"Avena fatua"
             ,"Calandrinia menziesii"
             ,"Centaurea melitensis"
             ,"Corethrogyne filaginifolia"
             ,"Erodium cicutarium"
             ,"Stipa pulchra"
             ,"Amsinckia menziesii"
             ,"Bromus arenarius"
             ,"Caulanthus coulteri"
             ,"Lupinus microcarpus"
             ,"Lupinus succulentus"
             ,"Phacelia tanacetifolia"
             ,"Mirabilis multiflora")   # your pie-slice variables
  ) +
scale_fill_manual(
    name = "Species",
    values = c(
             "Acmispon parviflorus" = colors[1]
             ,"Aira caryophyllea" = colors[2]
             ,"Bromus diandrus" = colors[3]
             ,"Bromus hordeaceus" = colors[4]
             ,"Clarkia purpurea" = colors[5]
             ,"Danthonia californica" = colors[6]
             ,"Daucus pusillus" = colors[7]
             ,"Eschscholzia californica" = colors[8]
             ,"Galium parisiense" = colors[9]
             ,"Hypochaeris glabra" = colors[10]
             ,"Madia gracilis" = colors[11]
             ,"Sanicula bipinnatifida" = colors[12]
             ,"Avena barbata" = colors[13]
             ,"Erodium botrys" = colors[14]
             ,"Geranium dissectum" = colors[15]
             ,"Lupinus bicolor" = colors[16]
             ,"Torilis arvensis" = colors[17]
             ,"Trifolium hirtum" = colors[18]
             ,"Vicia sativa" = colors[19]
             ,"Bromus madritensis" = colors[20]
             ,"Elymus caput-medusae" = colors[21]
             ,"Holocarpha virgata" = colors[22]
             ,"Micropus californicus" = colors[23]
             ,"Plantago erecta" = colors[24]
             ,"Avena fatua" = colors[25]
             ,"Calandrinia menziesii" = colors[26]
             ,"Centaurea melitensis" = colors[27]
             ,"Corethrogyne filaginifolia" = colors[28]
             ,"Erodium cicutarium" = colors[29]
             ,"Stipa pulchra" = colors[30]
             ,"Amsinckia menziesii" = colors[31]
             ,"Bromus arenarius" = colors[32]
             ,"Caulanthus coulteri" = colors[33]
             ,"Lupinus microcarpus" = colors[34]
             ,"Lupinus succulentus" = colors[35]
             ,"Phacelia tanacetifolia" = colors[36]
             ,"Mirabilis multiflora" = colors[37]
    )
    , guide = "none"
  ) +
  geom_sf(data = ca, fill = NA, color = "black",linewidth = 0.75) +
  coord_sf(xlim = c(-127, -113), ylim = c(32, 42)) +
  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = c(0.925,0.725)
  )  +
  geom_point(data = data.frame(lat = 38.544854, long = -121.784516), 
             aes(x = long, y = lat),
                 color = "red", 
                 size = 4)

ggsave(here::here("figs/CA_precip_comp_map.jpg"))

rm(ca,
   ca_vect,
   comp_reduced,
   comp_sum, 
   comp_wide,
   pl_list,
   plant_comp,
   PlList_2023,
   ppt_df)
```
## Reproductive allocation & End of Season biomass

Read in data, find outliers, fill in missing data

```{r}
#| eval: false
#| warning: false
## Read in Data
repro <- read_csv(here::here("data/Reproductive Allocation_2023.csv"))
str(repro)
summary(repro)
head(repro)

```

### Clean up and create variables

Remove rows that have NA in the Flower Count row and in the Veg Mass Row. 

```{r}
#| eval: false
# Remove individuals with no veg mass and no flower count
repro_narm <- repro %>% 
  drop_na(Flower_Count) %>% 
  drop_na(Veg_Mass)

# Read in plot data
plots <- read_csv(here::here("data/Plot_info.csv"))
str(plots)
summary(plots)
head(plots)

# Read in species data
spp <- read_csv(here::here("data/Species Codes.csv"))
unique(spp$Sp_Code)
unique(repro$Species)

# Pull plot code from Plant_ID, first two characters
repro_narm <- repro_narm %>% 
  mutate(Plot = substr(Plant_ID, start = 1, stop = 2))
unique(repro_narm$Plot)

# merge with plots
repro_full <- left_join(repro_narm,plots,by = join_by(Plot))

# Change name for consistency
repro_full$Sp_Code <- repro_full$Species

# merge with species data
repro_full <- left_join(repro_full,spp,by = join_by(Sp_Code))

# clean up dataset and re-order
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Perenniality,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Veg_Mass
  )


# repro_full[which(repro_full$Plant_ID == "G3.AVEFAT.1"),]
# two separate entries, go back and fix manually. Probably additive
# Additive! two flower bags and two veg bags.

# repro_full[which(repro_full$Plant_ID == "F2.BRODIA.1"),]
# was this counted twice?
# Counted twice, go with the heavier veg mass since debris was probably lost

# repro_full[which(repro_full$Plant_ID == "J1.BRODIA.1"),]
# was this counted twice?
# this was counted twice. The 4.01 mass is correct

rm(repro,repro_narm)
```

Now check for outliers
```{r}
#| eval: false
# Make a dataset so we can average flower mass per species x population x treatment and then find outliers

repro_outliers_vg <- repro_full %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(vg_avg = mean(Veg_Mass, na.rm = TRUE),
        vg_sd = sd(Veg_Mass, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Veg_Mass,
         vg_avg,
         vg_sd)

# remove any rows with NAs
repro_outliers_vg <- repro_outliers_vg[complete.cases(repro_outliers_vg),]

# get minimums and maximums using 3 standard deviations
veg_avgs <- repro_outliers_vg %>% 
  mutate(vg_max = vg_avg + 3*vg_sd,
         vg_min = vg_avg - 3*vg_sd)

# do the same for veg mass
vg_outliers <- veg_avgs[which(veg_avgs$Veg_Mass > veg_avgs$vg_max | veg_avgs$Veg_Mass < veg_avgs$vg_min),c("Plant_ID","Sp_Code","Population", "Veg_Mass","vg_avg","vg_sd","vg_min","vg_max")]
# Only 4 and they are just above the maximum, keep

# write_csv(vg_outliers, "vg_outliers.csv")

# now do separately for fls
repro_outliers_fl <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  select(Plant_ID,
         Sp_Code,
         Scientific_Name,
         Population,
         Flower,
         Flower_Mass_g,
         Veg_Mass) %>% 
  filter(!is.na(Flower_Mass_g))

# export dataset for intern/undergraduate practice
# write_csv(repro_outliers, "repro_intern.csv")

# Group by spp and population and then find averages and standard deviations
str(repro_outliers_fl)

repro_avgs_fl <- repro_outliers_fl %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(fl_avg = mean(Flower_Mass_g, na.rm = TRUE),
            fl_sd = sd(Flower_Mass_g, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Flower,
         Flower_Mass_g,
         fl_avg,
         fl_sd)

# remove any rows with NAs
repro_avgs_fl <- repro_avgs_fl[complete.cases(repro_avgs_fl),]

# get minimums and maximums using 3 standard deviations
repro_avgs_fl <- repro_avgs_fl %>% 
  mutate(fl_max = fl_avg + 3*fl_sd,
         fl_min = fl_avg - 3*fl_sd)

# Make a dataframe with rows of observations of flower masses that are either larger than the maximum or smaller than the minimum
fl_outliers <- repro_avgs_fl[which(repro_avgs_fl$Flower_Mass_g > repro_avgs_fl$fl_max | repro_avgs_fl$Flower_Mass_g < repro_avgs_fl$fl_min),c("Plant_ID","Sp_Code","Population", "Flower", "Flower_Mass_g","fl_avg","fl_sd","fl_min","fl_max")]

# make csvs to look deeper
# write_csv(fl_outliers, "fl_outliers.csv")

# find difference between flower mass and maximum
fl_outliers$dif <- fl_outliers$Flower_Mass_g-fl_outliers$fl_max

# how bad is it? anything bigger than 1 standard deviation difference is too big
fl_hi <- fl_outliers[which(fl_outliers$dif >= fl_outliers$fl_sd),]
# flower masses in fl_hi are way too big to include

# Make those flower masses NAs and replace in repro_full

# make repro_full long and then rbind and remove duplicates
fl_hi$Flower_Mass_g <- NA
fl_hi <- fl_hi %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    Flower,
    Flower_Mass_g) %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# make original dataset long
repro_full_long <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# create list of Plant_ID and flower number combos that need to become NA
nays <- fl_hi$ID_no

# create that dataset of 
repro_long_filt <- repro_full_long %>% 
  filter(ID_no %in% c(nays)) %>% 
  mutate(
    Flower_Mass_g = NA
  )
# 27 outliers converted to NA

# check your work, should return an NA
repro_long_filt$Flower_Mass_g[repro_long_filt$ID_no == sample(nays,1)]

# Now rbind, order, and remove duplicates
# 5620 in repro fill long, and 27 in repro long filt
# should have 5647
repro_fullfilt_long <- rbind(repro_long_filt,repro_full_long)

repro_long <- repro_fullfilt_long[-which(duplicated(repro_fullfilt_long$ID_no)),]

repro_full_2 <- repro_long %>% 
  pivot_wider(id_cols = Plant_ID,
              names_from = Flower,
              values_from = Flower_Mass_g)

# bind with original
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Veg_Mass
  )

repro_full_3 <- left_join(repro_full_2,repro_full,by=join_by(Plant_ID))

repro_full <- repro_full_3 %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Plot,
    Block, 
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Recorder,
    Collection_Date,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Flower_Count,
    Veg_Mass
  )

rm(repro_outliers_fl,
   repro_outliers_vg, 
   repro_avgs_fl,
   veg_avgs, 
   fl_outliers, 
   vg_outliers, 
   fl_hi, 
   repro_full_long, 
   nays, 
   repro_long_filt, 
   repro_fullfilt_long, 
   repro_long,
   repro_full_2,
   repro_full_3)
```

Now flower outliers have been removed, veg outliers have been checked.

Now fill in average flower mass for for each Species x population x treatment and fill in for individuals with no flower masses. 

```{r}
#| eval: false
# Average flower masses for each individual
firstfl = which(colnames(repro_full)=="Flower_1_g")
lastfl = which(colnames(repro_full)=="Flower_10_g")
repro_full[,c(firstfl:lastfl)]

repro_full$fl_average <- rowMeans(repro_full[,c(firstfl:lastfl)], na.rm = TRUE)

# Find the individuals with no flower mass
no_fls <- repro_full[which(repro_full$fl_average == "NaN"),]

# include everything except fl-Average
firstcol = which(colnames(no_fls)=="Plant_ID")
lastcol = which(colnames(no_fls)=="Veg_Mass")
no_fls <- no_fls[,c(firstcol:lastcol)]

# remove the individuals with no flower masses
repro_full <- repro_full[which(repro_full$fl_average != "NaN"),]

# detach(package:plyr)
# when plyr is loaded after dplyr, group_by doesn't work
avg_fls <- repro_full %>% 
  group_by(Sp_Code, Population, Treatment) %>% 
  summarise(fl_average = mean(fl_average, na.rm = TRUE))

# mast avg fls onto no_fls and then re-merge no-fls with larger dataset
rep_fls <- left_join(no_fls,avg_fls,by=join_by(Sp_Code,Population,Treatment))
repro_fill <- rbind(rep_fls,repro_full)

# check your work
repro_fill[which(is.na(repro_fill$fl_average)),]
# There's one without averages to fill in? Calmen in E6...investigate

repro_fill$total_fl_mass <- repro_fill$Flower_Count*repro_fill$fl_average

# There's an entry for ESCCAL which hasn't been done yet -- look into
# Later you can remove any species with only one entry

rm(repro_full,
   no_fls,
   avg_fls)

```
Now remove species with low data, less than three observations

```{r}
#| eval: false
unique(repro_fill$Sp_Code)

# Who are the species with low n?
summary <- repro_fill %>% 
  group_by(Sp_Code) %>% 
  summarise(n = n()) %>% 
  filter(n < 3)
# ACMPAR, CAUCOU, ESCCAL, LUPSUC

Sp_List_small <- summary$Sp_Code

# Remove species with less than 3 entries
repro_fill_hi_n <- repro_fill[-(which(repro_fill$Sp_Code %in% c(Sp_List_small))),]
unique(repro_fill_hi_n$Sp_Code)

summary(repro_fill_hi_n)

repro_fill_hi_n_narm <- repro_fill_hi_n[-(which(repro_fill_hi_n$total_fl_mass %in% c(NA, -Inf))),]

repro_fill_hi_n_narm_1 <- repro_fill_hi_n_narm

# Lastly add in precipitation
precip <- read_csv(here::here("data/Site_Precip.csv"))
repro_fill_hi_n_narm <- left_join(repro_fill_hi_n_narm_1,precip, by=join_by("Population"))

rm(summary,
   Sp_List_small,
   repro_fill,
   repro_fill_hi_n,
   precip,
   repro_fill_hi_n_narm_1
   )
```

Find and add in SPEI

```{r}
#| eval: false
install.packages("devtools")
require(devtools)
devtools::install_github('seschaub/getSpei')
require(getSpei)

workdir <- getwd()
setwd(workdir)

# create data frame:
Population  <- c("ANGELO", "MCL", "BORR","SEDG","WWP")
longitude    <- c(-123.6302, -121.72841, -122.42114, -120.0545, -119.16882)
latitude     <- c(39.75141, 37.37914, 38.87508, 34.70243, 34.91771)
locations_df <- data.frame(Population, longitude, latitude)

# specify arguments and run function:
d1 <- spec_spei(spei_files = "spei48", start_y = 2015, end_y = 2015, 
                locations = locations_df)

write.csv(d1, "SPEI_Pop.csv")
```

```{r}
#| eval: false
spei <- read_csv(here::here("data/SPEI_Pop.csv"))
spei <- spei %>% 
  filter(month > 2 & month < 6)


spei_avg <- spei %>% 
  group_by(location_id) %>% 
  summarise(spei_avg = mean(spei48)) %>% 
  mutate(Population = location_id)

spei_avg <- spei_avg[,c("Population","spei_avg")]

repro_fill_hi_n_narm_1 <- left_join(repro_fill_hi_n_narm,spei_avg, by=join_by("Population"))

repro_fill_hi_n_narm <- repro_fill_hi_n_narm_1

repro_fill_hi_n_narm$Veg_Mass_ln <- log(repro_fill_hi_n_narm$Veg_Mass)
repro_fill_hi_n_narm$total_fl_mass_ln <- log(repro_fill_hi_n_narm$total_fl_mass)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm[-(which(repro_fill_hi_n_narm$total_fl_mass_ln %in% c(NA, -Inf))),]

# Create ratio of fl to v
repro_fill_hi_n_narm$RV <- repro_fill_hi_n_narm$total_fl_mass_ln/repro_fill_hi_n_narm$Veg_Mass_ln

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  mutate(rating.n = (
    case_when(
      Rating == "Native" ~ 1
      , Rating == "Non-native" ~ 2
      , Rating == "Limited" ~ 3
      , Rating == "Moderate" ~ 4
      , Rating == "High" ~ 5
    )
  ))

rm(
  rep_fls
  , repro_fill_hi_n_narm_1
  , spei
  , spei_avg
)

write.csv(repro_fill_hi_n_narm, here::here("data/repro_fill_hi_n_narm.csv"))

```

We did it! Now graph and analyze. Use "repro_fill_hi_n_narm" for reproductive allocation.

Read in cleaned dataset to reduce computing power
```{r}
repro_fill_hi_n_narm <- read_csv(here::here("data/repro_fill_hi_n_narm.csv"))
```

## Q1: 1:1

```{r}
anova(lm(total_fl_mass_ln ~ Veg_Mass_ln*Scientific_Name, data = repro_fill_hi_n_narm))

mod <- lm(total_fl_mass_ln ~ Veg_Mass_ln * Scientific_Name, 
          data = repro_fill_hi_n_narm)


slopes <- emtrends(mod, ~ Scientific_Name, var = "Veg_Mass_ln")
slopes

slopes.df <- summary(slopes) %>% 
  arrange(desc(Veg_Mass_ln.trend))

test_slopes_vs1 <- summary(slopes, infer = TRUE, null = 1)
test_slopes_vs1[which(test_slopes_vs1$p.value<0.05),]

level_order <- unique(as.factor(slopes.df$Scientific_Name))


# Graph a bar plot of each species in order of slope with standard error bars and line at 1

#("ANGELO" = "#440154FF", "MCL" = "#3B528BFF", "BORR" = "#21908CFF", "SEDG" = "#5DC863FF", "WWP" = "#FDE725FF")) + 

# Make species names italic and 45 degrees
# Maybe make bar colors if significant

sigs <- test_slopes_vs1[which(test_slopes_vs1$p.value<0.05),1]

slopes.df$columnColor <- "N.S."

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[1])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[2])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[3])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[4])] <- "p-value < 0.05"

ggplot(slopes.df, aes(x = factor(Scientific_Name, level=c(level_order)), y = Veg_Mass_ln.trend
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.05" = "#440154FF")) +
  theme(
    axis.text.x = element_text(face = "italic", angle = 45, vjust = 1, hjust = 1),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp.jpg"))


ggplot(slopes.df, aes(x = factor(Scientific_Name, level=c(level_order)), y = Veg_Mass_ln.trend
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.05" = "#440154FF")) +
  theme(
    axis.text.x = element_blank(),
    legend.text =element_text(size=12, colour="black"),
    axis.text.y=element_text(size=12, colour="black"),
    axis.title=element_text(size=12,face="bold"),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp_prez.jpg"), width = 7,height = 3.5)

rm(
  mod
  , slopes
  , slopes.df
  , test_slopes_vs1
)

```

It's interesting that the three most widespread species are significant and low. And the arguable invasive native is the other significant.

## Q2: Growth Native

#### Model: RV
Model looking at what creates the ratio of log r and log v. Use species as random effect, and use that to determine best graph

```{r}
str(repro_fill_hi_n_narm)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  mutate(rating.n = (
    case_when(
      Rating == "Native" ~ 1
      , Rating == "Non-native" ~ 2
      , Rating == "Limited" ~ 3
      , Rating == "Moderate" ~ 4
      , Rating == "High" ~ 5
    )
  ))

m0 <- glm(RV ~ 1, family = gaussian, data = repro_fill_hi_n_narm)

m1 <- glmer(RV ~ 1 + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m2 <- glmer(RV ~ 1 + (1|Block/Shelter/Plot), family = gaussian, data = repro_fill_hi_n_narm)

m3 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment), family = gaussian, data = repro_fill_hi_n_narm)

m4 <- glmer(RV ~ 1 + (1|Block/Shelter/Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m5 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m6 <- glmer(RV ~ 1 + (1|Block) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

ICtab(m0,m1,m2,m3,m4,m5,m6)

m5 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m_ratingxpop <- glmer(RV ~ Rating*Population + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

ICtab(m5
      ,m_native
      ,m_rating
      ,m_growth
      ,m_growthxpop
      ,m_growthxprecip
      ,m_nativexgrowth
      ,m_nativexpop
      ,m_nativexprecip
      ,m_pdif
      ,m_precip
      ,m_spei
      ,m_ratingxgrowth
      ,m_ratingxpop
      ,m_ratingxprecip
      ,m_rating.n
      ,m_rating.nxgrowth
      ,m_rating.nxpop
      ,m_rating.nxprecip
      ,m_nativexspei
      ,m_ratingxspei
      ,m_growthxspei
      ,m_rating.nxspei)
lrtest(m_ratingxpop, m5)
Anova(m_ratingxpop)
summary(m_ratingxpop)

# try additive terms

repro_sub <- repro_fill_hi_n_narm %>% 
  select(
    RV,
    Native_Status,
    Rating,
    Rating_2,
    rating.n,
    Growth_Habit,
    Growth_2,
    Population,
    Precip_mm,
    precip_dif,
    spei_avg
  )

# Initialize an empty model
forward_model <- lm(RV ~ ., data = repro_sub)

# Forward stepwise regression
forward_model <- step(forward_model, direction = "forward", scope = formula(~ .))

# Initialize a model with all predictors
backward_model <- lm(RV ~ ., data = repro_sub)

# Backward stepwise regression
backward_model <- step(backward_model, direction = "backward")

# Initialize a model with all predictors
both_model <- lm(RV ~ ., data = repro_sub)

# Both-direction stepwise regression
both_model <- step(both_model, direction = "both")

# Check for overdispersion
# deviance/residual df


### Come back to this
### Kind of a mess

```

#### Model: Fl Mass

```{r}
str(repro_fill_hi_n_narm)

m0 <- glm(total_fl_mass_ln ~ 1, family = gaussian, data = repro_fill_hi_n_narm)

m1 <- glmer(total_fl_mass_ln ~ 1 + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m2 <- glmer(total_fl_mass_ln ~ 1 + (1|Block/Shelter/Plot), family = gaussian, data = repro_fill_hi_n_narm)

m3 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment), family = gaussian, data = repro_fill_hi_n_narm)

m4 <- glmer(total_fl_mass_ln ~ 1 + (1|Block/Shelter/Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m5 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m6 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m7 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

ICtab(m0,m1,m2,m3,m4,m5,m6,m7)

m7 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

mV <- glmer(total_fl_mass_ln ~ Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_native <- glmer(total_fl_mass_ln ~ Native_Status*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating <- glmer(total_fl_mass_ln ~ Rating*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_growth <- glmer(total_fl_mass_ln ~ Growth_Habit*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating.n <- glmer(total_fl_mass_ln ~ rating.n*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_growth2 <- glmer(total_fl_mass_ln ~ Growth_2*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating2 <- glmer(total_fl_mass_ln ~ Growth_2*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_nativexgrowth <- glmer(total_fl_mass_ln ~ Native_Status*Growth_Habit*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_ratingxgrowth <- glmer(total_fl_mass_ln ~ Rating*Growth_Habit*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating.nxgrowth <- glmer(total_fl_mass_ln ~ rating.n*Growth_Habit*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating2xgrowth <- glmer(total_fl_mass_ln ~ Rating_2*Growth_Habit*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_rating2xgrowth2 <- glmer(total_fl_mass_ln ~ Rating_2*Growth_2*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_nativexgrowth2 <- glmer(total_fl_mass_ln ~ Native_Status*Growth_2*Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_nativepgrowth <- glmer(total_fl_mass_ln ~ Native_Status + Growth_Habit + Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_nativep <- glmer(total_fl_mass_ln ~ Native_Status + Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

m_growthp <- glmer(total_fl_mass_ln ~ Growth_Habit + Veg_Mass_ln + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm)

# add in three way growth habit and three way native

ICtab(m7
      ,mV
      ,m_native
      ,m_rating
      ,m_growth
      ,m_rating.n
      ,m_growth2
      ,m_rating2
      ,m_nativexgrowth
      ,m_ratingxgrowth
      ,m_rating.nxgrowth
      ,m_rating2xgrowth
      ,m_rating2xgrowth2
      ,m_nativexgrowth2
      ,m_nativepgrowth
      ,m_nativep
      ,m_growthp)
anova(m_growth, mV)
Anova(m_growth)
summary(m_growth)
#mall

# Check assumptions

sim_res <- simulateResiduals(m_growth)
plot(sim_res)

plotResiduals(sim_res, form = repro_fill_hi_n_narm$Veg_Mass_ln)
plotResiduals(sim_res, form = repro_fill_hi_n_narm$Growth_Habit)
plotResiduals(sim_res, form = repro_fill_hi_n_narm$Veg_Mass_ln, group = repro_fill_hi_n_narm$Growth_Habit)
plotResiduals(sim_res, fitted(m_growth))




res <- resid(m_growth)
fitted <- fitted(m_growth)

plot(fitted, res)
abline(h = 0)

qqnorm(res)
qqline(res)

plot(fitted(m_growth), resid(m_growth))

plotResiduals(sim_res, form = dat$x1)

re <- ranef(m_growth)$group[,1]
qqnorm(re)
qqline(re)

testOutliers(sim_res)

car::vif(m_growth)


slopes <- emtrends(m_growth, ~ Growth_Habit, var = "Veg_Mass_ln")
slopes

slopes.df <- summary(slopes)

ggplot(slopes.df, aes(x = Growth_Habit, y = Veg_Mass_ln.trend
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity", fill = "#21908CFF") +
  # geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, size=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  labs(x = "Growth Habit",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/slopes_growth_pred.jpg"))




```




