---
title: "Wrangling & Analysis"
format: html
execute:
  cache: false       # disables caching
  eval: false        # optional: only enable when ready to run
---

## Background

These data come from a common garden drought experiment in Davis, California investigating the functional trait strategies of California grassland plants from five different grassland communities spanning California's latitudinal precipitation gradient. There 10-12 species per community or provenance, each provenance has 16 plots (8 drought, 8 control) across 10 shelters (5 drought, 5 control). In 2022-2023, we focused on collecting whole plant traits, including reproductive investment. This study focuses on how these traits relate to reproductive investment depending on native status, growth habit or provenance. This document will clean and wrangle the many datasets and images associated with this study as well as carry out analysis and creation of figures.

## Traits & Objectives

-   Reproductive allocation
-   End of Season biomass
-   Relative Growth Rate
-   Photosynthetic Rate
-   Leaf Mass Area
-   Leaf Area
-   WUE 
-   Height
-   Leaf Nitrogen
-   Date of first flower
-   Seed Mass

Limited Data:
-   Mid season biomass partial
-   Root biomass partial
-   Root:shoot ratio partial
-   Root diameter partial
-   Specific root length partial
-   Root nitrogen partial

Q0: Did drought treatment affect reproductive investment?
Q1: How much variation is there among annual California grassland plant in their allometric vegetative and reproductive scaling relationship?
Q2: Is there a difference in reproductive allometric scaling between annual forbs and grasses?
Q3: Does native status or provenance have an effect on reproductive scaling of forbs and grasse, separately?
Q4: How do aboveground traits associate with reproductive allocation and reproductive scaling?

## To-do list
- Clean up PCA step
- Break up final PCA lma into two steps

## Setup

```{r}
#| message: false
#| warning: false
#| echo: false

library(colorspace)
library(readr)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(lme4)
library(LeafArea)
#library(mapdata)
#library(maps)
#library(terra)
library(broom)
library(tidyverse)
#library(sf)
library(geodata)
library(tigris)
library(ggnewscale)
library(scatterpie)
library(lmtest)
library(car)
library(bbmle)
library(cropgrowdays)
library(emmeans)
library(here)
#install.packages("DHARMa")
library(DHARMa)
#install.packages("patchwork")
library(patchwork)
#install.packages("magick")
library(magick)
# install.packages("devtools")
require(devtools)
# devtools::install_github('seschaub/getSpei')
require(getSpei)
library(SPEI)
library(influence.ME)
# install.packages("MuMIn")
library(MuMIn)
library(ggeffects)
```

## Soil Sites

```{r}
#| eval: false
soil <- read_csv(here::here("data/Site Nitrogen.csv"))
str(soil)

soil <- soil |> 
  mutate(`NO3-N` = (
    case_when(
      `NO3-N` == "<0.10" ~ 0.05
      , TRUE ~ as.numeric(`NO3-N`)
    )
  )) |> 
  pivot_longer(cols = c(`NH4-N`, 
                        `NO3-N`),
               names_to = "Nitrogen_Type",
               values_to = "Concentration_ppm")

# Calculate mean and standard error for each site × nutrient
plot_dat <- soil %>%
  group_by(Site, Nitrogen_Type) %>%
  summarise(
    mean = mean(Concentration_ppm, na.rm = TRUE),
    se = sd(Concentration_ppm, na.rm = TRUE) / sqrt(sum(!is.na(Concentration_ppm))),
    .groups = "drop"
  )

# Specify which site should be red
red_site <- "OTL"

# Plot
ggplot(plot_dat, aes(x = mean, y = Nitrogen_Type, color = Site)) +
  geom_errorbar(
    aes(xmin = mean - se, xmax = mean + se),
    orientation = "y",
    width = 0,
    linewidth = 0.8
  ) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c(
      "ACR" = "black",
      "MCL" = "black",
      "BORR" = "black",
      "SEDG" = "black",
      "WWP" = "black",
      "OTL" = "red"
    )
  ) +
  labs(
    x = "Concentration (ppm)",
    y = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 11),
    axis.title.x = element_text(size = 12)
  )
  

library(dplyr)
library(ggplot2)

# Calculate mean and standard error for each site × nutrient
summary_data <- soil %>%
  group_by(Site, Nitrogen_Type) %>%
  summarise(
    Mean = mean(Concentration_ppm, na.rm = TRUE),
    SE = sd(Concentration_ppm, na.rm = TRUE) /
         sqrt(sum(!is.na(Concentration_ppm))),
    .groups = "drop"
  )

# Give each site a small vertical offset
summary_data <- summary_data %>%
  mutate(
    Nitrogen_Type = factor(
      Nitrogen_Type,
      levels = c("NO3-N", "NH4-N")
    ),
    Site = factor(
      Site,
      levels = c("ACR", "MCL", "BORR",
                 "SEDG", "WWP", "OTL")
    ),
    y_offset = as.numeric(Nitrogen_Type) +
               (as.numeric(Site) - 3.5) * 0.10
  )

# Plot
ggplot(summary_data, aes(
  x = Mean,
  y = y_offset,
  color = Site
)) +
  
  # Standard error lines
  geom_errorbar(
    aes(
      xmin = Mean - SE,
      xmax = Mean + SE
    ),
    orientation = "y",
    width = 0,
    linewidth = 0.8
  ) +
  
  # Mean points
  geom_point(size = 3) +
  
  # Five sites black, one site red
  scale_color_manual(
    values = c(
      "ACR" = "black",
      "MCL" = "black",
      "BORR" = "black",
      "SEDG" = "black",
      "WWP" = "black",
      "OTL" = "red"
    )
  ) +
  
  # Put the nutrient names back on the y axis
  scale_y_continuous(
    breaks = c(1, 2),
    labels = c("NO3-N", "NH4-N")
  ) +
  
  labs(
    x = "Soil Concentration (ppm)",
    y = NULL
  ) +
  
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 11),
    axis.title.x = element_text(size = 12)
  )

ggsave(here::here("figs/soil_nitrogen.jpg"), height = 3.5, width = 7)

```

## Water sites
Do the same thing as the nitrogen figure but for MAP (or current year?)

I could do soil moisture, same year rainfall, or MAP...
Below is soil moisture

```{r}
#| eval: false
sm <- read_csv(here::here("data/Site Soil Moisture.csv"))
str(sm)

# Calculate mean and SE for each site
summary_data <- sm %>%
  group_by(Site) %>%
  summarise(
    Mean = mean(sm, na.rm = TRUE),   # replace with your column name
    SE   = sd(sm, na.rm = TRUE) /
           sqrt(sum(!is.na(sm))),
    .groups = "drop"
  ) %>%
  arrange(Site) %>%                          # sets a consistent order
  mutate(
    row      = row_number(),
    n        = n(),
    y_offset = 1 + (row - (n + 1) / 2) * 0.10
  )

  
# Plot
ggplot(summary_data, aes(
  x     = Mean,
  y     = y_offset,
  color = Site
)) +
  geom_errorbar(
    aes(xmin = Mean - SE, xmax = Mean + SE),
    orientation = "y",
    width        = 0,
    linewidth    = 0.8
  ) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c(
      "Angelo"  = "black",
      "McL"  = "black",
      "BORR" = "black",
      "Sedgwick" = "black",
      "WW"  = "black",
      "OTL-Drought"  = "red",
      "OTL-Control" = "blue"
    )
  ) +
  scale_y_continuous(
    breaks = 1,
    labels = "Sites",
    limits = c(0.5, 1.5)    # padding so jittered points aren't clipped
  ) +
  labs(
    x = "Soil Moisture (% VWC)",   # adjust units as needed
    y = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.y     = element_text(size = 12),
    axis.text.x     = element_text(size = 11),
    axis.title.x    = element_text(size = 12)
  )

ggsave(here::here("figs/soil_moisture.jpg"), height = 3.5, width = 7)


```


## CA Map and Plant Comp

Make a map of California with the inside of the boundary as a heat map of an aridity index. Then add five pie charts of plant composition.

Precip data from WorlClim

```{r}
#| eval: false
#| echo: false
# ----- climate data -----
ppt <- worldclim_global(var = "prec", res = 2.5, path = "data/")
ppt_annual <- sum(ppt)
?worldclim_global
# ----- California boundary -----
ca <- states(cb = TRUE) |> subset(NAME == "California")
ca_vect <- vect(ca)

# ----- crop & mask -----
ppt_ca <- crop(ppt_annual, ca_vect)
ppt_ca <- mask(ppt_ca, ca_vect)

# ----- convert to df -----
ppt_df <- as.data.frame(ppt_ca, xy = TRUE)
names(ppt_df) <- c("x", "y", "ppt")

# ----- plot -----
ggplot(ppt_df) +
  geom_raster(aes(x = x, y = y, fill = ppt)) +
  geom_sf(data = ca, fill = NA, color = "black") +
  scale_fill_viridis_c(name = "Annual precip (mm)") +
  coord_sf() +
  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal()

rm(ppt_ca, ppt, ppt_annual)

# Put it all together now

## Need the pie chart data to have the following columns: ID, Lat, Long, Sp1 proportion, Sp2 proportion, etc.

# so we need to pivot wider the dataset and average across ten plots and add coordaintes for each site then use mapPies function
plant_comp <- read_csv(here::here("data/USDA_Site Composition.csv"))
str(plant_comp)

# First average across ten plots
comp_sum <- plant_comp %>%
  filter(!Species == "Grass") %>% 
  group_by(Site, Species) %>% 
  summarise(avg_cov = mean(Percent_Cover, na.rm = TRUE))

# get rid of bare ground, rock, and litter
comp_sum <- comp_sum %>%
  filter(!Species %in% c("Litter","Rock","Bare ground")) %>% 
  group_by(Site) %>% 
  arrange(desc(avg_cov), .by_group = TRUE)

PlList_2023 <- read_csv(here::here("data/USDA Drought Project 2023 Plant List.csv"))
pl_list <- PlList_2023 %>% 
  select(Site, Species)

comp_reduced <- left_join(pl_list, comp_sum, by=join_by(Site, Species))

comp_reduced$prop_cov <- comp_reduced$avg_cov/100

comp_reduced <- comp_reduced %>% 
  group_by(Site) %>% 
  mutate(prop_total = sum(prop_cov))

comp_reduced$prop_adj <- comp_reduced$prop_cov/comp_reduced$prop_total

comp_wide <- comp_reduced %>% 
  select(Site,
         Species,
         prop_adj) %>% 
  pivot_wider(names_from = Species,
              values_from = prop_adj)

## add lat and long columns
comp_wide$lat <- c(39.75141,37.37914,38.87508,34.70243,34.91771)
comp_wide$long <- c(-123.6302,-121.72841,-122.42114,-120.0545,-119.16882)
comp_wide$long_move <- comp_wide$long - 2
comp_wide$lat_move<- comp_wide$lat
comp_wide$lat_move[4] <- comp_wide$lat[4] + 1
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1
comp_wide$lat_move[1] <- comp_wide$lat[1] + 1
comp_wide$lat_move[2] <- comp_wide$lat[2] - 0.5
comp_wide$long_move[3] <- comp_wide$long[3] - 2.5
comp_wide$lat_move[3] <- comp_wide$lat[3] - 0.25
comp_wide$long_move[4] <- comp_wide$long[4] - 3.5
comp_wide$lat_move[4] <- comp_wide$lat[4] + 0.1
comp_wide$long_move[5] <- comp_wide$long[5] - 2.5
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1.5
#comp_wide$long_move[1] <- comp_wide$long[1] + 0.2
comp_wide

comp_wide[is.na(comp_wide)] <- 0
my_colors <- rainbow_hcl(37)


# ----- plot -----
ggplot() +
  geom_raster(data = ppt_df, aes(x = x, y = y, fill = ppt)) +
    scale_fill_viridis_c(name = "Annual precip (mm)",
                         direction = -1) +
  new_scale_fill() +
 geom_segment(
  data = comp_wide,
  aes(x = long, y = lat, xend = long_move, yend = lat_move),
  color = "black", linewidth = 0.75
) +
   geom_scatterpie(
    data = comp_wide,
    aes(long_move, lat_move, r = 1),
    cols = c("Acmispon parviflorus"
             ,"Aira caryophyllea"
             ,"Bromus diandrus"
             ,"Bromus hordeaceus"
             ,"Clarkia purpurea"
             ,"Danthonia californica"
             ,"Daucus pusillus"
             ,"Eschscholzia californica"
             ,"Galium parisiense"
             ,"Hypochaeris glabra"
             ,"Madia gracilis"
             ,"Sanicula bipinnatifida"
             ,"Avena barbata"
             ,"Erodium botrys"
             ,"Geranium dissectum"
             ,"Lupinus bicolor"
             ,"Torilis arvensis"
             ,"Trifolium hirtum"
             ,"Vicia sativa"
             ,"Bromus madritensis"
             ,"Elymus caput-medusae"
             ,"Holocarpha virgata"
             ,"Micropus californicus"
             ,"Plantago erecta"
             ,"Avena fatua"
             ,"Calandrinia menziesii"
             ,"Centaurea melitensis"
             ,"Corethrogyne filaginifolia"
             ,"Erodium cicutarium"
             ,"Stipa pulchra"
             ,"Amsinckia menziesii"
             ,"Bromus arenarius"
             ,"Caulanthus coulteri"
             ,"Lupinus microcarpus"
             ,"Lupinus succulentus"
             ,"Phacelia tanacetifolia"
             ,"Mirabilis multiflora")   # your pie-slice variables
  ) +
scale_fill_manual(
    name = "Species",
    values = c(
             "Acmispon parviflorus" = my_colors[1]
             ,"Aira caryophyllea" = my_colors[2]
             ,"Bromus diandrus" = my_colors[3]
             ,"Bromus hordeaceus" = my_colors[4]
             ,"Clarkia purpurea" = my_colors[5]
             ,"Danthonia californica" = my_colors[6]
             ,"Daucus pusillus" = my_colors[7]
             ,"Eschscholzia californica" = my_colors[8]
             ,"Galium parisiense" = my_colors[9]
             ,"Hypochaeris glabra" = my_colors[10]
             ,"Madia gracilis" = my_colors[11]
             ,"Sanicula bipinnatifida" = my_colors[12]
             ,"Avena barbata" = my_colors[13]
             ,"Erodium botrys" = my_colors[14]
             ,"Geranium dissectum" = my_colors[15]
             ,"Lupinus bicolor" = my_colors[16]
             ,"Torilis arvensis" = my_colors[17]
             ,"Trifolium hirtum" = my_colors[18]
             ,"Vicia sativa" = my_colors[19]
             ,"Bromus madritensis" = my_colors[20]
             ,"Elymus caput-medusae" = my_colors[21]
             ,"Holocarpha virgata" = my_colors[22]
             ,"Micropus californicus" = my_colors[23]
             ,"Plantago erecta" = my_colors[24]
             ,"Avena fatua" = my_colors[25]
             ,"Calandrinia menziesii" = my_colors[26]
             ,"Centaurea melitensis" = my_colors[27]
             ,"Corethrogyne filaginifolia" = my_colors[28]
             ,"Erodium cicutarium" = my_colors[29]
             ,"Stipa pulchra" = my_colors[30]
             ,"Amsinckia menziesii" = my_colors[31]
             ,"Bromus arenarius" = my_colors[32]
             ,"Caulanthus coulteri" = my_colors[33]
             ,"Lupinus microcarpus" = my_colors[34]
             ,"Lupinus succulentus" = my_colors[35]
             ,"Phacelia tanacetifolia" = my_colors[36]
             ,"Mirabilis multiflora" = my_colors[37]
    )
    , guide = "none"
  ) +
  geom_sf(data = ca, fill = NA, color = "black",linewidth = 0.75) +
  coord_sf(xlim = c(-127, -113), ylim = c(32, 42)) +
#  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = c(0.925,0.725)
  )  +
  geom_point(data = data.frame(lat = 38.544854, long = -121.784516), 
             aes(x = long, y = lat),
                 color = "red", 
                 size = 4)


ggsave(here::here("figs/CA_precip_comp_map.jpg"))

 rm(ca,
   ca_vect,
   comp_reduced,
   comp_sum, 
   comp_wide,
   pl_list,
   plant_comp,
   PlList_2023,
   ppt_df)
```

```{r}
#| eval: false
library(ggplot2)

species <- c(
  "Acmispon parviflorus",
  "Aira caryophyllea",
  "Bromus diandrus",
  "Bromus hordeaceus",
  "Clarkia purpurea",
  "Danthonia californica*",
  "Daucus pusillus",
  "Eschscholzia californica*",
  "Galium parisiense",
  "Hypochaeris glabra",
  "Madia gracilis",
  "Sanicula bipinnatifida*",
  "Avena barbata",
  "Erodium botrys",
  "Geranium dissectum",
  "Lupinus bicolor",
  "Torilis arvensis",
  "Trifolium hirtum",
  "Vicia sativa",
  "Bromus madritensis",
  "Elymus caput-medusae",
  "Holocarpha virgata",
  "Micropus californicus",
  "Plantago erecta",
  "Avena fatua",
  "Calandrinia menziesii",
  "Centaurea melitensis",
  "Corethrogyne filaginifolia*",
  "Erodium cicutarium",
  "Stipa pulchra*",
  "Amsinckia menziesii",
  "Bromus arenarius",
  "Caulanthus coulteri",
  "Lupinus microcarpus",
  "Lupinus succulentus",
  "Phacelia tanacetifolia",
  "Mirabilis multiflora*"
)

legend_df <- data.frame(
  Species = factor(species, levels = species),
  x = 1,
  y = seq_along(species)
)

legend_plot <- ggplot(legend_df,
                      aes(x, y, fill = Species)) +
  geom_point(shape = 21, size = 5) +
  scale_fill_manual(
    values = c(
              "Acmispon parviflorus" = my_colors[1]
             ,"Aira caryophyllea" = my_colors[2]
             ,"Bromus diandrus" = my_colors[3]
             ,"Bromus hordeaceus" = my_colors[4]
             ,"Clarkia purpurea" = my_colors[5]
             ,"Danthonia californica*" = my_colors[6]
             ,"Daucus pusillus" = my_colors[7]
             ,"Eschscholzia californica*" = my_colors[8]
             ,"Galium parisiense" = my_colors[9]
             ,"Hypochaeris glabra" = my_colors[10]
             ,"Madia gracilis" = my_colors[11]
             ,"Sanicula bipinnatifida*" = my_colors[12]
             ,"Avena barbata" = my_colors[13]
             ,"Erodium botrys" = my_colors[14]
             ,"Geranium dissectum" = my_colors[15]
             ,"Lupinus bicolor" = my_colors[16]
             ,"Torilis arvensis" = my_colors[17]
             ,"Trifolium hirtum" = my_colors[18]
             ,"Vicia sativa" = my_colors[19]
             ,"Bromus madritensis" = my_colors[20]
             ,"Elymus caput-medusae" = my_colors[21]
             ,"Holocarpha virgata" = my_colors[22]
             ,"Micropus californicus" = my_colors[23]
             ,"Plantago erecta" = my_colors[24]
             ,"Avena fatua" = my_colors[25]
             ,"Calandrinia menziesii" = my_colors[26]
             ,"Centaurea melitensis" = my_colors[27]
             ,"Corethrogyne filaginifolia*" = my_colors[28]
             ,"Erodium cicutarium" = my_colors[29]
             ,"Stipa pulchra*" = my_colors[30]
             ,"Amsinckia menziesii" = my_colors[31]
             ,"Bromus arenarius" = my_colors[32]
             ,"Caulanthus coulteri" = my_colors[33]
             ,"Lupinus microcarpus" = my_colors[34]
             ,"Lupinus succulentus" = my_colors[35]
             ,"Phacelia tanacetifolia" = my_colors[36]
             ,"Mirabilis multiflora*" = my_colors[37]
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(face = "bold"),
  legend.text = element_text(face = "bold.italic")
  )

ggsave(here::here("figs/SpeciesLegend.jpg"),
       legend_plot,
       width = 4,
       height = 10)
```

Combine the two in powerpoint

```{r}
#| label: fig-plot
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "plot"

# save image as ggplot
my_plot_ggplot <- image_read(here::here("figs/CA_precip_comp_map.jpg")) %>% 
  image_ggplot(interpolate = TRUE)
my_plot_ggplot
rm(my_plot_ggplot)
```
Figure 1. Plant composition of 12 most dominant species in 5 grasslands along latitudinal precipitation gradient in California. Colors in pie chart reflect distinct species. 

## Reproductive allocation & End of Season biomass

Read in data, find outliers, fill in missing data

```{r}
#| eval: false
#| warning: false
#| echo: false
## Read in Data
repro <- read_csv(here::here("data/Reproductive Allocation_2023.csv"))
str(repro)
summary(repro)
head(repro)

```

### Clean up and create variables

Removed rows that have NA in the Flower Count row and in the Veg Mass Row. 

```{r}
#| eval: false
#| echo: false
# Remove individuals with no veg mass and no flower count
repro_narm <- repro %>% 
  drop_na(Flower_Count) %>% 
  drop_na(Veg_Mass)

# Read in plot data
plots <- read_csv(here::here("data/Plot_info.csv"))
str(plots)
summary(plots)
head(plots)

# Read in species data
spp <- read_csv(here::here("data/Species Codes.csv"))
unique(spp$Sp_Code)
unique(repro$Species)

# Pull plot code from Plant_ID, first two characters
repro_narm <- repro_narm %>% 
  mutate(Plot = substr(Plant_ID, start = 1, stop = 2))
unique(repro_narm$Plot)

# merge with plots
repro_full <- left_join(repro_narm,plots,by = join_by(Plot))

# Change name for consistency
repro_full$Sp_Code <- repro_full$Species

# merge with species data
repro_full <- left_join(repro_full,spp,by = join_by(Sp_Code))

# Remove perennials
repro_full <- repro_full %>% 
  filter(!Sp_Code %in% c("ESCCAL","STIPUL"))

# clean up dataset and re-order
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Perenniality,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Veg_Mass
  )


# repro_full[which(repro_full$Plant_ID == "G3.AVEFAT.1"),]
# two separate entries, go back and fix manually. Probably additive
# Additive! two flower bags and two veg bags.

# repro_full[which(repro_full$Plant_ID == "F2.BRODIA.1"),]
# was this counted twice?
# Counted twice, go with the heavier veg mass since debris was probably lost

# repro_full[which(repro_full$Plant_ID == "J1.BRODIA.1"),]
# was this counted twice?
# this was counted twice. The 4.01 mass is correct

rm(repro,repro_narm)
```
Checked for outliers
```{r}
#| eval: false
#| echo: false
# Make a dataset so we can average flower mass per species x population x treatment and then find outliers

repro_outliers_vg <- repro_full %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(vg_avg = mean(Veg_Mass, na.rm = TRUE),
        vg_sd = sd(Veg_Mass, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Veg_Mass,
         vg_avg,
         vg_sd)

# remove any rows with NAs
repro_outliers_vg <- repro_outliers_vg[complete.cases(repro_outliers_vg),]

# get minimums and maximums using 3 standard deviations
veg_avgs <- repro_outliers_vg %>% 
  mutate(vg_max = vg_avg + 3*vg_sd,
         vg_min = vg_avg - 3*vg_sd)

# do the same for veg mass
vg_outliers <- veg_avgs[which(veg_avgs$Veg_Mass > veg_avgs$vg_max | veg_avgs$Veg_Mass < veg_avgs$vg_min),c("Plant_ID","Sp_Code","Population", "Veg_Mass","vg_avg","vg_sd","vg_min","vg_max")]
# Only 4 and they are just above the maximum, keep

# write_csv(vg_outliers, "vg_outliers.csv")

# now do separately for fls
repro_outliers_fl <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  select(Plant_ID,
         Sp_Code,
         Scientific_Name,
         Population,
         Flower,
         Flower_Mass_g,
         Veg_Mass) %>% 
  filter(!is.na(Flower_Mass_g))

# export dataset for intern/undergraduate practice
# write_csv(repro_outliers, "repro_intern.csv")

# Group by spp and population and then find averages and standard deviations
str(repro_outliers_fl)

repro_avgs_fl <- repro_outliers_fl %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(fl_avg = mean(Flower_Mass_g, na.rm = TRUE),
            fl_sd = sd(Flower_Mass_g, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Flower,
         Flower_Mass_g,
         fl_avg,
         fl_sd)

# remove any rows with NAs
repro_avgs_fl <- repro_avgs_fl[complete.cases(repro_avgs_fl),]

# get minimums and maximums using 3 standard deviations
repro_avgs_fl <- repro_avgs_fl %>% 
  mutate(fl_max = fl_avg + 3*fl_sd,
         fl_min = fl_avg - 3*fl_sd)

# Make a dataframe with rows of observations of flower masses that are either larger than the maximum or smaller than the minimum
fl_outliers <- repro_avgs_fl[which(repro_avgs_fl$Flower_Mass_g > repro_avgs_fl$fl_max | repro_avgs_fl$Flower_Mass_g < repro_avgs_fl$fl_min),c("Plant_ID","Sp_Code","Population", "Flower", "Flower_Mass_g","fl_avg","fl_sd","fl_min","fl_max")]

# make csvs to look deeper
# write_csv(fl_outliers, "fl_outliers.csv")

# find difference between flower mass and maximum
fl_outliers$dif <- fl_outliers$Flower_Mass_g-fl_outliers$fl_max

# how bad is it? anything bigger than 1 standard deviation difference is too big
fl_hi <- fl_outliers[which(fl_outliers$dif >= fl_outliers$fl_sd),]
# flower masses in fl_hi are way too big to include

# Make those flower masses NAs and replace in repro_full

# make repro_full long and then rbind and remove duplicates
fl_hi$Flower_Mass_g <- NA
fl_hi <- fl_hi %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    Flower,
    Flower_Mass_g) %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# make original dataset long
repro_full_long <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# create list of Plant_ID and flower number combos that need to become NA
nays <- fl_hi$ID_no

# create that dataset of 
repro_long_filt <- repro_full_long %>% 
  filter(ID_no %in% c(nays)) %>% 
  mutate(
    Flower_Mass_g = NA
  )
# 27 outliers converted to NA

# check your work, should return an NA
repro_long_filt$Flower_Mass_g[repro_long_filt$ID_no == sample(nays,1)]

# Now rbind, order, and remove duplicates
# 5620 in repro fill long, and 27 in repro long filt
# should have 5647
repro_fullfilt_long <- rbind(repro_long_filt,repro_full_long)

repro_long <- repro_fullfilt_long[-which(duplicated(repro_fullfilt_long$ID_no)),]

repro_full_2 <- repro_long %>% 
  pivot_wider(id_cols = Plant_ID,
              names_from = Flower,
              values_from = Flower_Mass_g)

# bind with original
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Veg_Mass
  )

repro_full_3 <- left_join(repro_full_2,repro_full,by=join_by(Plant_ID))

repro_full <- repro_full_3 %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Plot,
    Block, 
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Recorder,
    Collection_Date,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Flower_Count,
    Veg_Mass
  )

rm(repro_outliers_fl,
   repro_outliers_vg, 
   repro_avgs_fl,
   veg_avgs, 
   fl_outliers, 
   vg_outliers, 
   fl_hi, 
   repro_full_long, 
   nays, 
   repro_long_filt, 
   repro_fullfilt_long, 
   repro_long,
   repro_full_2,
   repro_full_3)
```
Now flower outliers have been removed, veg outliers have been checked.

Then filled in average flower mass for for each Species x population x treatment and fill in for individuals with no flower masses. 

```{r}
#| eval: false
#| echo: false
# Average flower masses for each individual
firstfl = which(colnames(repro_full)=="Flower_1_g")
lastfl = which(colnames(repro_full)=="Flower_10_g")
repro_full[,c(firstfl:lastfl)]

repro_full$fl_average <- rowMeans(repro_full[,c(firstfl:lastfl)], na.rm = TRUE)

# Find the individuals with no flower mass
no_fls <- repro_full[which(repro_full$fl_average == "NaN"),]

# include everything except fl-Average
firstcol = which(colnames(no_fls)=="Plant_ID")
lastcol = which(colnames(no_fls)=="Veg_Mass")
no_fls <- no_fls[,c(firstcol:lastcol)]

# remove the individuals with no flower masses
repro_full <- repro_full[which(repro_full$fl_average != "NaN"),]

# detach(package:plyr)
# when plyr is loaded after dplyr, group_by doesn't work
avg_fls <- repro_full %>% 
  group_by(Sp_Code, Population, Treatment) %>% 
  summarise(fl_average = mean(fl_average, na.rm = TRUE))

# mast avg fls onto no_fls and then re-merge no-fls with larger dataset
rep_fls <- left_join(no_fls,avg_fls,by=join_by(Sp_Code,Population,Treatment))
repro_fill <- rbind(rep_fls,repro_full)

# check your work
repro_fill[which(is.na(repro_fill$fl_average)),]
# There's one without averages to fill in? Calmen in E6...investigate

repro_fill$total_fl_mass <- repro_fill$Flower_Count*repro_fill$fl_average

# There's an entry for ESCCAL which hasn't been done yet -- look into
# Later you can remove any species with only one entry

rm(repro_full,
   no_fls,
   avg_fls)

```
Kept species with low data

```{r}
#| eval: false
unique(repro_fill$Sp_Code)

# Who are the species with low n?
#summary <- repro_fill %>% 
#  group_by(Sp_Code) %>% 
#  summarise(n = n()) %>% 
#  filter(n < 3)
# ACMPAR, CAUCOU, ESCCAL, LUPSUC

#Sp_List_small <- summary$Sp_Code

# Remove species with less than 3 entries
#repro_fill_hi_n <- repro_fill[-(which(repro_fill$Sp_Code %in% c(Sp_List_small))),]
#unique(repro_fill_hi_n$Sp_Code)

summary(repro_fill)

repro_fill_hi_n_narm <- repro_fill[-(which(repro_fill$total_fl_mass %in% c(NA, -Inf))),]

repro_fill_hi_n_narm_1 <- repro_fill_hi_n_narm
# Lastly add in precipitation
precip <- read_csv(here::here("data/Site_Precip.csv"))
repro_fill_hi_n_narm <- left_join(repro_fill_hi_n_narm_1,precip, by=join_by("Population"))

rm(summary,
   Sp_List_small,
   repro_fill,
   repro_fill_hi_n,
   precip,
   repro_fill_hi_n_narm_1
   )
```

Found and added in SPEI using package `getSPEI`

HERE'S where we replace

Try using AI and CWD from TerraClimate. This gives climate data for roughly 4 km by 4 km grid cell

```{r}
#| eval: false

#install.packages("remotes")  # if you don't have it
#remotes::install_github("mikejohnson51/AOI")
#remotes::install_github("mikejohnson51/climateR")
library(AOI)
library(climateR)
library(sf)
library(terra)
library(dplyr)

# Define your five points
sites <- data.frame(
  site = c("SEDG", "WWP", "MCL", "BORR", "ANGELO"),
  lon  = c(-120.1, -119.2, -121.7, -122.4, -123.6),
  lat  = c(34.7,   34.9,   37.4,   38.9,   39.8)
)

# ── KEY FIX: convert to sf before passing to getTerraClim ────────
pts_sf <- st_as_sf(sites, coords = c("lon", "lat"), crs = 4326)

# Pull TerraClimate — returns a named list of SpatRasters
tc_raw <- getTerraClim(
  AOI       = pts_sf,
  varname   = c("def", "pet", "ppt"),
  startDate = "1958-01-01",
  endDate   = "2023-12-31"
)

# ── Extract point values from each SpatRaster ─────────────────────
pts_vect <- vect(pts_sf)   # terra needs SpatVector, not sf

extract_var <- function(var_name) {
  r   <- tc_raw[[var_name]]
  ex  <- terra::extract(r, pts_vect)           # rows = sites, cols = time steps
  ex  <- ex[, -1]                              # drop ID column
  
  # Build a time index — one row per month across all years
  n_layers <- nlyr(r)
  yr_start <- 1958
  dates    <- seq(as.Date(paste0(yr_start, "-01-01")), by = "month",
                  length.out = n_layers)
  
  # Reshape: sites as rows → pivot to long format
  ex_t <- as.data.frame(t(ex))                # now rows = months, cols = sites
  colnames(ex_t) <- sites$site
  ex_t$date     <- dates
  ex_t$variable <- var_name
  
  tidyr::pivot_longer(ex_t, cols = all_of(sites$site),
                      names_to = "site", values_to = "value")
}

tc_long <- bind_rows(lapply(c("def", "pet", "ppt"), extract_var))

# ── Annual summaries + Aridity Index ─────────────────────────────
annual <- tc_long |>
  mutate(year = as.integer(format(date, "%Y"))) |>
  group_by(site, year, variable) |>
  summarise(annual_mm = sum(value, na.rm = TRUE), .groups = "drop") |>
  tidyr::pivot_wider(names_from = variable, values_from = annual_mm) |>
  mutate(aridity_index = ppt / pet)

# Save
write.csv(tc_long, here::here("data/terraclimate_monthly.csv"), row.names = FALSE)
write.csv(annual, here::here("data/terraclimate_annual.csv"),  row.names = FALSE)
```

def is CWD

Old SPEI workflow

merge with dataset

```{r}
#| eval: false
#| echo: false

cwd_ai <- read_csv(here::here("data/terraclimate_annual.csv"))
cwd_ai_sum <- cwd_ai %>% 
  group_by(site) |> 
  summarise(cwd_mean = mean(def),
            ai_mean = mean(aridity_index)) |> 
  mutate(Population = site) |> 
  select(
    Population,
    cwd_mean,
    ai_mean
  )

repro_fill_hi_n_narm_1 <- left_join(repro_fill_hi_n_narm,cwd_ai_sum, by=join_by("Population"))

repro_fill_hi_n_narm <- repro_fill_hi_n_narm_1

write.csv(repro_fill_hi_n_narm, here::here("data/repro_fill_hi_n_narm.csv"))

```


```{r}
#| eval: false

workdir <- getwd()
setwd(workdir)

# create data frame:
Population  <- c("ANGELO", "MCL", "BORR","SEDG","WWP")
longitude    <- c(-123.6302, -121.72841, -122.42114, -120.0545, -119.16882)
latitude     <- c(39.75141, 37.37914, 38.87508, 34.70243, 34.91771)
locations_df <- data.frame(Population, longitude, latitude)

# specify arguments and run function:
d1 <- spec_spei(spei_files = "spei48", start_y = 2015, end_y = 2015, 
                locations = locations_df)

write.csv(d1, "SPEI_Pop.csv")
```

```{r}
#| eval: false
#| echo: false

spei <- read_csv(here::here("data/SPEI_Pop.csv"))
spei <- spei %>% 
  filter(month > 2 & month < 6)


spei_avg <- spei %>% 
  group_by(location_id) %>% 
  summarise(spei_avg = mean(spei48)) %>% 
  mutate(Population = location_id)

spei_avg <- spei_avg[,c("Population","spei_avg")]

repro_fill_hi_n_narm_1 <- left_join(repro_fill_hi_n_narm,spei_avg, by=join_by("Population"))

repro_fill_hi_n_narm <- repro_fill_hi_n_narm_1

repro_fill_hi_n_narm$Veg_Mass_ln <- log(repro_fill_hi_n_narm$Veg_Mass)
repro_fill_hi_n_narm$total_fl_mass_ln <- log(repro_fill_hi_n_narm$total_fl_mass)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm[-(which(repro_fill_hi_n_narm$total_fl_mass_ln %in% c(NA, -Inf))),]

# Create ratio of fl to v
repro_fill_hi_n_narm$RV <- repro_fill_hi_n_narm$total_fl_mass_ln/repro_fill_hi_n_narm$Veg_Mass_ln

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  mutate(rating.n = (
    case_when(
      Rating == "Native" ~ 1
      , Rating == "Non-native" ~ 2
      , Rating == "Limited" ~ 3
      , Rating == "Moderate" ~ 4
      , Rating == "High" ~ 5
    )
  ))

rm(
  rep_fls
  , repro_fill_hi_n_narm_1
  , spei
  , spei_avg
)

write.csv(repro_fill_hi_n_narm, here::here("data/repro_fill_hi_n_narm.csv"))

```

Reproductive dataset is ready to go. Use "repro_fill_hi_n_narm" for reproductive allocation.

Read in cleaned dataset to reduce computing power.
```{r}
#| warning: false
#| message: false
repro_fill_hi_n_narm <- read_csv(here::here("data/repro_fill_hi_n_narm.csv"))

# remove Lupinus succulentus because only 1 datapoint
# Remove three grass individuals because they were doubles or mis-identified
# Remove two grass individuals because they are outliers with high influential, likely weeds

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  filter(!Sp_Code %in% c("LUPSUC"
                         #, "ACMPAR"
                         #,"CAUCOU"
                         )) %>% 
   filter(!Plant_ID %in% c("I8.BRODIA.1"
                          ,"B2.AVEBAR.1"
                          ,"C1.AVEBAR.1"
                          )) %>% 
  filter(!Plant_ID %in% c("J5.BRODIA.1"
                          ,"F5.BROHOR.1"))

unique(repro_fill_hi_n_narm$Sp_Code)
unique(repro_fill_hi_n_narm$Scientific_Name)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
    mutate(Scientific_Name = case_when(
         Scientific_Name == "Bromus hordeaceaus" ~ "Bromus hordeaceus"
        ,TRUE ~ Scientific_Name
        )
    )
```

## Q0: Drought

```{r}
head(repro_fill_hi_n_narm)

anova(lm(total_fl_mass_ln ~ Veg_Mass_ln*Treatment, data = repro_fill_hi_n_narm))
```
```{r}
#| eval: false
sm <- read_csv(here::here("data/Site Soil Moisture.csv"))
str(sm)

sm <- sm |> 
  filter(Site %in% c("OTL-Drought","OTL-Control"))
unique(sm$Site)


library(dplyr)
library(ggplot2)

# Define treatment and block for each shelter
shelter_info <- tibble(
  ID        = c("A",       "B",       "C",       "D",       "E",
                "F",       "G",       "H",       "I",       "J"),
  Treatment = c("Drought", "Drought", "Drought", "Control", "Control",
                "Control", "Control", "Control", "Drought", "Drought"),
  Block     = c("Block 1", "Block 2", "Block 3", "Block 4", "Block 5",
                "Block 1", "Block 2", "Block 3", "Block 4", "Block 5")
)

# Summarise hourly soil moisture to one mean ± SE per shelter
summary_data <- sm %>%
  left_join(shelter_info, by = "ID") %>%
  group_by(ID, Treatment, Block) %>%
  summarise(
    Mean = mean(sm, na.rm = TRUE),
    SE   = sd(sm, na.rm = TRUE) / sqrt(sum(!is.na(sm))),
    .groups = "drop"
  )

ggplot(summary_data, aes(
  x    = Block,
  y    = Mean,
  fill = Treatment
)) +
  geom_col(
    position = position_dodge(width = 0.9),
    width    = 0.8
  ) +
  geom_errorbar(
    aes(ymin = Mean - SE, ymax = Mean + SE),
    position  = position_dodge(width = 0.9),
    width     = 0.2,
    linewidth = 0.8
  ) +
 # geom_text(
  #  aes(label = ID, y = Mean + SE + 0.005),
   # position    = position_dodge(width = 0.9),
  #  size        = 3.5,
  #  show.legend = FALSE
  #) +
  scale_fill_viridis_d(option = "cividis") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))  # starts at 0, small padding at top
  ) +
  labs(
    x    = "Block",
    y    = "Mean Soil Moisture",
    fill = "Treatment"
  ) +
  theme_classic() +
  theme(
    axis.text    = element_text(size = 11),
    axis.title   = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 11)
  )

ggsave(here::here("figs/drought.jpg"), height = 3.5, width = 7)


```


## Q1: 1:1

Q1: How much variation is there among annual California grassland plants in their allometric vegetative and reproductive scaling relationship?

Does this group of species have different allometric scaling slopes?

```{r}
anova(lm(total_fl_mass_ln ~ Veg_Mass_ln*Scientific_Name, data = repro_fill_hi_n_narm))
```

No, the effect of vegetative mass on reproductive mass does not depend on species (p-value = 0.2807. Although species does effect the slope intercept (p-value = <2e-16).

Herbaceous species are thought to have a 1:1 scaling relationship between vegetative biomass and reproductive biomass because their lack of woody tissue. Do any species have slopes significantly diffferent than 1?

```{r}
# ChatGPT Prompt, May 11, 2026: "I have reproductive biomass and vegetative biomass data for about 500 individual plants across 26 species. For each species, I want to know if the scaling relationship between reproductive biomass and vegetative biomass is different than a slope of 1"

# Option 2: SMA Regression to account for error of both x and y axis
# Standardized Major Axis Regression
# install.packages("smatr")
library(smatr)

order(unique(repro_fill_hi_n_narm$Sp_Code))
repro_fill_hi_n_narm %>%
  count(Growth_Habit, sort = TRUE)


repro_fill_hi_n_narm_2 <- repro_fill_hi_n_narm %>% 
  filter(!Sp_Code %in% c("ACMPAR","CAUCOU"))

AIRCAR <- repro_fill_hi_n_narm_2 %>% 
  filter(Sp_Code == "AIRCAR")
a <- sma(total_fl_mass_ln ~ Veg_Mass_ln, data = AIRCAR,
    slope.test = 1)
a$pval
a$groupsummary[["Slope_test_p"]]
coef(a)[["slope"]]
a$groupsummary[["Slope_lowCI"]]
a$groupsummary[["Slope_highCI"]]

AMSMEN <- repro_fill_hi_n_narm %>% 
  filter(Sp_Code == "AMSMEN")

library(ggplot2)

ggplot(AMSMEN, aes(x = Veg_Mass_ln,
                y = total_fl_mass_ln,
                label = Plant_ID)) +

  # points
  geom_point(size = 2) +

  # labels
  geom_text(vjust = -0.5, size = 3) +
  
    # regression line
  geom_smooth(method = "lm",
              se = FALSE,
              color = "blue",
              linewidth = 1) +

  # line with slope = 1 through origin
  geom_abline(intercept = 0,
              slope = 1,
              color = "black",
              linewidth = 1) +

  labs(
    x = "ln(Vegetative mass)",
    y = "ln(Reproductive mass)"
  ) +

  theme_classic()

results <- repro_fill_hi_n_narm_2 %>%
  group_by(Scientific_Name) %>%
  do({
    fit <- sma(total_fl_mass_ln ~ Veg_Mass_ln,
               data = .,
               slope.test = 1)

    tibble(
      slope = coef(fit)[["slope"]],
      lower = fit$groupsummary[["Slope_lowCI"]],
      upper = fit$groupsummary[["Slope_highCI"]],
      intercept = coef(fit)[["elevation"]],
      p_value = fit$groupsummary[["Slope_test_p"]]
    )
  }) %>%
  ungroup %>% 
  mutate(p_adj = p.adjust(p_value, method = "fdr"))

AMSMEN <- AMSMEN %>% 
  select(
    Plant_ID,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Flower_Count,
    Veg_Mass,
    fl_average,
    total_fl_mass
  )
write_csv(AMSMEN,"AMSMEN.csv")

```

Three species have slopes significantly different than 1 (p-value < 0.1)

```{r}
#| echo: false

level_order <- unique(as.factor(results$Scientific_Name))
# Graph a bar plot of each species in order of slope with standard error bars and line at 1

#("ANGELO" = "#440154FF", "MCL" = "#3B528BFF", "BORR" = "#21908CFF", "SEDG" = "#5DC863FF", "WWP" = "#FDE725FF")) + 

# Make species names italic and 45 degrees
# Maybe make bar colors if significant

sigs <- results[which(results$p_adj<0.1),1]

results$columnColor <- "N.S."

results <- arrange(results,desc(slope))

#results$columnColor[which(results$Scientific_Name == "Amsinckia menziesii")] <- "p-value < 0.05"

results$columnColor[which(results$Scientific_Name == "Bromus arenarius")] <- "p-value < 0.1"

results$columnColor[which(results$Scientific_Name == "Bromus diandrus")] <- "p-value < 0.1"

results$columnColor[which(results$Scientific_Name == "Erodium botrys")] <- "p-value < 0.1"

#results$columnColor[which(results$Scientific_Name == "Madia gracilis")] <- "p-value < 0.05"

#results$columnColor[which(results$Scientific_Name == "Micropus californicus")] <- "p-value < 0.05"

ggplot(results, aes(x = factor(Scientific_Name, level=c(level_order)), y = slope
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.1" = "#440154FF")) +
  theme(
    axis.text.x = element_text(face = "italic", angle = 45, vjust = 1, hjust = 1),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "R:V Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp_SMA.jpg"))

rm(level_order
   , sigs
   , AIRCAR
   , AMSMEN
   , a
   , repro_fill_hi_n_narm_2)
```

Fig 2. SMA reproductive scaling slopes of 26 annual grassland species. Black dotted line indicates slope of 1 or a 1:1 relationship of vegetative to reproductive biomass. Teal bars indicate species slope is not significantly different than 1. Purple bars indicate slopes significantly different than 1.

Need to look into AMSMEN. Why so large and disproportionate. See graph above. B8 and H3 seem to be skewing it heavily. Both died in April. Technically produced flowers so they're on the board but got arrested early? Feels  wrong to omit them but their effect is disproportionate.
Same iwth A2 and A4 potentially.

Maybe go check out all AMSMEN...

I'm also curious if southern populations of BRODIA are higher than 1.

```{r}
#| echo: false
#| eval: false

# This creates a graph for presentations. Not yet updated with SMA

ggplot(slopes.df, aes(x = factor(Scientific_Name, level=c(level_order)), y = Veg_Mass_ln.trend
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.05" = "#440154FF")) +
  theme(
    axis.text.x = element_blank(),
    legend.text =element_text(size=12, colour="black"),
    axis.text.y=element_text(size=12, colour="black"),
    axis.title=element_text(size=12,face="bold"),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp_prez.jpg"), width = 7,height = 3.5)
```


## Q2: Grass vs Forb Scaling

Q2: Is there a difference in reproductive allometric scaling between annual forbs and grasses?

Let's separate them anyway since all grasses are not native, then forbs if invasive matters and if site matters. And ask grasses if invasive level matters and if site matters

```{r}
# Use a mixed effects linear regression to reflect experimental structure
mxg <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
              data = repro_fill_hi_n_narm)

summary(mxg)
anova(mxg)
#install.packages("lmerTest")
library(lmerTest)
anova(mxg)

# Check assumptions
sim_res <- simulateResiduals(mxg)
plot(sim_res)

testDispersion(sim_res)
# good on dispersion, p-value high
testOutliers(sim_res)
# good on outliers, p-value high
testUniformity(sim_res)
# Poor on normality but data is robust, as good as it gets

# Check for funnel shape (bad), we want constant variance
plot(fitted(mxg), resid(mxg))
abline(h = 0)
# good

# check for dramatic S-shape (bad)
qqnorm(resid(mxg))
qqline(resid(mxg))
# for the most part hugs the line but is a bit skewy

hist(resid(mxg), breaks = 30)      # see the shape
plot(mxg)                           # residuals vs fitted — look for patterns

lattice::qqmath(ranef(mxg, condVar = TRUE))

# not sure what this does
plot(mxg)

library(DHARMa)

sim_res <- simulateResiduals(fittedModel = mxg, plot = TRUE)

# Individual diagnostic tests
testUniformity(sim_res)      # overall uniformity (like normality)
testDispersion(sim_res)      # over/underdispersion
testOutliers(sim_res)        # outlier detection

# Residuals vs each predictor (important!)
plotResiduals(sim_res, repro_fill_hi_n_narm$Veg_Mass_ln)
plotResiduals(sim_res, repro_fill_hi_n_narm$Growth_Habit)

# From Clade on Oct 5 2026:
# The histogram looks normally distributed with long tails on both sides

# That's actually a pretty benign situation — what you're describing is leptokurtosis (heavy tails), not skewness. The KS test flagged it, but the distribution is symmetric, which means your model isn't systematically over- or under-predicting in one direction.

#Why this happens: Long tails on both sides after log-transforming reproductive mass is very common in plant ecology. A few individuals allocate extremely high or extremely low mass to reproduction relative to what the model predicts — biological outliers rather than data errors.

#Why it's not a big deal here:

#Your fixed effect estimates and standard errors are largely unaffected by symmetric heavy tails
#The bigger concern would be asymmetric skew (biased predictions) or a pattern in residuals vs. fitted values (model misspecification)
#KS tests are sensitive with larger sample sizes and will flag even minor deviations that don't affect inference

#Quick sanity check — look at the residuals vs. fitted:

#plot(mxg)

#If that plot shows a roughly horizontal band of points with no curve or fan shape, your model is well-specified and the heavy tails are just a property of biological variability in your data. That's the more important diagnostic.

#If you want to be thorough for a paper, you can note in your methods that residuals showed mild leptokurtosis but were symmetric, fixed effects were robust to this given the sample size, and that log-transformation was applied to meet linearity and homoscedasticity assumptions. Reviewers rarely push back on this for ecological data.

library(emmeans)

slopes <- emtrends(
  mxg,
  ~ Growth_Habit,
  var = "Veg_Mass_ln"
)

slopes

slopes_df <- as.data.frame(slopes)

library(ggplot2)

ps <- ggplot(slopes_df,
       aes(x = Growth_Habit,
           y = Veg_Mass_ln.trend)) +

  geom_point(size = 3) +

  geom_errorbar(
    aes(ymin = lower.CL,
        ymax = upper.CL),
    width = 0.2
  ) +

 # geom_hline(yintercept = 1,
  #           linetype = "dashed",
   #          color = "black") +

  labs(
    y = "RV Scaling slope",
    x = "Functional Type"
  ) +

  theme_classic()
# These are 95% confidence intervals
ggsave("Fxl_type_slope.jpg")

intercepts <- emmeans(mxg, ~ Growth_Habit,
                      at = list(Veg_Mass_ln = 0))

intercepts.df <- summary(intercepts)

pi <- ggplot(intercepts.df,
       aes(x = Growth_Habit,
           y = emmean)) +

  geom_point(size = 3) +

  geom_errorbar(
    aes(ymin = lower.CL,
        ymax = upper.CL),
    width = 0.2
  ) +

 # geom_hline(yintercept = 1,
  #           linetype = "dashed",
   #          color = "black") +

  labs(
    y = "RV Scaling intercept",
    x = "Functional Type"
  ) +

  theme_classic()
# These are 95% confidence intervals
pi
ggsave("Fxl_type_intercept.jpg")

library(smatr)

fit <- sma(
  total_fl_mass_ln ~ Veg_Mass_ln * Growth_Habit,
  data = repro_fill_hi_n_narm
)

summary(fit)


library(ggeffects)

# Status predictions (2 lines)
pred_habit <- ggpredict(mxg, terms = c("Veg_Mass_ln", "Growth_Habit"))

p2 <- ggplot() +

  # Raw data points
  geom_point(
    data = repro_fill_hi_n_narm,
    aes(x = Veg_Mass_ln,
        y = total_fl_mass_ln,
        color = Growth_Habit),
    alpha = 0.3
  ) +

  # Confidence bands
  geom_ribbon(
    data = pred_habit,
    aes(x = x,
        ymin = conf.low,
        ymax = conf.high,
        fill = group,
        group = group),
    alpha = 0.2,
    color = NA
  ) +

  # Predicted lines
  geom_line(
    data = pred_habit,
    aes(x = x,
        y = predicted,
        color = group,
        group = group),
    linewidth = 1.2
  ) +

  # Shared continuous scale
  scale_color_manual(
    name = "Functional Type",
    values = c(
    "Forb" = "#3B528BFF",
    "Grass" = "#5DC863FF"
  )) +
    scale_fill_manual(
      name = "Functional Type",
      values = c(
    "Forb" = "#3B528BFF",
    "Grass" = "#5DC863FF"
  )) +

  labs(
    x = "Ln(Vegetative Biomass)",
    y = "Predicted ln(Reproductive Biomass)"
  ) +

  theme_classic() +
  theme(legend.position = "bottom")

p2

library(patchwork)

p_all <- (p2 | (ps / pi)) +
  plot_layout(
    widths = c(3, 1))
p_all

ggsave(here::here("figs/F3_fxltype_panel.jpg"), p_all, width = 7, height = 4)

rm(
  fit
  , slopes
  , slopes_df
  , intercepts.df
  , intercepts 
  , mxg
  , sim_res
  , pi
  , ps
  , p_all
)



```

There appears to be a difference between grasses and forbs (forbs have a steeper slope but this interactive effect is not significant). 

SMA regression can't include random effects which are important. SMA still shows forbs have higher slopes and says slopes are different. 

# Grasses

```{r}
# scale cwd and ai
repro_fill_hi_n_narm$cwd_sc <- scale(repro_fill_hi_n_narm$cwd_mean, center = TRUE, scale = FALSE)
repro_fill_hi_n_narm$ai_sc <- scale(repro_fill_hi_n_narm$ai_mean, center = TRUE, scale = FALSE)
repro_fill_hi_n_narm$pp_sc <- scale(repro_fill_hi_n_narm$Precip_mm, center = TRUE, scale = FALSE)
repro_fill_hi_n_narm$pp_dif_sc <- scale(repro_fill_hi_n_narm$precip_dif, center = TRUE, scale = FALSE)

repro_grass <- repro_fill_hi_n_narm %>% 
  filter(Growth_Habit == "Grass")

# When comparing different fixed effects, use REML = false. Then use REML true for model estimates and assumptions. Also use REML false for anovas and likelihood ratio tests (anything model comparison)
# When comparing same fixed effects but different random effects, use REML true

mxrxc <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating*cwd_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr_c <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + cwd_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r_c <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + cwd_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr2_c <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + cwd_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r2_c <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + cwd_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

m_r <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

mxr2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_grass)

m_r2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_grass)

mxc <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*cwd_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

m_c <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + cwd_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

mV <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + 
              (1 | Scientific_Name) +
              (1 | Shelter),
           REML = FALSE,
              data = repro_grass)

mxrxa <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating*ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxa <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*ai_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

m_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + ai_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

mxrxp <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating*pp_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr_p <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + pp_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r_p <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + pp_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr2_p <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + pp_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r2_p <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + pp_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxp <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*pp_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

m_p <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + pp_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_grass)

mxrxpd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating*pp_dif_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr_pd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + pp_dif_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r_pd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + pp_dif_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxr2_pd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + pp_dif_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

m_r2_pd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + pp_dif_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_grass)

mxpd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*pp_dif_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_grass)

m_pd <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + pp_dif_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_grass)


ICtab(mxrxc
      , mxr_c
      , m_r_c 
      , mxr2_c
      , m_r2_c
      , mxr
      , m_r
      , mxr2
      , m_r2
      , mxc
      , m_c
      , mxrxa
      , mxr_a
      , m_r_a 
      , mxr2_a
      , m_r2_a
      , mxa
      , m_a
      , mxrxp
      , mxr_p
      , m_r_p 
      , mxr2_p
      , m_r2_p
      , mxp
      , m_p
      , mxrxpd
      , mxr_pd
      , m_r_pd 
      , mxr2_pd
      , m_r2_pd
      , mxpd
      , m_pd)

# For the supp
ICtab(mV
      , mxr
      , m_r
      , mxr2
      , m_r2
      , mxrxa
      , mxr_a
      , m_r_a 
      , mxr2_a
      , m_r2_a
      , mxa
      , m_a)

# Winners are m_a, mxr2_a
# Runner up is m_r2_a, and m_c

lrtest(mV,m_a)
lrtest(m_a,mxr2_a)

# for estimates now
m_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + ai_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
              data = repro_grass)

# check assumptions of m_a
# Check assumptions
sim_res <- simulateResiduals(m_a)
plot(sim_res)

testDispersion(sim_res)
# good on dispersion, p-value high
testOutliers(sim_res)
# good on outliers, p-value high
testUniformity(sim_res)
# Poor on normality but data is robust, as good as it gets

# Check for funnel shape (bad), we want constant variance
plot(fitted(m_a), resid(m_a))
abline(h = 0)
# good

# check for dramatic S-shape (bad)
qqnorm(resid(m_a))
qqline(resid(m_a))
# for the most part hugs the line but is a bit skewy

hist(resid(m_a), breaks = 30)      # see the shape
plot(m_a)                           # residuals vs fitted — look for patterns

lattice::qqmath(ranef(m_a, condVar = TRUE))

# not sure what this does
plot(m_a)

library(DHARMa)

sim_res <- simulateResiduals(fittedModel = m_a, plot = TRUE)

# Individual diagnostic tests
testUniformity(sim_res)      # overall uniformity (like normality)
testDispersion(sim_res)      # over/underdispersion
testOutliers(sim_res)        # outlier detection

# Residuals vs each predictor (important!)
plotResiduals(sim_res, repro_grass$Veg_Mass_ln)
plotResiduals(sim_res, repro_grass$ai_sc)


# m_a wins
slopes <- emtrends(m_a, ~ ai_sc, var = "Veg_Mass_ln")
intercepts <- emmeans(m_a, ~ ai_sc,
                      at = list(Veg_Mass_ln = 0))

slopes.df <- summary(slopes)

# Use plot from below
library(ggeffects)

# SPEI predictions (3 lines: low, mean, high)
pred_ai <- ggpredict(m_a, terms = c("Veg_Mass_ln", "ai_sc [-1,0,1]"))

library(dplyr)

pred_ai <- as.data.frame(pred_ai) %>%
  mutate(panel = "Aridity Index")

pred_ai$ai_sc <-as.numeric(pred_ai$group) 
pred_ai <- pred_ai %>% 
  mutate(ai_sc =
    case_when(
      ai_sc == 1 ~ -1,
      ai_sc == 2 ~ 0,
      ai_sc == 3 ~ 1
    )
  )

pg <- ggplot() +

  # Raw data points
  geom_point(
    data = repro_grass,
    aes(x = Veg_Mass_ln,
        y = total_fl_mass_ln,
        color = ai_sc),
    alpha = 0.3
  ) +

  # Confidence bands
  geom_ribbon(
    data = pred_ai,
    aes(x = x,
        ymin = conf.low,
        ymax = conf.high,
        fill = ai_sc,
        group = ai_sc),
    alpha = 0.2,
    color = NA
  ) +

  # Predicted lines
  geom_line(
    data = pred_ai,
    aes(x = x,
        y = predicted,
        color = ai_sc,
        group = ai_sc),
    linewidth = 1.2
  ) +

  # Shared continuous scale
  scale_color_viridis_c(
    direction = -1,
    name = "Aridity Index",
    option = "D"
  ) +
  scale_fill_viridis_c(
    direction = -1,
    name = "Aridity Index",
    option = "D",
    guide = "none"   # prevents duplicate legend
  ) +

  labs(
    x = "Ln(Vegetative Biomass)",
    y = "Predicted Ln(Reproductive Biomass)",
    tag = "A"
  ) +

  theme_classic() +
  theme(legend.position = "bottom")


pg

repro_fill_hi_n_narm$group <- repro_fill_hi_n_narm$Growth_Habit

rm(mxrxc
      , mxr_c
      , m_r_c 
      , mxr2_c
      , m_r2_c
      , mxr
      , m_r
      , mxr2
      , m_r2
      , mxc
      , m_c
      , mxrxa
      , mxr_a
      , m_r_a 
      , mxr2_a
      , m_r2_a
      , mxa
      , m_a
      , mxrxp
      , mxr_p
      , m_r_p 
      , mxr2_p
      , m_r2_p
      , mxp
      , m_p
      , mxrxpd
      , mxr_pd
      , m_r_pd 
      , mxr2_pd
      , m_r2_pd
      , mxpd
      , m_pd
      , pred_ai)

```

# Forbs

```{r}
repro_forb <- repro_fill_hi_n_narm %>% 
  filter(Growth_Habit == "Forb")

mxg2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

m_g2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Growth_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)


mxg2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_2 + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE, 
             data = repro_forb)

m_g2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Growth_2 +
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_forb)

m_g2_n <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Growth_2 + Native_Status +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxg2_n <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_2 + Native_Status +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

m_g2xn <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Growth_2*Native_Status +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxn_g2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Native_Status + Growth_2 +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxg2xn <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Native_Status*Growth_2 +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxn_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Native_Status + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

m_n_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Native_Status + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxn <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Native_Status + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

m_n <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Native_Status +
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

mxr_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

m_r_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxr2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

m_r2_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Shelter),
              REML = FALSE,
              data = repro_forb)

mxr <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

m_r <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

mxr2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Rating_2 + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_forb)

m_r2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Rating_2 + 
              (1 | Scientific_Name) +
              (1 | Shelter),
             REML = FALSE,
              data = repro_forb)

mxa <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*ai_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

m_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + ai_sc + 
              (1 | Scientific_Name) +
              (1 | Shelter),
            REML = FALSE,
              data = repro_forb)

mV <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + 
              (1 | Scientific_Name) +
              (1 | Shelter),
           REML = FALSE,
              data = repro_forb)

ICtab(mxg2_a
      , m_g2_a
      , mxg2
      , m_g2
      , m_g2_n
      , mxg2_n
      , m_g2xn
      , mxn_g2
      , mxg2xn
      , mxn_a
      , m_n_a
      , mxn
      , m_n
      , mxr_a
      , m_r_a 
      , mxr2_a
      , m_r2_a
      , mxr
      , m_r
      , mxr2
      , m_r2
      , mxa
      , m_a,
      mV)

# Winners are mV, m_g2, mxa

lrtest(mV,m_g2)
lrtest(mV,mxa)

mV <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + 
              (1 | Scientific_Name) +
              (1 | Shelter),
              data = repro_forb)

# mV wins
summary(mV)

# check assumptions
# check assumptions of m_a
# Check assumptions
sim_res <- simulateResiduals(mV)
plot(sim_res)

testDispersion(sim_res)
# good on dispersion, p-value high
testOutliers(sim_res)
# good on outliers, p-value high
testUniformity(sim_res)
# Poor on normality but data is robust, as good as it gets

# Check for funnel shape (bad), we want constant variance
plot(fitted(mV), resid(mV))
abline(h = 0)
# good

# check for dramatic S-shape (bad)
qqnorm(resid(mV))
qqline(resid(mV))
# for the most part hugs the line but is a bit skewy

hist(resid(mV), breaks = 30)      # see the shape
plot(mV)                           # residuals vs fitted — look for patterns

lattice::qqmath(ranef(mV, condVar = TRUE))

# not sure what this does
plot(m_a)

library(DHARMa)

sim_res <- simulateResiduals(fittedModel = mV, plot = TRUE)

# Individual diagnostic tests
testUniformity(sim_res)      # overall uniformity (like normality)
testDispersion(sim_res)      # over/underdispersion
testOutliers(sim_res)        # outlier detection

# Residuals vs each predictor (important!)
plotResiduals(sim_res, repro_forb$Veg_Mass_ln)


# SPEI predictions (3 lines: low, mean, high)
pred_line <- ggpredict(mV, terms = c("Veg_Mass_ln"))

library(dplyr)

pred_line <- as.data.frame(pred_line)

pf <- ggplot() +

  # Raw data points
  geom_point(
    data = repro_forb,
    aes(x = Veg_Mass_ln,
        y = total_fl_mass_ln),
    alpha = 0.3
  ) +

  # Confidence bands
  geom_ribbon(
    data = pred_line,
    aes(x = x,
        ymin = conf.low,
        ymax = conf.high),
    alpha = 0.2,
    color = NA
  ) +

  # Predicted lines
  geom_line(
    data = pred_line,
    aes(x = x,
        y = predicted),
    linewidth = 1.2
  ) +

  labs(
    x = "Ln(Vegetative Biomass)",
    y = "Predicted Ln(Reproductive Biomass)",
    tag = "B"
  ) +

  theme_classic() +
  theme(legend.position = "bottom")

 pf

p3 <- pg | pf
p3

# Go back and make another map of CA SPEI, maybe even side by side with precip

ggsave(here::here("figs/model_slopes_1.jpg"),p3, width = 7, height = 4)

rm(mxg2_a
      , m_g2_a
      , mxg2
      , m_g2
      , m_g2_n
      , mxg2_n
      , m_g2xn
      , mxn_g2
      , mxg2xn
      , mxn_a
      , m_n_a
      , mxn
      , m_n
      , mxr_a
      , m_r_a 
      , mxr2_a
      , m_r2_a
      , mxr
      , m_r
      , mxr2
      , m_r2
      , mxa
      , m_a,
      mV)

```

Considered including phylogenetics in model but tried using phyr but 29 out of 30 species were not found in the phylogeny

## Q4: Traits

Q4: How do aboveground traits associate with reproductive allocation and reproductive scaling?

First read in all the traits we need and clean them up

-   Relative Growth Rate (leaf, height, volume)
-   Photosynthetic Rate
-   Leaf Mass Area
-   Leaf Area
-   WUE instantaneous
-   WUE intrinsic
-   Height
-   Leaf Nitrogen
-   C:N
-   Date of first flower
-   Seed Mass

-   Mid season biomass partial
-   Root biomass partial
-   Root:shoot ratio partial
-   Root diameter partial
-   Specific root length partial
-   Root nitrogen partial

## Relative Growth Rate

### Clean up dataset
```{r}
#| eval: false
#| echo: false
rgr <- read_csv(here::here("data/Relative Growth Rate 2023.csv"))
plots <- read_csv(here::here("data/Plot_info.csv"))
str(rgr)
summary(rgr)
head(rgr)
unique(rgr$Month)

# All IDs from May and June are missing ".1" at the end
# Extract species code from ID
# Extract plot, shelter letter from ID, add in block and population
rgr2 <- rgr %>% 
  mutate(Plant_ID = if_else(Month == "May" | Month == "June", true = paste0(Plant_ID, ".1"), false = Plant_ID),
         Sp_Code = substring(Plant_ID, 4,9),
         Plot = substring(Plant_ID, 1,2),) %>% 
  filter(!Plant_ID == "E2.PHATAN.1B") %>% 
  left_join(plots,by = join_by(Plot))
rgr[which(rgr$Plant_ID == "E2.PHATAN.1B"),]
rgr2[which(rgr2$Plant_ID == "E2.PHATAN.1B"),]
str(rgr2)  

unique(rgr2$Sp_Code)
unique(rgr2$Shelter)
unique(rgr2$Plot)

# Find gaps, investigate
# Fixed, entering errors besides A1.ACMPAR in May which is missing leaf count

incomplete_rows <- rgr2[,c("Plant_ID","Month","Height_cm","Leaf_number","canopy_x","canopy_y")]
incomplete_rows_2 <- incomplete_rows[!complete.cases(incomplete_rows),]
incomplete_rows_2 <- incomplete_rows_2[rowSums(is.na(incomplete_rows_2)) < 4,] 

# Find and investigate outliers
# Good job for interns
# Find average and standard deviation for each species, each growth variable (make area first), per population per month
str(rgr2)
rgr2$area <- as.numeric(rgr2$canopy_x)*as.numeric(rgr2$canopy_y)

rgr_avgs <- rgr2 %>% 
  group_by(Sp_Code,
           Month,
           Population) %>% 
  mutate(ht_avg = mean(Height_cm, na.rm = TRUE),
            ht_sd = sd(Height_cm, na.rm = TRUE),
            lf_avg = mean(Leaf_number, na.rm = TRUE),
            lf_sd = sd(Leaf_number, na.rm = TRUE),
            area_avg = mean(area, na.rm = TRUE),
            area_sd = sd(area, na.rm = TRUE),
            x_avg = mean(canopy_x, na.rm = TRUE),
            x_sd = sd(canopy_x, na.rm = TRUE),
            y_avg = mean(canopy_y, na.rm = TRUE),
            y_sd = sd(canopy_y, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Month,
         Population,
         Height_cm,
         ht_avg,
         ht_sd,
         Leaf_number,
         lf_avg,
         lf_sd,
         area,
         area_avg,
         area_sd,
         canopy_x,
         canopy_y,
         x_avg,
         x_sd,
         y_avg,
         y_sd)

rgr_avgs <- rgr_avgs[complete.cases(rgr_avgs),]

rgr_avgs <- rgr_avgs %>% 
  mutate(ht_max = ht_avg + 3*ht_sd,
         ht_min = ht_avg - 3*ht_sd,
         lf_max = lf_avg + 3*lf_sd,
         lf_min = lf_avg - 3*lf_sd,
         area_max = area_avg + 3*area_sd,
         area_min = area_avg - 3*area_sd,
         x_max = x_avg + 3*x_sd,
         x_min = x_avg - 3*x_sd,
         y_max = y_avg + 3*y_sd,
         y_min = y_avg - 3*y_sd)

ht_outliers <- rgr_avgs[which(rgr_avgs$Height_cm > rgr_avgs$ht_max | rgr_avgs$Height_cm < rgr_avgs$ht_min),c("Plant_ID","Sp_Code","Month","Population", "Height_cm","ht_avg","ht_sd","ht_min","ht_max")]

lf_outliers <- rgr_avgs[which(rgr_avgs$Leaf_number > rgr_avgs$lf_max | rgr_avgs$Leaf_number < rgr_avgs$lf_min),c("Plant_ID","Sp_Code","Month","Population", "Leaf_number","lf_avg","lf_sd","lf_min","lf_max")]

area_outliers <- rgr_avgs[which(rgr_avgs$area > rgr_avgs$area_max | rgr_avgs$area < rgr_avgs$area_min),c("Plant_ID","Sp_Code","Month","Population", "area","area_avg","area_sd","area_min","area_max")]

x_outliers <- rgr_avgs[which(rgr_avgs$canopy_x > rgr_avgs$x_max | rgr_avgs$canopy_x < rgr_avgs$x_min),c("Plant_ID","Sp_Code","Month","Population", "canopy_x","x_avg","x_sd","x_min","x_max")]

y_outliers <- rgr_avgs[which(rgr_avgs$canopy_y > rgr_avgs$y_max | rgr_avgs$canopy_y < rgr_avgs$y_min),c("Plant_ID","Sp_Code","Month","Population", "canopy_y","y_avg","y_sd","y_min","y_max")]

## Now go and investigate outliers using original data
```

### Create variables
```{r}
#| eval: false

# Calculate canopy volume (height x canopy area)
head(rgr2)
rgr2$volume <- rgr2$Height_cm*rgr2$canopy_x*rgr2$canopy_y
rgr2$Date <- as.Date(rgr2$Date, tryFormats = "%m/%d/%y")
str(rgr2)

# Separate by metric

rgr_h_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         Height_cm) %>% 
  filter(!is.na(Height_cm)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, Height_cm)) %>% 
  mutate(rgr_h_FM = (log(Height_cm_March)-log(Height_cm_February))/as.numeric((Date_March-Date_February)),
         rgr_h_MA = (log(Height_cm_April)-log(Height_cm_March))/as.numeric((Date_April-Date_March)),
         rgr_h_AM = (log(Height_cm_May)-log(Height_cm_April))/as.numeric((Date_May-Date_April)),
         rgr_h_MJ = (log(Height_cm_June)-log(Height_cm_May))/as.numeric((Date_June-Date_May)))

rgr_l_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         Leaf_number) %>% 
  filter(!is.na(Leaf_number)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, Leaf_number)) %>% 
  mutate(rgr_l_FM = (log(Leaf_number_March)-log(Leaf_number_February))/as.numeric((Date_March-Date_February)),
         rgr_l_MA = (log(Leaf_number_April)-log(Leaf_number_March))/as.numeric((Date_April-Date_March)),
         rgr_l_AM = (log(Leaf_number_May)-log(Leaf_number_April))/as.numeric((Date_May-Date_April)),
         rgr_l_MJ = (log(Leaf_number_June)-log(Leaf_number_May))/as.numeric((Date_June-Date_May)))

rgr_v_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         volume) %>% 
  filter(!is.na(volume)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, volume)) %>% 
  mutate(rgr_v_FM = (log(volume_March)-log(volume_February))/as.numeric((Date_March-Date_February)),
         rgr_v_MA = (log(volume_April)-log(volume_March))/as.numeric((Date_April-Date_March)),
         rgr_v_AM = (log(volume_May)-log(volume_April))/as.numeric((Date_May-Date_April)),
         rgr_v_MJ = (log(volume_June)-log(volume_May))/as.numeric((Date_June-Date_May)))
  
# Join all three together but only keep rgrs, full join by Plant_ID

rgr_h_l <- full_join(rgr_h_wide,rgr_l_wide,by = join_by(Plant_ID))

rgr_hlv <- full_join(rgr_h_l,rgr_v_wide,by = join_by(Plant_ID))

rgr_hlv2 <- rgr_hlv %>% 
  select(
    Plant_ID,
    rgr_h_FM,
    rgr_h_MA,
    rgr_h_AM,
    rgr_h_MJ,
    rgr_l_FM,
    rgr_l_MA,
    rgr_l_AM,
    rgr_l_MJ,
    rgr_v_FM,
    rgr_v_MA,
    rgr_v_AM,
    rgr_v_MJ
  )

## Wooooo we did it!
## Some values have -Inf, investigate that at some point. For now you're done!

write_csv(rgr_hlv2,here::here("data/RGR_clean.csv"))
```

### Exploratory Graphs

We have height, leaf number, area, and volume

```{r}
rgr_hlv2 <- read_csv(here::here("data/RGR_clean.csv"))

rgr_hlv2_long <- rgr_hlv2 %>% 
  pivot_longer(cols = c("rgr_h_FM":"rgr_v_MJ"),
               names_to = "var_mo",
               values_to = "rate") %>%
  filter(!is.na(rate))

rgr_sum <- rgr_hlv2_long %>%
  filter(!rate %in% c("-Inf","NaN","Inf")) %>% 
  group_by(var_mo) %>% 
  summarise(avg = mean(rate),
            stdev = sd(rate)) %>% 
  arrange(desc(avg))
```

Highest growth rate for volume is Feb-Mar (0.0957), then Mar-Apr (0.0827). For height, highest growth rate is Mar-Apr (0.0405), then Feb-Mar (0.0367). For leaves, highest growth rate is 0.0291 in Feb-Mar, then May-Jun 0.0198 (seems wrong). 

Remember we didn't know who each grass was in February. So might need to throw those out OR separate by growth habit.

### Select Reference times

One predictive variable of allometric slope might be relative growth rate timing and/or maximum relative growth rate. 

### Explore RGR

```{r}
#| eval: false

# parse out plant ID for species, plot, Population
plots <- read_csv(here::here("data/Plot_info.csv"))
str(plots)
summary(plots)
head(plots)

spp <- read_csv(here::here("data/Species Codes.csv"))
unique(spp$Sp_Code)

# need to pull out first two characters in string of plant ID to create a "plot" column

rgr_hlv <- rgr_hlv2 %>% 
  mutate(Plot = substr(Plant_ID, start = 1, stop = 2),
         Species = substr(Plant_ID, start = 4, stop = 9))
unique(rgr_hlv$Plot)
unique(rgr_hlv$Species)

rgr_hlv_full <- left_join(rgr_hlv,plots,by = join_by(Plot))

rgr_hlv_full$Sp_Code <- rgr_hlv$Species

rgr_hlv_full <- left_join(rgr_hlv_full,spp,by = join_by(Sp_Code))

rgr_h <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_h_FM,
    rgr_h_MA,
    rgr_h_AM,
    rgr_h_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_h_FM":"rgr_h_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_h,here::here("data/RGR_height.csv"))

rgr_l <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_l_FM,
    rgr_l_MA,
    rgr_l_AM,
    rgr_l_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_l_FM":"rgr_l_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_l,here::here("data/RGR_leaves.csv"))

rgr_v <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_v_FM,
    rgr_v_MA,
    rgr_v_AM,
    rgr_v_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_v_FM":"rgr_v_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_v,here::here("data/RGR_volume.csv"))

```

Start graphing, what do I want to see? growth rate by rating or native status across months

```{r}
#| message: false
#| warning: false
#| echo: false
# Order Invasive rating: Native, Non-native, Limited, Moderate, High
# Order Month rates
rgr_h <- read_csv(here::here("data/RGR_height.csv"))
rgr_l <- read_csv(here::here("data/RGR_leaves.csv"))
rgr_v <- read_csv(here::here("data/RGR_volume.csv"))

rgr_h$Rating<- factor(rgr_h$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_h$var_mo<- factor(rgr_h$var_mo, levels = c("rgr_h_FM", "rgr_h_MA","rgr_h_AM","rgr_h_MJ"))

# Change violin plot colors by groups
ggplot(rgr_h, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()


rgr_l$Rating<- factor(rgr_l$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_l$var_mo<- factor(rgr_l$var_mo, levels = c("rgr_l_FM", "rgr_l_MA","rgr_l_AM","rgr_l_MJ"))

ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

rgr_v$Rating<- factor(rgr_v$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_v$var_mo<- factor(rgr_v$var_mo, levels = c("rgr_v_FM", "rgr_v_MA","rgr_v_AM","rgr_v_MJ"))

ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

# Recreate species x slope graph




```

Should almost look at each of these graphs by population, so it's limited to community. It's possible an invasive from Angelo has the same growth rate as a native from santa barbara.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: false
rgr_h_ACR <- rgr_h %>% 
  filter(Population == "ANGELO")

rgr_h_MCL <- rgr_h %>% 
  filter(Population == "MCL")

rgr_h_BORR <- rgr_h %>% 
  filter(Population == "BORR")

rgr_h_SEDG <- rgr_h %>% 
  filter(Population == "SEDG")

# Change violin plot colors by groups
ggplot(rgr_h_ACR, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_ACR, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_MCL, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_MCL, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_BORR, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_BORR, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_SEDG, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_SEDG, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()



ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()
```
These graphs should give us clues about invasive coverage and niche dynamics and competition within each plot.

Now curious how these relate to reproductive allocation/scaling

Remember we didn't know who each grass was in February. So might need to throw those out OR separate by growth habit.

## Photosynthetic Rate

```{r}
#| eval: false
#| echo: false
#| warning: false
#| message: false
# how to tell it the header and which rows to skip
# how to read through the filenames in a certain folder and make a list and then read those in 1 by 1
getwd()
licor_files <- list.files("/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/licor")
licor_files[4]
licor_files_path <- paste0("data/licor/",licor_files)
licor_files_path[4]

trial <- read_csv(here::here(licor_files_path[4]), skip = 1, col_select = "obs":"gtc")

lic <- list() # creates a list
for (k in 1:length(licor_files_path)){
 lic[[k]] <- read_csv(here::here(licor_files_path[k]), skip = 1, col_select = "obs":"gtc")
}

lic <- lapply(lic, function(tbl) {
  tbl$WUE <- as.numeric(tbl$WUE)
  return(tbl)
})

licor <- bind_rows(lic)
licor <- licor %>% 
  filter(!is.na(obs))

write_csv(licor,here::here("data/licor_clean.csv"))

```

Now I need to bring in Licor leaf area csv and join. Also need to double check numbers for each species.

GALPAR, Holvir, miccal not working -- fixed, just needed to adjust minimum size.

```{r}
#| eval: false
#| echo: false
acmpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ACMPAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
acmpar<-acmpar$summary

aircar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AIRCAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
aircar<-aircar$summary

amsmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AMSMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
amsmen<-amsmen$summary

avebar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AVEBAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
avebar<-avebar$summary

broare = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROARE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
broare<-broare$summary

brodia = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BRODIA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brodia<-brodia$summary

brohor = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROHOR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brohor<-brohor$summary

bromad = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROMAD",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
bromad<-bromad$summary

calmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CALMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
calmen<-calmen$summary

caucou = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CAUCOU",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
caucou<-caucou$summary

cenmel = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CENMEL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
cenmel<-cenmel$summary

clapur = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CLAPUR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
clapur<-clapur$summary

corfil = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CORFIL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
corfil<-corfil$summary

daupus = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/DAUPUS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
daupus<-daupus$summary

elycap = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ELYCAP",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
elycap<-elycap$summary

erobot = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/EROBOT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
erobot<-erobot$summary

erocic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/EROCIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erocic<-erocic$summary

esccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ESCCAL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
esccal<-esccal$summary


galpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/GALPAR", 
                distance.pixel = 355, 
                known.distance = 3,
                low.size = 0.005,
                log = TRUE)
          #      save.image = TRUE)
galpar<-galpar$summary

gerdis = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/GERDIS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
gerdis<-gerdis$summary

holvir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/HOLVIR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
                # save.image = TRUE)
holvir<-holvir$summary

hypgla = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/HYPGLA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
hypgla<-hypgla$summary

lupbic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPBIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
lupbic<-lupbic$summary

lupmic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPMIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupmic<-lupmic$summary

lupsuc = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPSUC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupsuc<-lupsuc$summary

madgra = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/MADGRA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
madgra<-madgra$summary

miccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/MICCAL", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
miccal<-miccal$summary

phatan = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/PHATAN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
phatan<-phatan$summary

plaere = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/PLAERE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
plaere<-plaere$summary

stipul = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/STIPUL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
stipul<-stipul$summary

torarv = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/TORARV",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
torarv<-torarv$summary

trihir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/TRIHIR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
trihir<-trihir$summary

vicsat = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/VICSAT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
vicsat<-vicsat$summary

# missing galpar, miccal, holvir
leaf_licor_area <- rbind(acmpar,
                         aircar,
                         amsmen,
                         avebar,
                         broare,
                         brodia,
                         brohor,
                         bromad,
                         calmen,
                         caucou,
                         cenmel,
                         clapur,
                         corfil,
                         daupus,
                         elycap,
                         erobot,
                         erocic,
                         esccal,
                         galpar,
                         gerdis,
                         holvir,
                         hypgla,
                         lupbic,
                         lupmic,
                         lupsuc,
                         madgra,
                         miccal,
                         phatan,
                         plaere,
                         stipul,
                         torarv,
                         trihir,
                         vicsat)

write_csv(leaf_licor_area, here::here("data/leaf_licor_area.csv"))
```

Double check that the numbers of values per species aligns with number of images per species folder. -- All good

Then merge with Licor dataset by plant ID.

```{r}
#| eval: false

# Separate date from plant ID, make in date format
leaf_licor_area <- read_csv(here::here("data/leaf_licor_area.csv"))
head(leaf_licor_area)
licor <- read_csv(here::here("data/licor_clean.csv"))
head(licor)

licor$Date <- as.Date(substr(licor$date, start=1, stop=10))
str(licor$Date)

leaf_licor_area$Date <- as.Date(substr(leaf_licor_area$sample, start=11, stop=18), "%Y%m%d")
str(licor$Date)

# separate plant ID from same and add periods to plant id
leaf_licor_area$Plant_ID <- substr(leaf_licor_area$sample, start=1,stop=9)
leaf_licor_area$Plant_ID <- paste0(
  substr(leaf_licor_area$Plant_ID, 1, 2),
  ".",
  substr(leaf_licor_area$Plant_ID, 3, nchar(leaf_licor_area$Plant_ID))
)
leaf_licor_area$Plant_ID

leaf_licor_area$Plant_ID <- paste0(
  substr(leaf_licor_area$Plant_ID, 1, 9),
  ".",
  substr(leaf_licor_area$Plant_ID, 10, nchar(leaf_licor_area$Plant_ID))
)
leaf_licor_area$Plant_ID

# capitalize Licor plant id
licor$Plant_ID <- toupper(licor$`plant id`)

# NAs to fill in
# C6.PHATAN.1 is missing, no photo
# H1.AVEBAR.1 is missing, no photo
# G8.VICSAT.2 is missing, no photo
# A2.BROMAD.2 is missing, no photo
# H1.BROMAD.2 is missing, no photo
# C7.ESCCAL.2 is missing, no photo
# D4.TRIHIR.2 is missing, no photo
# C4.TRIHIR.1 is missing, no photo
# A6.STIPUL.2 is missing, no photo
# D7.CORFIL.2 is missing, no photo

# Now I need to interpolate leaf area for the missing ones
leaf_licor_area$Species <- substr(leaf_licor_area$Plant_ID,4,9)
lf_area_avg <- leaf_licor_area %>% 
  group_by(Species) %>% 
  summarise(lf_avg = mean(total.leaf.area))

lf_NAs <- data.frame(
  Plant_ID = c(
    "C6.PHATAN.1",
    "H1.AVEBAR.1",
    "G8.VICSAT.2",
    "A2.BROMAD.2",
    "H1.BROMAD.2",
    "C7.ESCCAL.2",
    "D4.TRIHIR.2",
    "C4.TRIHIR.1",
    "A6.STIPUL.2",
    "D7.CORFIL.2"
  ))

lf_NAs$Species <- substr(lf_NAs$Plant_ID,4,9)
lf_NAs_fill <- left_join(lf_NAs,lf_area_avg,by = c("Species" = "Species"))

str(leaf_licor_area)
str(lf_NAs_fill)

lf_NAs_fill$Date <- NA
lf_NAs_fill$sample <- NA
lf_NAs_fill <- lf_NAs_fill %>% 
  mutate(total.leaf.area = lf_avg) %>% 
  select(
    sample,
    total.leaf.area,
    Date,
    Plant_ID,
    Species
  )

leaf_licor_area <- rbind(leaf_licor_area, lf_NAs_fill)
### Should I have actually done this by species x population


# merge with licor leaf area
licor2 <- left_join(licor,leaf_licor_area,by=c("Plant_ID" = "Plant_ID"))

# Come back and investigate NAs for licor2 -- why are some leaf areas missing
licor.na <- licor2 %>% 
 filter(is.na(total.leaf.area)) %>% 
  select(
    date
    , A
    , `plant id`
    , Plant_ID
  )

licor2$E_mmol <- licor2$E*1000

licor2$WUE_new <- licor2$A/licor2$E_mmol

# multiply focal data by leaf area fraction
licor2$leaf_ratio <- 6/licor2$total.leaf.area
licor2 <- licor2 %>% 
  filter(!is.na(total.leaf.area)) %>% # filter NAs for now
  mutate(WUE_adj = leaf_ratio*WUE_new,
         E_adj = leaf_ratio*as.numeric(E),
         A_adj = leaf_ratio*as.numeric(A),
         Ci_adj = leaf_ratio*as.numeric(Ci),
         gsw_adj = leaf_ratio*as.numeric(gsw))
str(licor2)

# remove licor data with gsw below threshold
# licor2 2315
licor2_b <- licor2[-which(licor2$gsw_adj < 0.03),]

# test for licor outliers by species, population, and treatment
# Maybe I do this later though? Because what if an outlier in photosynthetic rate goes with a reproductive outlier?

# average licor data by plant_id
licor3 <- licor2_b %>% 
  group_by(Plant_ID) %>% 
  summarise(A_adj_av = mean(A_adj, na.rm=TRUE),
            WUE_adj_av= mean(WUE_adj, na.rm = TRUE),
            E_adj_av=mean(E_adj, na.rm = TRUE),
            Ci_adj_av = mean(Ci_adj, na.rm = TRUE),
            gsw_adj_av = mean(gsw_adj, na.rm = TRUE))

# extract species and plot from plant id
licor3$Sp_Code <- substr(licor3$Plant_ID,4,9)
licor3$Plot <- substr(licor3$Plant_ID,1,2)

# merge with plot and species info
plots <- read_csv(here::here("data/Plot_info.csv"))
licor4 <- left_join(licor3,plots, by = join_by(Plot))
spp <- read_csv(here::here("data/Species Codes.csv"))
licor5 <- left_join(licor4,spp, by = join_by(Sp_Code))

write_csv(licor5,here::here("data/licor5.csv"))
```

### Explore licor

```{r}
#| echo: false
#| message: false
#| warning: false

licor5 <- read_csv(here::here("data/licor5.csv"))

summary(licor5)
hist(licor5$WUE_adj_av)
licor5[which(licor5$WUE_adj_av > 50),"Plant_ID"]

licor5b <- licor5 %>% 
  filter(!WUE_adj_av > 50)

licor5 <- licor5b

# okay some things are way off with WUE, and others -- will need to deal with outliers at some point

p <- ggplot(licor5, aes(x=Sp_Code, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

p <- ggplot(licor5, aes(x=Rating, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

p <- ggplot(licor5, aes(x=Native_Status, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

```

At some point I need to do a correlation table with all the traits.

## Leaf Mass Area
```{r}
#| eval: false
#| echo: false
acmpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/ACMPAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
acmpar<-acmpar$summary

aircar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AIRCAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
aircar<-aircar$summary

amsmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AMSMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
amsmen<-amsmen$summary

avebar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AVEBAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
avebar<-avebar$summary

broare = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROARE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
broare<-broare$summary

brodia = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BRODIA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brodia<-brodia$summary

brohor = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROHOR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brohor<-brohor$summary

bromad = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROMAD",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
bromad<-bromad$summary

calmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CALMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
calmen<-calmen$summary

caucou = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CAUCOU",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
caucou<-caucou$summary

cenmel = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CENMEL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
cenmel<-cenmel$summary

clapur = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CLAPUR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
clapur<-clapur$summary

daupus = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/DAUPUS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
daupus<-daupus$summary

elycap = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/ELYCAP",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
elycap<-elycap$summary

erobot = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/EROBOT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erobot<-erobot$summary

erocic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/EROCIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erocic<-erocic$summary

galpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/GALPAR", 
                distance.pixel = 355, 
                known.distance = 3,
                low.size = 0.005,
                log = TRUE)
          #      save.image = TRUE)
galpar<-galpar$summary

gerdis = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/GERDIS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
gerdis<-gerdis$summary

holvir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/HOLVIR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
                # save.image = TRUE)
holvir<-holvir$summary

hypgla = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/HYPGLA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
hypgla<-hypgla$summary

lupbic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPBIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
lupbic<-lupbic$summary

lupmic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPMIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupmic<-lupmic$summary

lupsuc = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPSUC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupsuc<-lupsuc$summary

madgra = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/MADGRA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
madgra<-madgra$summary

miccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/MICCAL", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
miccal<-miccal$summary

phatan = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/PHATAN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
phatan<-phatan$summary

plaere = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/PLAERE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
plaere<-plaere$summary

torarv = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/TORARV",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
torarv<-torarv$summary

trihir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/TRIHIR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
trihir<-trihir$summary

vicsat = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/VICSAT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
vicsat<-vicsat$summary

# missing galpar, miccal, holvir
lma_lf_area <- rbind(acmpar,
                         aircar,
                         amsmen,
                         avebar,
                         broare,
                         brodia,
                         brohor,
                         bromad,
                         calmen,
                         caucou,
                         cenmel,
                         clapur,
                         daupus,
                         elycap,
                         erobot,
                         erocic,
                         galpar,
                         gerdis,
                         holvir,
                         hypgla,
                         lupbic,
                         lupmic,
                         lupsuc,
                         madgra,
                         miccal,
                         phatan,
                         plaere,
                         torarv,
                         trihir,
                         vicsat)

write_csv(lma_lf_area, here::here("data/LMA_lf_area.csv"))


```


```{r}

lf_area <- read_csv(here::here("data/LMA_lf_area.csv"))
# There may be errors in this -- not sure how to check

## Extract plant id, species, plot
lf_area$Date <- as.Date(substr(lf_area$sample, start=11, stop=18), "%Y%m%d")

# separate plant ID from same and add periods to plant id
lf_area$Plant_ID <- substr(lf_area$sample, start=1,stop=9)
lf_area$Plant_ID <- paste0(
  substr(lf_area$Plant_ID, 1, 2),
  ".",
  substr(lf_area$Plant_ID, 3, nchar(lf_area$Plant_ID))
)
lf_area$Plant_ID

lf_area$Plant_ID <- paste0(
  substr(lf_area$Plant_ID, 1, 9),
  ".",
  substr(lf_area$Plant_ID, 10, nchar(lf_area$Plant_ID))
)
lf_area$Plant_ID

lf_area$Sp_Code <- substr(lf_area$Plant_ID,4,9)


# merge with plot and species info
#lf_area2 <- left_join(lf_area,plots, by = join_by(Plot))
#lf_area3 <- left_join(lf_area2,spp, by = join_by(Sp_Code))

# Now merge with leaf mass and calculate LMA
mass <- read_csv(here::here("data/Leaf Mass_2023_20231009.csv"))
head(mass)

plots <- read_csv(here::here("data/Plot_info.csv"))
spp <- read_csv(here::here("data/Species Codes.csv"))

# Remove perennials
mass_a <- mass %>% 
  filter(!Sp_Code %in% c("CORFIL","ESCCAL","STIPUL","DANCAL"))

# 696 masses
# 691 leaf areas

# I want to know how many observations per species for each dataset, start there

library(dplyr)
ct_area <- lf_area %>% count(Sp_Code)
ct_mass <- mass_a %>% count(Sp_Code)

join_check <- full_join(lf_area,mass_a,by=join_by("Plant_ID"))
join_check %>% filter(is.na(total.leaf.area))
join_check %>% filter(is.na(Dry_leaf_mass_g))
# Missing:
# H1.AVEBAR.2 area
# H3.BRODIA.2 Mass
# G8.VICSAT.2 area

str(lf_area)
str(mass_a)

LMA <- full_join(lf_area,mass_a, by = join_by(Plant_ID, Sp_Code == Sp_Code))
# LMA is measured in g/m^2

LMA$area_m2 <- LMA$total.leaf.area/10000
LMA$lma_g_m2 <- LMA$area_m2/LMA$Dry_leaf_mass_g

summary(LMA)
LMA[which(is.na(LMA$Dry_leaf_mass_g)),"Plant_ID"]
LMA[which(is.na(LMA$total.leaf.area)),"Plant_ID"]
# interpolate LMA based on spp x population averages
# First make sure they are NA so they don't effect averages
LMA$lma_g_m2[which(LMA$Plant_ID == "H1.AVEBAR.2")] <- NA
LMA$lma_g_m2[which(LMA$Plant_ID == "H3.BRODIA.2")] <- NA
LMA$lma_g_m2[which(LMA$Plant_ID == "G8.VICSAT.2")] <- NA
# H1.AVEBAR.2 area
# H3.BRODIA.2 Mass
# G8.VICSAT.2 area
LMA[which(is.na(LMA$lma_g_m2)),"Plant_ID"]

# merge with plot and species info
LMA$Plot <- substr(LMA$Plant_ID,1,2)
LMA2 <- left_join(LMA,plots, by = join_by(Plot))
LMA <- left_join(LMA2,spp, by = join_by(Sp_Code))

pop_spp_avg <- LMA %>% 
  group_by(Sp_Code,Population) %>% 
  summarise(avg_lma = mean(lma_g_m2,na.rm = TRUE))

pop <- LMA[which(LMA$Plant_ID == "H1.AVEBAR.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "AVEBAR" & pop_spp_avg$Population == "SEDG"),"avg_lma"]
LMA[which(LMA$Plant_ID == "H1.AVEBAR.2"),"lma_g_m2"] <- avg

pop <- LMA[which(LMA$Plant_ID == "H3.BRODIA.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "BRODIA" & pop_spp_avg$Population == "WWP"),"avg_lma"]
LMA[which(LMA$Plant_ID == "H3.BRODIA.2"),"lma_g_m2"] <- avg

pop <- LMA[which(LMA$Plant_ID == "G8.VICSAT.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "VICSAT" & pop_spp_avg$Population == "BORR"),"avg_lma"]
LMA[which(LMA$Plant_ID == "G8.VICSAT.2"),"lma_g_m2"] <- avg

LMA[which(is.na(LMA$lma_g_m2)),"Plant_ID"]

# Check for duplicates
LMA[which(duplicated(LMA$Plant_ID)),]


```

### Explore LMA
```{r}
#| warning: false
summary(LMA)
# okay some things are way off with WUE, and others -- will need to deal with outliers at some point

ggplot(LMA, aes(x=Sp_Code, y=lma_g_m2)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )


ggplot(LMA, aes(x=Rating, y=lma_g_m2)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

```

## Leaf Area

```{r}
#| eval: false

galpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Single Leaf Area/GALPAR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
galpar<-galpar$summary

holvir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Single Leaf Area/HOLVIR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
holvir<-holvir$summary

miccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Single Leaf Area/MICCAL", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
miccal<-miccal$summary

singles <- rbind(galpar,holvir,miccal)

lf_area_rm <- lf_area %>% 
  filter(!Sp_Code %in% c("GALPAR","HOLVIR","MICCAL"))

str(lf_area_rm)
str(singles)

## Extract plant id, species, plot
singles$Date <- as.Date(substr(singles$sample, start=11, stop=18), "%Y%m%d")

# separate plant ID from same and add periods to plant id
singles$Plant_ID <- substr(singles$sample, start=1,stop=9)
singles$Plant_ID <- paste0(
  substr(singles$Plant_ID, 1, 2),
  ".",
  substr(singles$Plant_ID, 3, nchar(singles$Plant_ID))
)
singles$Plant_ID

singles$Plant_ID <- paste0(
  substr(singles$Plant_ID, 1, 9),
  ".",
  substr(singles$Plant_ID, 10, nchar(singles$Plant_ID))
)
singles$Plant_ID

singles$Sp_Code <- substr(singles$Plant_ID,4,9)

lf_area <- rbind(lf_area_rm,singles)

lf_area$Plot <- substr(lf_area$Plant_ID,1,2)

write_csv(lf_area,here::here("data/single_lf_area.csv"))

```


```{r}
lf_area <- read_csv(here::here("data/single_lf_area.csv"))

summary(lf_area)

ggplot(lf_area, aes(x=Sp_Code, y=total.leaf.area)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
```

## WUE

I believe we can get WUE from Licor and from the CN data. 

```{r}

summary(licor5)

WUE <- licor5 %>% 
  select(Plant_ID,
         WUE_adj_av,
         Sp_Code,
         Plot,
         Block,
         Shelter,
         Plot_number,
         Population,
         Treatment,
         Scientific_Name,
         Native_Status,
         Growth_Habit,
         Growth_2,
         Perenniality,
         Rating,
         Rating_2,
         Source)

summary(WUE)
hist(WUE$WUE_adj_av)

ggplot(WUE, aes(x=Sp_Code, y=WUE_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

# Bizarre outliers for DANCAL, AIRCAR, STIPUL

### Look up in plant physiology book

# What are normal ranges for short and long term WUE?
# Ask Jen
```

## Leaf Nitrogen
```{r}
cn <- read_csv(here::here("data/CN_2023.csv"))

summary(cn)
unique(cn$Error)

cn2 <- cn %>% 
  filter(is.na(Error) | Error == "new")

# break out species from sample ID
cn2$Sp_Code <- substr(cn2$Sample_ID,4,9)

ct_cn <- cn2 %>% count(Sp_Code)

cn2$Plant_ID <- substr(cn2$Sample_ID, start=1,stop=11)
cn2$Plant_ID <- paste0(
  substr(cn2$Plant_ID, 1, 2),
  ".",
  substr(cn2$Plant_ID, 4, 9),
  ".",
  substr(cn2$Plant_ID, 11, 11)
)

cn2$Plot <- substr(cn2$Plant_ID, 1, 2)

cn2$N_per <- (cn2$Total_N_µg/(cn2$Sample_Weight_mg_from_Sample_List*1000))*100

cn2$CN_ratio <- cn2$Total_C_µg/cn2$Total_N_µg

# calculate intrinsic WUE

# discrimination during diffusion, assume 4.4
a <- 4.4

# discrimination by rubisco, assume 27
b <- 27

# reference C02 carbon 13, assume -8 
delta_air <- -8

# ambinet CO2
C_a <- 420  # µmol/mol

cn2 <- cn2 %>%
  mutate(
    # Step 1: carbon isotope discrimination
    Delta = (delta_air - δ13CVPDB_permil) / (1 + δ13CVPDB_permil / 1000),
    
    # Step 2: estimate Ci/Ca
    Ci_Ca = (Delta - a) / (b - a),
    
    # Step 3: estimate intrinsic WUE (µmol CO2 / mol H2O)
    iWUE = C_a * (1 - Ci_Ca) / 1.6
  )


```

## Height
```{r}
#| message: false
#| warning: false
rgr <- read_csv(here::here("data/Relative Growth Rate 2023.csv"))
plots <- read_csv(here::here("data/Plot_info.csv"))
str(rgr)
summary(rgr)
head(rgr)
unique(rgr$Month)

# All IDs from May and June are missing ".1" at the end
# Extract species code from ID
# Extract plot, shelter letter from ID, add in block and population
rgr2 <- rgr %>% 
  mutate(Plant_ID = if_else(Month == "May" | Month == "June", true = paste0(Plant_ID, ".1"), false = Plant_ID),
         Sp_Code = substring(Plant_ID, 4,9),
         Plot = substring(Plant_ID, 1,2),) %>%
  filter(!Plant_ID == "E2.PHATAN.1B") %>% 
  left_join(plots,by = join_by(Plot))
str(rgr2)  

unique(rgr2$Sp_Code)
unique(rgr2$Shelter)
unique(rgr2$Plot)

head(rgr2)
rgr_max_h <- rgr2 %>% 
  select(
    Plant_ID,
    Month,
    Height_cm,
    Sp_Code,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment
  ) %>% 
  filter(!is.na(Height_cm))

rgr_max_h2 <- rgr_max_h %>% 
  group_by(Plant_ID) %>% 
  summarise(h_max = max(Height_cm, na.rm = TRUE),
            Sp_Code = first(Sp_Code),
            Plot = first(Plot),
            Population = first(Population),
            Treatment = first(Treatment)
            ) 
# rgr_max_h2 <- rgr_max_h2[-which(duplicated(rgr_max_h2)),]


rgr_max_h3 <- left_join(rgr_max_h2,plots, by = join_by(Plot))
rgr_max_h4 <- left_join(rgr_max_h3,spp, by = join_by(Sp_Code))

rgr_max_h4[which(rgr_max_h4$Sp_Code == "PHATAN"),]

ggplot(rgr_max_h4, aes(x=Sp_Code, y=h_max)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

ggplot(rgr_max_h4, aes(x=Rating, y=h_max)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

```

## Date of First Flower

```{r}
#| message: false
#| warning: false
pheno <- read_csv(here::here("data/Phenology_2023.csv"))
str(pheno)


# Check distribution of species
pheno_narm <- pheno %>% 
  filter(!is.na(Flower_Date))
ct_ph <- pheno_narm %>% count(Species, na.rm = TRUE)
# These should line up with reproductive allocation IDs

# Change date to date format
str(pheno_narm)
pheno_narm$Flower_Date <- as.Date(pheno_narm$Flower_Date, tryFormats = "%m/%d/%y")

# Change date into DOY
pheno_narm$doy <- day_of_year(pheno_narm$Flower_Date)

# Extract plot
pheno_narm$Plot <- substr(pheno_narm$Plant_ID,1,2)
pheno_narm$Sp_Code <- pheno_narm$Species

# merge plot info and species info
pheno2 <- left_join(pheno_narm, plots, by=join_by("Plot"))
pheno3 <- left_join(pheno2, spp, by=join_by("Sp_Code"))

# check for duplicates
pheno3[which(duplicated(pheno3$Plant_ID)),]


ggplot(pheno3, aes(x=Sp_Code, y=doy)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

ggplot(pheno3, aes(x=Rating, y=doy)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
```

## Seed Mass
```{r}
#| warning: false

seed <- read_csv(here::here("data/Seed inventory_weight.csv"))

str(seed)

seed <- seed %>%
  mutate(seed_mass = seed_mass_avg) %>% 
  select(
    Sp_Code,
    Population,
    seed_mass
  ) %>% 
  filter(!is.na(Sp_Code)
  ) %>% 
  filter(!is.na(seed_mass)) %>% 
  filter(!Sp_Code == "LUPSUC")

seed$seed_mass <- as.numeric(seed$seed_mass)

ggplot(seed, aes(x=Sp_Code, y=seed_mass)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )



```

## PCA
Datasets:
repro_fill_hi_n_narm
rgr_h
rgr_l
rgr_v
licor5
LMA
lf_area
cn2
rgr_max_h3
pheno3
seed

This code prepares each dataset for merging and then joins them one by one until you get the dataframe "full"

```{r}
## Each dataset should remove problem samples and perennials
## not sure how best to fill in gaps, after merging

# merge and scale
str(repro_fill_hi_n_narm)
repro <- repro_fill_hi_n_narm %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    Flower_Count,
    fl_average,
    Veg_Mass_ln,
    total_fl_mass_ln,
    RV
  ) %>% 
  mutate(Plant_ID = substr(Plant_ID,1,9)) %>% 
  filter(!is.na(Flower_Count)) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR"))

which(rgr_h$rate == "Inf")
str(rgr_h)
rgrh <- rgr_h %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    var_mo,
    rate
  ) %>% 
  filter(!is.na(rate)) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>% 
  group_by(Plant_ID) %>% 
  summarise(max_rate_ht = max(rate,na.rm= TRUE)) %>% 
  mutate(Plant_ID = substr(Plant_ID,1,9))

rgr_l[which(rgr_l$rate == "Inf"),"rate"] <- NA
rgrl <- rgr_l %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    var_mo,
    rate
  ) %>% 
  filter(!is.na(rate)) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>% 
  group_by(Plant_ID) %>% 
  summarise(max_rate_lf = max(rate,na.rm= TRUE)) %>% 
  mutate(Plant_ID = substr(Plant_ID,1,9))

rgr_v[which(rgr_v$rate == "Inf"),"rate"] <- NA
rgrv <- rgr_v %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    var_mo,
    rate
  ) %>% 
  filter(!is.na(rate)) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>% 
  group_by(Plant_ID) %>% 
  summarise(max_rate_vol = max(rate,na.rm= TRUE)) %>% 
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(licor5)
licor <- licor5 %>% 
  select(
    Plant_ID,
    Population,
    A_adj_av,
    WUE_adj_av,
    gsw_adj_av,
    Sp_Code
  ) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(LMA)
lma <- LMA %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    lma_g_m2
  ) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(lf_area)
lf_area <- lf_area %>% 
  select(
    Plant_ID,
    Sp_Code,
    total.leaf.area
  ) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(cn2)
cn <- cn2 %>% 
  select(
    Plant_ID,
    Sp_Code,
    N_per,
    CN_ratio,
    iWUE
  ) %>%
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(rgr_max_h4)
ht <- rgr_max_h4 %>% 
  select(
    Plant_ID,
    Sp_Code,
    h_max
  ) %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(pheno3)
pheno <- pheno3 %>% 
  select(
    Plant_ID,
    Sp_Code,
    doy
  ) %>%
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR")) %>%
  mutate(Plant_ID = substr(Plant_ID,1,9))

str(seed)
seed_a <- seed %>% 
  filter(!Sp_Code %in% c("ESCCAL","DANCAL","STIPUL","CORFIL","SANBIP","MIRMUL","CLAPUR","HOLVIR"))

str(repro)
str(rgrh)
full <- full_join(repro,rgrh,by=join_by(Plant_ID))

full2 <- full_join(full,rgrl,by=join_by(Plant_ID))

full3 <- full_join(full2,rgrv,by=join_by(Plant_ID))

str(full3)
str(licor)
full4 <- full_join(full3,licor,by=join_by(Plant_ID,Sp_Code,Population))

str(lma)
full5 <- full_join(full4,lma,by=join_by(Plant_ID,Sp_Code,Population))

str(lf_area)
full6 <- full_join(full5,lf_area,by=join_by(Plant_ID,Sp_Code))

str(cn)
full7 <- full_join(full6,cn,by=join_by(Plant_ID,Sp_Code))

full7[487,]
ht[which(ht$Plant_ID == "E2.PHATAN"),]

ht[556,]
full7[which(full7$Plant_ID == "I3.EROCIC"),]

str(ht)
full8 <- full_join(full7,ht,by=join_by(Plant_ID,Sp_Code))

full8[124,]
pheno[which(pheno$Plant_ID == "G3.AVEFAT"),]

pheno[230,]
full8[which(full8$Plant_ID == "E2.EROCIC"),]

str(pheno)
full9 <- full_join(full8,pheno,by=join_by(Plant_ID,Sp_Code))

# Actually need to join seed mass by Species and population

str(full9)
str(seed_a)
full <- full_join(full9,seed_a,by=join_by(Sp_Code,Population))

full[which(is.na(full$Population)),]
# 124 plats are NA for species and population

# Remove duplicates and error pls
full <- full %>% 
  filter(!duplicated(Plant_ID)) %>% 
  filter(!Plant_ID %in% c("D4.AVEFAT","D4.AVEBAR","H4.AVEBAR","H4.AVEFAT","D7.AVEFAT")) 

full <- full[!c(rowSums(is.na(full)) > 10),]

# Duplicates of E2.PHATAN in ht -- investigate, choose A. It's not E4.PHATAN.1-E2. Removed E2.PHATAN.1B
# Duplicates of I3.EROCIC in CN, chose A
# 3 duplicates of E2.EROCIC in full8 -- 2 E2.EROCIC in CN, plus a rando from RGR

# Fill in NA gaps
# Go get seed masses
# Go get phenology

rm(cn
  , cn2
  , ct_area
  , ct_cn
  , ct_mass
  , ct_ph
  , fit
  , full2
  , full3
  , full4
  , full5
  , full6
  , full7
  , full8
  , full9
  # , ht used later
  , intercepts
  , intercepts.df
  , join_check
  , lf_area
  , licor
  , licor5
  , licor5b
  , lma
  , LMA
  , LMA2
  , mass
  , mass_a
  , mxg
  , pheno
  , pheno_narm
  , pheno2
  , pheno3
  , pop
  , pred_habit
  , pred_line
  , results
  , rgr
  , rgr_h
  , rgr_hlv2
  , rgr_hlv2_long
  , rgr_l
  , rgr_max_h
  , rgr_max_h2
  , rgr_max_h3
  , rgr_max_h4
  , rgr_sum
  , rgr_v
  , rgr2
 # , rgrh used later
  , rgrl
  , rgrv
  , seed
  #, seed_a used later
  , sim_res
  , slopes
  , slopes_df
  , slopes.df
  , WUE
)


```

This code creates averages for each species x population and imputates them into missing rows
```{r}
## not sure how best to fill in gaps
# 301 total incomplete cases
# 200 Flower Count, Veg Mass etc
# 105 for max rate ht,lf,vol
# 215 for licor
# 203 for lma
# 204 for leaf area
# 256 for N
# 222 for pheno
colnames(full)
unique(full$Sp_Code)

incomplete_rows_2 <- full[!complete.cases(full),]

full_avgs <- full %>% 
  group_by(Sp_Code,Population) %>% 
  summarise(max_rate_ht = mean(max_rate_ht, na.rm = TRUE),
            max_rate_lf = mean(max_rate_lf, na.rm = TRUE),
            max_rate_vol = mean(max_rate_vol, na.rm = TRUE),
            A_adj_av = mean(A_adj_av, na.rm = TRUE),
            WUE_adj_av = mean(WUE_adj_av, na.rm = TRUE),
            gsw_adj_av = mean(gsw_adj_av, na.rm = TRUE),
            lma_g_m2 = mean(lma_g_m2, na.rm = TRUE),
            total.leaf.area = mean(total.leaf.area, na.rm = TRUE),
            N_per = mean(N_per, na.rm = TRUE),
            CN_ratio = mean(CN_ratio, na.rm = TRUE),
            iWUE = mean(iWUE, na.rm = TRUE),
            h_max = mean(h_max, na.rm = TRUE),
            doy = mean(doy, na.rm = TRUE),
            seed_mass = mean(seed_mass, na.rm = TRUE))

trait_cols <- c("max_rate_ht"
                ,"max_rate_lf"
                ,"max_rate_vol"
                ,"A_adj_av"
                ,"WUE_adj_av"
                ,"gsw_adj_av"
                ,"lma_g_m2"
                ,"total.leaf.area"
                ,"N_per"
                ,"CN_ratio"
                ,"iWUE"
                ,"h_max"
                ,"doy"
                ,"seed_mass")

full_filled <- full %>% 
  left_join(full_avgs, by = c("Sp_Code","Population"),suffix = c("","_avg")) %>% 
  mutate(
    across(all_of(trait_cols),
           ~ ifelse(is.na(.),
                    get(paste0(cur_column(),"_avg")),
                    .))
  ) %>% 
  select(-ends_with("_avg"))

colnames(full_filled) 
# It worked! Now just some NaNs
which(full_filled$N_per == "NaN" | is.na(full_filled$N_per))
# 25 with NaN
which(full_filled$seed_mass == "NaN" | is.na(full_filled$seed_mass))
# 72 with missing seed mass

# Need to fill in missing seed mass and look into missing leaf N
# Why don't I have any leaf N for AMSMEN?

```

### Correlation Table

### PCA and/or NMDS

Run raw PCA assuming each line of values are from the same individual (not true) and include all traits

```{r}
# There are NaN values in this column
# force them to be NAs
full_filled[which(full_filled$N_per == "NaN"), "N_per"] <- NA

full_filled$N_per

full_filled[which(full_filled$CN_ratio == "NaN"), "CN_ratio"] <- NA

full_filled$CN_ratio

full_filled[which(full_filled$iWUE == "NaN"), "iWUE"] <- NA

full_filled$iWUE

full_filled[which(full_filled$seed_mass == "NaN"), "seed_mass"] <- NA

full_filled$seed_mass

full_anon <- full_filled %>% 
  select(
    !c(Plant_ID,Sp_Code,Population)
  ) 
str(full_anon)

library(mice)

imp <- mice(full_anon,
            m = 1,
            method = "pmm",
            maxit = 20,
            seed = 123)

dat_imp <- complete(imp)

pca <- prcomp(dat_imp, scale. = TRUE)

# View summary
summary(pca)

# View loadings
pca$rotation

loadings_pc1 <- pca$rotation[, 1]
loadings_pc1
sort(abs(loadings_pc1), decreasing = TRUE)
contrib_pc1 <- loadings_pc1^2
sort(contrib_pc1, decreasing = TRUE)

loadings_pc2 <- pca$rotation[, 2]
loadings_pc2
contrib_pc2 <- loadings_pc2^2
sort(contrib_pc2, decreasing = TRUE)

loadings_pc3 <- pca$rotation[, 3]
loadings_pc3
contrib_pc3 <- loadings_pc3^2
sort(contrib_pc3, decreasing = TRUE)

biplot(pca)

pca_scores <- as.data.frame(pca$x)
pca_scores$Plant_ID <- full_filled$Plant_ID   
# Add grouping variable
pca_scores$Sp_Code <- substr(pca_scores$Plant_ID,4,9)
pca_scores_2 <- left_join(pca_scores,spp, by=join_by("Sp_Code"))

ggplot(pca_scores_2, aes(PC1, PC2, color = Native_Status)) +
  geom_point(size = 3, alpha = 0.8) +
  theme_minimal() +
  labs(title = "PCA of Iris Dataset",
       x = paste0("PC1 (", round(summary(pca)$importance[2,1] * 100, 1), "% variance)"),
       y = paste0("PC2 (", round(summary(pca)$importance[2,2] * 100, 1), "% variance)"))


# Install devtools if needed
library(ggbiplot)

ggbiplot(pca, groups = pca_scores_2$Rating_2, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

ggsave(here::here("figs/pca_rating2.jpg"))

ggbiplot(pca, groups = pca_scores_2$Growth_2, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

ggsave(here::here("figs/pca_growth2.jpg"))

ggbiplot(pca, groups = pca_scores_2$Sp_Code, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

ggsave(here::here("figs/pca_spp.jpg"))

# Make points smaller
# how to report raw loadings

# why doesn't AMSMEN have N?
# Fill in seeds 
# Fill in phenology

# Try PCA by species x population averages and add slope -- does it change
# Can we use that PCA to predict allometric scaling relationship
# Try PCA by species averages

# need to figure out if PCAs are used to model reproductive allocation and how to do that. How does a PCA interact with a dependent variable



```


Now try reducing to species x population and only using slope and intercept. 

```{r}
#Need slopes and intercepts from model then merge

head(full_filled)

full_collapse <- full_filled %>% 
  group_by(Sp_Code,Population) %>% 
  summarise(Flower_Count = mean(Flower_Count, na.rm = TRUE)
            , fl_average = mean(fl_average, na.rm = TRUE)
            , max_rate_ht = mean(max_rate_ht, na.rm = TRUE)
            , max_rate_lf = mean(max_rate_lf, na.rm = TRUE)
            , max_rate_vol = mean(max_rate_vol, na.rm = TRUE)
            , A_adj_av = mean(A_adj_av, na.rm = TRUE)
            , WUE_adj_av = mean(WUE_adj_av, na.rm = TRUE)
            , gsw_adj_av = mean(gsw_adj_av, na.rm = TRUE)
            , lma_g_m2 = mean(lma_g_m2, na.rm = TRUE)
            , total.leaf.area = mean(total.leaf.area, na.rm = TRUE)
            , N_per = mean(N_per, na.rm = TRUE)
            , CN_ratio = mean(CN_ratio, na.rm = TRUE)
            , iWUE = mean(iWUE, na.rm = TRUE)
            , h_max = mean(h_max, na.rm = TRUE)
            , doy = mean(doy, na.rm = TRUE)
            , seed_mass = mean(seed_mass, na.rm = TRUE)
    
  )

# convert NaN to NAs
full_collapse[which(full_collapse$N_per == "NaN"), "N_per"] <- NA

full_collapse$N_per

full_collapse[which(full_collapse$CN_ratio == "NaN"), "CN_ratio"] <- NA

full_collapse$CN_ratio

full_collapse[which(full_collapse$iWUE == "NaN"), "iWUE"] <- NA

full_collapse$iWUE

full_collapse[which(full_collapse$seed_mass == "NaN"), "seed_mass"] <- NA

full_collapse$seed_mass

full_collapse_anon <- full_collapse %>% 
  ungroup() %>% 
    select(
    !c(Sp_Code,
       Population)
  )

library(mice)

imp <- mice(full_collapse_anon,
            m = 1,
            method = "pmm",
            seed = 123)

dat_imp <- complete(imp)

pca <- prcomp(dat_imp, scale. = TRUE)

# View summary
summary(pca)

# View loadings
pca$rotation

loadings_pc1 <- pca$rotation[, 1]
loadings_pc1
sort(abs(loadings_pc1), decreasing = TRUE)
contrib_pc1 <- loadings_pc1^2
sort(contrib_pc1, decreasing = TRUE)

loadings_pc2 <- pca$rotation[, 2]
loadings_pc2
contrib_pc2 <- loadings_pc2^2
sort(contrib_pc2, decreasing = TRUE)

loadings_pc3 <- pca$rotation[, 3]
loadings_pc3
contrib_pc3 <- loadings_pc3^2
sort(contrib_pc3, decreasing = TRUE)

pca_scores <- as.data.frame(pca$x)
pca_scores <- pca_scores[, 1:3]  # example: PC1–PC3

pca_scores <- as.data.frame(pca$x)
pca_scores$Sp_Code <- full_collapse$Sp_Code   
# Add grouping variable
pca_scores$Population <- full_collapse$Population
pca_scores_2 <- left_join(pca_scores,spp, by=join_by("Sp_Code"))
```

Now try full_collapse, with reduced traits to see if it's better coverage. Make it anonymous. Try out different traits to get best coverage.

Then use axes in model with veg mass (independent) and fl mass (dependent)

```{r}
str(full_collapse)

# Limit it to a handful of traits:
# growth: RGR-height
# Size: Height
# Leaf physiology: A
# Phenology
# Seed mass
# photosynthetic rate (higher score than WUE)
# LMA
# This grouping explains 68% total variation on first three axes. 50%, 20%, 18%. Highest score even when swapping alternatives

full_collapse_anon_ltd <- full_collapse_anon %>% 
  select(
    max_rate_ht
    , h_max
    , N_per
    , doy
    , seed_mass
    , A_adj_av
    , lma_g_m2
  )

library(mice)

imp <- mice(full_collapse_anon_ltd,
            m = 1,
            method = "pmm",
            seed = 123)

dat_imp <- complete(imp)

pca <- prcomp(dat_imp, scale. = TRUE)

# View summary
summary(pca)

# View loadings
pca$rotation

loadings_pc1 <- pca$rotation[, 1]
loadings_pc1
sort(abs(loadings_pc1), decreasing = TRUE)
contrib_pc1 <- loadings_pc1^2
sort(contrib_pc1, decreasing = TRUE)

loadings_pc2 <- pca$rotation[, 2]
loadings_pc2
contrib_pc2 <- loadings_pc2^2
sort(contrib_pc2, decreasing = TRUE)

loadings_pc3 <- pca$rotation[, 3]
loadings_pc3
contrib_pc3 <- loadings_pc3^2
sort(contrib_pc3, decreasing = TRUE)

pca_scores <- as.data.frame(pca$x)
pca_scores <- pca_scores[, 1:3]  # example: PC1–PC3

pca_scores <- as.data.frame(pca$x)
pca_scores$Sp_Code <- full_collapse$Sp_Code   
# Add grouping variable
pca_scores$Population <- full_collapse$Population
pca_scores_2 <- left_join(pca_scores,spp, by=join_by("Sp_Code"))

# Install devtools if needed
library(ggbiplot)

p1 <- ggbiplot(pca, groups = pca_scores_2$Native_Status, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

#ggsave(here::here("figs/pca_native.jpg"))

p2 <- ggbiplot(pca, groups = pca_scores_2$Population, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

#ggsave(here::here("figs/pca_pop.jpg"))

p3 <- ggbiplot(pca, groups = pca_scores_2$Growth_Habit, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

#ggsave(here::here("figs/pca_growth.jpg"))

p4 <- ggbiplot(pca, groups = pca_scores_2$Rating, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

#ggsave(here::here("figs/pca_spp.jpg"))

p4 <- ggbiplot(pca, groups = pca_scores_2$Rating, ellipse = TRUE, circle = TRUE) +
  scale_color_discrete(name = "") +
  theme_minimal()

new_names <- c(
  "Height Growth Rate",
  "Maximum Height",
  "Leaf N",
  "Flower Date",
  "Seed Mass",
  "Photosynthetic Rate",
  "Leaf Mass Area"
)

rownames(pca$rotation) <- new_names

p <- ggbiplot(
  pca,
  obs.scale = 1,
  var.scale = 1,
  var.axes = TRUE,
  ellipse = FALSE
) +
  scale_x_continuous(expand = expansion(mult = 0.35)) +
  scale_y_continuous(expand = expansion(mult = 0.3))

# Make the points more transparent
p$layers[[2]]$aes_params$alpha <- 0.3

p + theme_classic()

ggsave(here::here("figs/base_pca.jpg"))

# Facet wrap four PCA graphs to show that it was a good idea to use grass v forbs

library(patchwork)

# assuming your plots are:
# p1, p2, p3, p4

combined_plot <- (p1 | p2) /
                 (p3 | p4)

combined_plot

ggsave(here::here("figs/combined_pca.jpg"))
```

```{r}
# use PCs to model fl_mass~veg mass
# Need to merge PCs with repro_fill_hi_narm

analysis_df <- repro_fill_hi_n_narm %>%
  left_join(pca_scores_2, by = c("Sp_Code"
                                 , "Population"
                                 , "Scientific_Name"
                                 , "Native_Status"
                                 , "Growth_Habit"
                                 , "Growth_2"
                                 , "Rating"
                                 , "Rating_2"
                                 , "Source"))

ht_plus <- ht %>% 
  mutate(Plant_ID = paste0(Plant_ID,".1"))
rgrh_plus <- rgrh %>% 
  mutate(Plant_ID = paste0(Plant_ID,".1"))

analysis_df_plus <- analysis_df %>% 
  left_join(seed_a, by = c("Sp_Code","Population")) %>% 
  left_join(rgrh_plus, by = c("Plant_ID")) %>% 
  left_join(ht_plus, by = c("Plant_ID", "Sp_Code"))

# so it is individual data, except for seed mass
# Return here to figure out models and testing
# Scale seed mass, height, and growth rate
# include growth form with raw variables
# include population

analysis_df_plus$seed_mass_sc <- scale(analysis_df_plus$seed_mass, center = TRUE, scale = TRUE)
analysis_df_plus$max_rate_ht_sc <- scale(analysis_df_plus$max_rate_ht, center = TRUE, scale = TRUE)
analysis_df_plus$h_max_sc <- scale(analysis_df_plus$h_max, center = TRUE, scale = TRUE)

m_1_2_3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC1 + PC2 + PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

mV <- lmer(total_fl_mass_ln ~ Veg_Mass_ln +
              (1 | Scientific_Name) +
              (1 | Plot),
           REML = FALSE,
              data = analysis_df_plus)

m_1 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC1 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

mx1 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC1 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

m_2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC2 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

mx2 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC2 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

m_3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

mx3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

ICtab(m_1_2_3
      , mV
      , m_1
      , mx1
      , m_2
      , mx2
      , m_3
      , mx3
)

anova(m_3, mV)
anova(m_3, mx3)
## LEFT OFF HERE ##

## Asked Chat GPT on June 1: "I'm trying to figure out which plant traits best predict or explain the reproductive allometric scaling relationship for 28 different species. The species list is comprised of native and invasive species, grasses and forbs, and come from 1-5 different populations from along a precipitation gradient. I ran a PCA with eight traits that explains 69% of the variation in traits. I then used the PCA axes in linear mixed effects models. The third PC alone axes consistently was part of the best fit model (using ICtab). I broke down the third PC axes into it's top three traits and tried those in different combinations in the model, along with growth habit (forbs v grasses) and population. The result is that I have five models with similar dAIC scores (within 2) with different predictors and I'm not sure how to proceed. The models include natural log of vegetative biomass, maximum height, relative height growth rate, population, and growth habit. How should I proceed?" 
# Followed up with "natural log of reproductive biomass is the response variable"

# Now compare PC3 with components, and then bring in ai and functional type

# are the componenets from the individual from averages? 

# TRying to remember why 1|Shelter vs 1|Plot
# Fit your population aridity model with (1|Plot)
m_aridity <- lmer(total_fl_mass_ln ~ Veg_Mass_ln * ai_sc + 
                    (1|Scientific_Name) + (1|Plot),
                  data = analysis_df_plus, REML = TRUE)

isSingular(m_aridity)   # TRUE means plot variance collapsed to ~0
VarCorr(m_aridity)      # look at the Plot variance estimate

# ln(R) ~ ln(V)
mV <- lmer(total_fl_mass_ln ~ Veg_Mass_ln +
              (1 | Scientific_Name) +
              (1 | Plot),
           REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + height
m_h <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + h_max_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + growth rate
m_r <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + max_rate_ht_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + seed mass
m_s <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + seed_mass_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + PC3
m_3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + height + ln(V):height
mxh <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*h_max_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + height + ln(V):rate
mxr <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*max_rate_ht_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + height + ln(V):seed
mxs <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*seed_mass_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + habit + ln(V):PC3
mx3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

ICtab(
  mV
  , m_h
  , m_r
  , m_s
  , m_3
  , mxh
  , mxr
  , mxs
  , mx3
)

# ln(R) ~ ln(V) + habit
m_g <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

m_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + ai_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
             REML = FALSE,
              data = analysis_df_plus)

# ln(R) ~ ln(V) + habit + ln(V):habit
mxg <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = FALSE,
              data = analysis_df_plus)

mxa <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*ai_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
             REML = FALSE,
              data = analysis_df_plus)

mxa_3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*ai_sc + PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

mx3_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC3 + ai_sc + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3 + ai_sc + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3xa <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*ai_sc + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

mxg_3 <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_Habit + PC3 +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

mx3_g <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC3 + Growth_Habit + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3_g <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*PC3 + Growth_Habit + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3xg <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*Growth_Habit + 
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3xg_a <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*Growth_Habit + ai_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

m_3xa_g <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*ai_sc + Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = FALSE,
              data = analysis_df_plus)

ICtab(
  mV
  , m_3
  , m_g
  , m_a
  , mx3
  , mxg
  , mxa
  , mxa_3
  , mx3_a
  , m_3_a
  , m_3xa
  , mxg_3
  , mx3_g
  , m_3_g
  , m_3xg
  , m_3xg_a
  , m_3xa_g
)

mxg_reml <- lmer(total_fl_mass_ln ~ Veg_Mass_ln*Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
            REML = TRUE,
              data = analysis_df_plus)
VarCorr(mxg_reml)

ggplot(analysis_df_plus, aes(x=PC3, fill=Growth_Habit)) + geom_density(alpha=0.5)

cor.test(analysis_df_plus$PC3, analysis_df_plus$Growth_Habit)

summary(lm(PC3 ~ Growth_Habit, data = analysis_df_plus))

m_3xg_a_REML <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*Growth_Habit + ai_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = TRUE,
              data = analysis_df_plus)
library(car)
vif(m_3xg_a_REML)

m_3xg_REML <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3*Growth_Habit +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = TRUE,
              data = analysis_df_plus)
library(car)
vif(m_3xg_REML)

library(lmerTest)

# Convert your existing model (no need to refit from scratch)
m_3xg_test <- as_lmerModLmerTest(m_3xg_REML)
summary(m_3xg_test)

m_3_a_REML <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + PC3 + ai_sc +
              (1 | Scientific_Name) +
              (1 | Plot),
              REML = TRUE,
              data = analysis_df_plus)
library(car)
vif(m_3_a_REML)

library(lmerTest)

# Convert your existing model (no need to refit from scratch)
m_3_a_test <- as_lmerModLmerTest(m_3_a_REML)
summary(m_3_a_test)

# Check assumptions
sim_res <- simulateResiduals(m_3xg_REML)
plot(m_3xg_REML)

testDispersion(sim_res)
# good on dispersion, p-value high
testOutliers(sim_res)
# good on outliers, p-value high
testUniformity(sim_res)
# Poor on normality but data is robust, as good as it gets

# Check for funnel shape (bad), we want constant variance
plot(fitted(m_3xg_REML), resid(m_3xg_REML))
abline(h = 0)
# good

# check for dramatic S-shape (bad)
qqnorm(resid(m_3xg_REML))
qqline(resid(m_3xg_REML))
# for the most part hugs the line but is a bit skewy

hist(resid(m_3xg_REML), breaks = 30)      # see the shape
plot(m_3xg_REML)                           # residuals vs fitted — look for patterns

lattice::qqmath(ranef(m_3xg_REML, condVar = TRUE))

# not sure what this does
plot(m_3xg_REML)

library(DHARMa)

sim_res <- simulateResiduals(fittedModel = m_3xg_REML, plot = TRUE)

# Individual diagnostic tests
testUniformity(sim_res)      # overall uniformity (like normality)
testDispersion(sim_res)      # over/underdispersion
testOutliers(sim_res)        # outlier detection

# Residuals vs each predictor (important!)
plotResiduals(sim_res, analysis_df_plus$Veg_Mass_ln)
plotResiduals(sim_res, analysis_df_plus$Growth_Habit)
plotResiduals(sim_res, analysis_df_plus$PC3)

leveneTest(resid(m_3xg_REML) ~ analysis_df_plus$Growth_Habit)

# Check if a quadratic term helps for continuous predictors
m_3xg_quad <- lmer(total_fl_mass_ln ~ Veg_Mass_ln + I(Veg_Mass_ln^2) + 
                     PC3*Growth_Habit +
                     (1|Scientific_Name) + (1|Plot),
                   REML = TRUE, data = analysis_df_plus)

# Compare residual plots
sim_res_quad <- simulateResiduals(m_3xg_quad)
plotResiduals(sim_res_quad, analysis_df_plus$Veg_Mass_ln)


analysis_df_plus |> 
  distinct(Scientific_Name, Population) |>  # or however population is coded
  nrow()

sp_pop_summary <- analysis_df_plus |>
  group_by(Scientific_Name, Population) |>
  summarise(PC3 = first(PC3), ai_sc = first(ai_sc), 
            Growth_Habit = first(Growth_Habit), .groups = "drop")

cor(sp_pop_summary$PC3, sp_pop_summary$ai_sc)

sp_pop_summary <- analysis_df_plus |>
  group_by(Scientific_Name, Population) |>  # adjust Population column name as needed
  summarise(PC3 = first(PC3), 
            Growth_Habit = first(Growth_Habit), 
            .groups = "drop")

# Point-biserial correlation (equivalent to Pearson with a binary variable)
cor(sp_pop_summary$PC3, as.numeric(as.factor(sp_pop_summary$Growth_Habit)))

# Or a t-test, which gives you more interpretable output
t.test(PC3 ~ Growth_Habit, data = sp_pop_summary)

# Visualize it
ggplot(sp_pop_summary, aes(x = Growth_Habit, y = PC3)) +
  geom_boxplot() +
  geom_jitter(width = 0.1)


library(ggplot2)

ggplot(analysis_df_plus, aes(spei_c, PC3)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_classic()

library(lme4)

m_PC3 <- lmer(
  total_fl_mass_ln ~ Veg_Mass_ln + PC3 + (1|Scientific_Name) + (1|Shelter),
  data = analysis_df_plus
)

m_arid <- lmer(
  total_fl_mass_ln ~ Veg_Mass_ln + spei_c + (1|Scientific_Name) + (1|Shelter),
  data = analysis_df_plus
)

m_both <- lmer(
  total_fl_mass_ln ~ Veg_Mass_ln + spei_c + PC3 + (1|Scientific_Name) + (1|Shelter),
  data = analysis_df_plus
)

install.packages("AICcmodavg")
library(AICcmodavg)

aictab(
  list(m_PC3, m_arid, m_both),
  modnames = c("PC3", "Aridity", "PC3+Aridity")
)
ICtab(m_PC3, m_arid, m_both)

summary(m_both)

library(performance)

check_collinearity(m_both)
vif(m_both)

ggplot(analysis_df_plus,
       aes(Veg_Mass_ln, total_fl_mass_ln, color = spei_c)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_viridis_c() +
  theme_classic()

```

Make a graph of predicted values one panel is PC3 lines and the other is SPEI.

```{r}
library(ggeffects)
library(ggplot2)
library(viridis)

pred_pc3 <- ggpredict(m_both, terms = c("Veg_Mass_ln", "PC3"))

library(dplyr)

pred_pc3 <- as.data.frame(pred_pc3) %>%
  mutate(panel = "PC3")

pred_pc3$PC3 <-as.numeric(pred_pc3$group) 
pred_pc3 <- pred_pc3 %>% 
  mutate(PC3 =
    case_when(
      PC3 == 1 ~ -1.19,
      PC3 == 2 ~ -0.08,
      PC3 == 3 ~ 1.03
    )
  )

p_main <- ggplot() +

  # Raw data points
  geom_point(
    data = analysis_df_plus,
    aes(x = Veg_Mass_ln,
        y = total_fl_mass_ln),
    alpha = 0.3
  ) +

  # Predicted lines
  geom_line(
    data = pred_pc3,
    aes(x = x,
        y = predicted,
        color = PC3,
        group = PC3),
    linewidth = 1.2
  ) +

  # Shared continuous scale
  scale_color_viridis_c(
    direction = -1,
    name = "PC3",
    option = "plasma"
  ) +
  scale_fill_viridis_c(
    direction = -1,
    name = "PC3",
    option = "plasma",
    guide = "none"   # prevents duplicate legend
  ) +
   annotate(
    "text",
    x = Inf,
    y = -Inf,
    label = "Short",
    hjust = 1,
    vjust = -0.5,
    size = 4
  ) +
   annotate(
    "text",
    x = Inf,
    y = -Inf,
    label = "Fast vertical growth",
    hjust = 1.0,
    vjust = -1.9,
    size = 4
  ) +
  annotate(
    "text",
    x = Inf,
    y = -Inf,
    label = "Small seeds",
    hjust = 1.0,
    vjust = -3.3,
    size = 4
  ) +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    label = "Large seeds",
    hjust = -0.2,
    vjust = 1.5,
    size = 4
  ) +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    label = "Slow vertical growth",
    hjust = -0.125,
    vjust = 2.9,
    size = 4
  ) +
  annotate(
    "text",
    x = -Inf,
    y = Inf,
    label = "Tall",
    hjust = -0.7,
    vjust = 4.3,
    size = 4
  ) +
  labs(
    x = "Ln(Vegetative Biomass)",
    y = "Predicted\nLn(Reproductive Biomass)",
    tag = "A"
  ) +

  theme_classic() +
  theme(legend.position = "bottom")

ggsave(here::here("figs/model_PC3.jpg"), p_main, width = 7, height = 4)
```

I wonder if I could make a graph with two small panels showing the effect of SPEI and PC3 to the right and to left is this big graph with PC lines

```{r}
library(ggeffects)

pc3_eff <- ggpredict(m_both, terms = "PC3")
spei_eff <- ggpredict(m_both, terms = "spei_c")

p_pc3 <- ggplot(pc3_eff,
                aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = conf.low,
        ymax = conf.high),
    alpha = 0.2
  ) +
  labs(
    x = "PC3",
    y = "Predicted\nreproductive biomass (ln)",
    tag = "B"
  ) +
  theme_classic()

p_spei <- ggplot(spei_eff,
                 aes(x, predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(
    aes(ymin = conf.low,
        ymax = conf.high),
    alpha = 0.2
  ) +
  labs(
    x = "SPEI",
    y = "Predicted\nreproductive biomass",
    tag = "C"
  ) +
  theme_classic()


#p_spei <- p_spei +
#  scale_x_continuous(
#    breaks = c(-0.4, 0, 0.4),
#    labels = scales::label_number()
 # )


p_pc3 <- p_pc3 +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p_spei <- p_spei +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


library(patchwork)

p_all <- (p_main | (p_pc3 / p_spei)) +
  plot_layout(
    widths = c(3, 1))
p_all

ggsave(here::here("figs/m_both_PC3_spei.jpg"), p_all, width = 7, height = 4)
```


```{r}
# SPEI predictions (3 lines: low, mean, high)
pred_pc3 <- ggpredict(m_both, terms = c("Veg_Mass_ln", "PC3"))

library(dplyr)

pred_pc3 <- as.data.frame(pred_pc3) %>%
  mutate(panel = "PC3")

pred_pc3$PC3 <-as.numeric(pred_pc3$group) 
pred_pc3 <- pred_pc3 %>% 
  mutate(PC3 =
    case_when(
      PC3 == 1 ~ -1.19,
      PC3 == 2 ~ -0.08,
      PC3 == 3 ~ 1.03
    )
  )

pg <- ggplot() +

  # Raw data points
  geom_point(
    data = analysis_df_plus,
    aes(x = Veg_Mass_ln,
        y = total_fl_mass_ln),
    alpha = 0.3
  ) +

  # Predicted lines
  geom_line(
    data = pred_pc3,
    aes(x = x,
        y = predicted,
        color = PC3,
        group = PC3),
    linewidth = 1.2
  ) +

  # Shared continuous scale
  scale_color_viridis_c(
    direction = -1,
    name = "PC3",
    option = "plasma"
  ) +
  scale_fill_viridis_c(
    direction = -1,
    name = "PC3",
    option = "plasma",
    guide = "none"   # prevents duplicate legend
  ) +

  labs(
    x = "Ln(Vegetative Biomass)",
    y = "Predicted ln(Reproductive Biomass)"
  ) +

  theme_classic() +
  theme(legend.position = "bottom")

pg
```
Big error when trying to re-run code and PCA. Need to go back through with fine-toothed comb. Left off on trying to get PCA graph labels to spread out for readability. Grrrrr




