---
title: "Wrangling & Analysis"
format: html
execute:
  cache: false       # disables caching
  eval: false        # optional: only enable when ready to run
---

## Background

These data come from a common garden drought experiment in Davis, California investigating the functional trait strategies of California grassland plants from five different grassland communities spanning California's latitudinal precipitation gradient. There 10-12 species per community or provenance, each provenance has 16 plots (8 drought, 8 control) across 10 shelters (5 drought, 5 control). In 2022-2023, we focused on collecting whole plant traits, including reproductive investment. This study focuses on how these traits relate to reproductive investment depending on native status, growth habit or provenance. This document will clean and wrangle the many datasets and images associated with this study.

## Traits & Objectives

-   Reproductive allocation
-   End of Season biomass
-   Relative Growth Rate
-   Photosynthetic Rate
-   Leaf Mass Area
-   Leaf Area
-   WUE Height
-   Leaf Nitrogen
-   Date of first flower
-   Mid season biomass partial
-   Root biomass partial
-   Root:shoot ratio partial
-   Root diameter partial
-   Specific root length partial
-   Root nitrogen partial

## To-do list
- Check for effects of drought treatment
- Check assumptions for models
- Fix Map orientation
- Licor outliers and mismatches
- Include seed mass in PCA
- Figure out if PCA is predicting something
- check for outliers in not RGR and non reproductive stuff

## Setup

```{r}
#| message: false
#| warning: false
#| echo: false

library(colorspace)
library(readr)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(lme4)
library(LeafArea)
#library(mapdata)
#library(maps)
#library(terra)
library(broom)
library(tidyverse)
#library(sf)
#library(geodata)
library(tigris)
library(ggnewscale)
#library(scatterpie)
library(lmtest)
library(car)
library(bbmle)
library(cropgrowdays)
library(emmeans)
library(here)
#install.packages("DHARMa")
library(DHARMa)
```

## CA Map and Plant Comp

Make a map of California with the inside of the boundary as a heat map of an aridity index. Then add five pie charts of plant composition.

Precip data from WorlClim

```{r}
#| eval: false
# ----- climate data -----
ppt <- worldclim_global(var = "prec", res = 2.5, path = "data/")
ppt_annual <- sum(ppt)

# ----- California boundary -----
ca <- states(cb = TRUE) |> subset(NAME == "California")
ca_vect <- vect(ca)

# ----- crop & mask -----
ppt_ca <- crop(ppt_annual, ca_vect)
ppt_ca <- mask(ppt_ca, ca_vect)

# ----- convert to df -----
ppt_df <- as.data.frame(ppt_ca, xy = TRUE)
names(ppt_df) <- c("x", "y", "ppt")

# ----- plot -----
ggplot(ppt_df) +
  geom_raster(aes(x = x, y = y, fill = ppt)) +
  geom_sf(data = ca, fill = NA, color = "black") +
  scale_fill_viridis_c(name = "Annual precip (mm)") +
  coord_sf() +
  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal()

rm(ppt_ca, ppt, ppt_annual)

# Put it all together now

## Need the pie chart data to have the following columns: ID, Lat, Long, Sp1 proportion, Sp2 proportion, etc.

# so we need to pivot wider the dataset and average across ten plots and add coordaintes for each site then use mapPies function
plant_comp <- read_csv(here::here("data/USDA_Site Composition.csv"))
str(plant_comp)

# First average across ten plots
comp_sum <- plant_comp %>%
  filter(!Species == "Grass") %>% 
  group_by(Site, Species) %>% 
  summarise(avg_cov = mean(Percent_Cover, na.rm = TRUE))

# get rid of bare ground, rock, and litter
comp_sum <- comp_sum %>%
  filter(!Species %in% c("Litter","Rock","Bare ground")) %>% 
  group_by(Site) %>% 
  arrange(desc(avg_cov), .by_group = TRUE)

PlList_2023 <- read_csv(here::here("data/USDA Drought Project 2023 Plant List.csv"))
pl_list <- PlList_2023 %>% 
  select(Site, Species)

comp_reduced <- left_join(pl_list, comp_sum, by=join_by(Site, Species))

comp_reduced$prop_cov <- comp_reduced$avg_cov/100

comp_reduced <- comp_reduced %>% 
  group_by(Site) %>% 
  mutate(prop_total = sum(prop_cov))

comp_reduced$prop_adj <- comp_reduced$prop_cov/comp_reduced$prop_total

comp_wide <- comp_reduced %>% 
  select(Site,
         Species,
         prop_adj) %>% 
  pivot_wider(names_from = Species,
              values_from = prop_adj)

## add lat and long columns
comp_wide$lat <- c(39.75141,37.37914,38.87508,34.70243,34.91771)
comp_wide$long <- c(-123.6302,-121.72841,-122.42114,-120.0545,-119.16882)
comp_wide$long_move <- comp_wide$long - 2
comp_wide$lat_move<- comp_wide$lat
comp_wide$lat_move[4] <- comp_wide$lat[4] + 1
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1
comp_wide$lat_move[1] <- comp_wide$lat[1] + 1
comp_wide$lat_move[2] <- comp_wide$lat[2] - 0.5
comp_wide$long_move[3] <- comp_wide$long[3] - 2.5
comp_wide$lat_move[3] <- comp_wide$lat[3] - 0.25
comp_wide$long_move[4] <- comp_wide$long[4] - 3.5
comp_wide$lat_move[4] <- comp_wide$lat[4] + 0.1
comp_wide$long_move[5] <- comp_wide$long[5] - 2.5
comp_wide$lat_move[5] <- comp_wide$lat[5] - 1.5
#comp_wide$long_move[1] <- comp_wide$long[1] + 0.2
comp_wide

comp_wide[is.na(comp_wide)] <- 0
colors <- rainbow_hcl(37)


# ----- plot -----
ggplot() +
  geom_raster(data = ppt_df, aes(x = x, y = y, fill = ppt)) +
    scale_fill_viridis_c(name = "Annual precip (mm)") +
  new_scale_fill() +
 geom_segment(
  data = comp_wide,
  aes(x = long, y = lat, xend = long_move, yend = lat_move),
  color = "black", linewidth = 0.75
) +
   geom_scatterpie(
    data = comp_wide,
    aes(long_move, lat_move, r = 1),
    cols = c("Acmispon parviflorus"
             ,"Aira caryophyllea"
             ,"Bromus diandrus"
             ,"Bromus hordeaceus"
             ,"Clarkia purpurea"
             ,"Danthonia californica"
             ,"Daucus pusillus"
             ,"Eschscholzia californica"
             ,"Galium parisiense"
             ,"Hypochaeris glabra"
             ,"Madia gracilis"
             ,"Sanicula bipinnatifida"
             ,"Avena barbata"
             ,"Erodium botrys"
             ,"Geranium dissectum"
             ,"Lupinus bicolor"
             ,"Torilis arvensis"
             ,"Trifolium hirtum"
             ,"Vicia sativa"
             ,"Bromus madritensis"
             ,"Elymus caput-medusae"
             ,"Holocarpha virgata"
             ,"Micropus californicus"
             ,"Plantago erecta"
             ,"Avena fatua"
             ,"Calandrinia menziesii"
             ,"Centaurea melitensis"
             ,"Corethrogyne filaginifolia"
             ,"Erodium cicutarium"
             ,"Stipa pulchra"
             ,"Amsinckia menziesii"
             ,"Bromus arenarius"
             ,"Caulanthus coulteri"
             ,"Lupinus microcarpus"
             ,"Lupinus succulentus"
             ,"Phacelia tanacetifolia"
             ,"Mirabilis multiflora")   # your pie-slice variables
  ) +
scale_fill_manual(
    name = "Species",
    values = c(
             "Acmispon parviflorus" = colors[1]
             ,"Aira caryophyllea" = colors[2]
             ,"Bromus diandrus" = colors[3]
             ,"Bromus hordeaceus" = colors[4]
             ,"Clarkia purpurea" = colors[5]
             ,"Danthonia californica" = colors[6]
             ,"Daucus pusillus" = colors[7]
             ,"Eschscholzia californica" = colors[8]
             ,"Galium parisiense" = colors[9]
             ,"Hypochaeris glabra" = colors[10]
             ,"Madia gracilis" = colors[11]
             ,"Sanicula bipinnatifida" = colors[12]
             ,"Avena barbata" = colors[13]
             ,"Erodium botrys" = colors[14]
             ,"Geranium dissectum" = colors[15]
             ,"Lupinus bicolor" = colors[16]
             ,"Torilis arvensis" = colors[17]
             ,"Trifolium hirtum" = colors[18]
             ,"Vicia sativa" = colors[19]
             ,"Bromus madritensis" = colors[20]
             ,"Elymus caput-medusae" = colors[21]
             ,"Holocarpha virgata" = colors[22]
             ,"Micropus californicus" = colors[23]
             ,"Plantago erecta" = colors[24]
             ,"Avena fatua" = colors[25]
             ,"Calandrinia menziesii" = colors[26]
             ,"Centaurea melitensis" = colors[27]
             ,"Corethrogyne filaginifolia" = colors[28]
             ,"Erodium cicutarium" = colors[29]
             ,"Stipa pulchra" = colors[30]
             ,"Amsinckia menziesii" = colors[31]
             ,"Bromus arenarius" = colors[32]
             ,"Caulanthus coulteri" = colors[33]
             ,"Lupinus microcarpus" = colors[34]
             ,"Lupinus succulentus" = colors[35]
             ,"Phacelia tanacetifolia" = colors[36]
             ,"Mirabilis multiflora" = colors[37]
    )
    , guide = "none"
  ) +
  geom_sf(data = ca, fill = NA, color = "black",linewidth = 0.75) +
  coord_sf(xlim = c(-127, -113), ylim = c(32, 42)) +
  labs(title = "Annual Precipitation in California (WorldClim v2.1)") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = c(0.925,0.725)
  )  +
  geom_point(data = data.frame(lat = 38.544854, long = -121.784516), 
             aes(x = long, y = lat),
                 color = "red", 
                 size = 4)

ggsave(here::here("figs/CA_precip_comp_map.jpg"))

rm(ca,
   ca_vect,
   comp_reduced,
   comp_sum, 
   comp_wide,
   pl_list,
   plant_comp,
   PlList_2023,
   ppt_df)
```
## Reproductive allocation & End of Season biomass

Read in data, find outliers, fill in missing data

```{r}
#| eval: false
#| warning: false
## Read in Data
repro <- read_csv(here::here("data/Reproductive Allocation_2023.csv"))
str(repro)
summary(repro)
head(repro)

```

### Clean up and create variables

Remove rows that have NA in the Flower Count row and in the Veg Mass Row. 

```{r}
#| eval: false
# Remove individuals with no veg mass and no flower count
repro_narm <- repro %>% 
  drop_na(Flower_Count) %>% 
  drop_na(Veg_Mass)

# Read in plot data
plots <- read_csv(here::here("data/Plot_info.csv"))
str(plots)
summary(plots)
head(plots)

# Read in species data
spp <- read_csv(here::here("data/Species Codes.csv"))
unique(spp$Sp_Code)
unique(repro$Species)

# Pull plot code from Plant_ID, first two characters
repro_narm <- repro_narm %>% 
  mutate(Plot = substr(Plant_ID, start = 1, stop = 2))
unique(repro_narm$Plot)

# merge with plots
repro_full <- left_join(repro_narm,plots,by = join_by(Plot))

# Change name for consistency
repro_full$Sp_Code <- repro_full$Species

# merge with species data
repro_full <- left_join(repro_full,spp,by = join_by(Sp_Code))

# clean up dataset and re-order
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Perenniality,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Veg_Mass
  )


# repro_full[which(repro_full$Plant_ID == "G3.AVEFAT.1"),]
# two separate entries, go back and fix manually. Probably additive
# Additive! two flower bags and two veg bags.

# repro_full[which(repro_full$Plant_ID == "F2.BRODIA.1"),]
# was this counted twice?
# Counted twice, go with the heavier veg mass since debris was probably lost

# repro_full[which(repro_full$Plant_ID == "J1.BRODIA.1"),]
# was this counted twice?
# this was counted twice. The 4.01 mass is correct

rm(repro,repro_narm)
```

Now check for outliers
```{r}
#| eval: false
# Make a dataset so we can average flower mass per species x population x treatment and then find outliers

repro_outliers_vg <- repro_full %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(vg_avg = mean(Veg_Mass, na.rm = TRUE),
        vg_sd = sd(Veg_Mass, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Veg_Mass,
         vg_avg,
         vg_sd)

# remove any rows with NAs
repro_outliers_vg <- repro_outliers_vg[complete.cases(repro_outliers_vg),]

# get minimums and maximums using 3 standard deviations
veg_avgs <- repro_outliers_vg %>% 
  mutate(vg_max = vg_avg + 3*vg_sd,
         vg_min = vg_avg - 3*vg_sd)

# do the same for veg mass
vg_outliers <- veg_avgs[which(veg_avgs$Veg_Mass > veg_avgs$vg_max | veg_avgs$Veg_Mass < veg_avgs$vg_min),c("Plant_ID","Sp_Code","Population", "Veg_Mass","vg_avg","vg_sd","vg_min","vg_max")]
# Only 4 and they are just above the maximum, keep

# write_csv(vg_outliers, "vg_outliers.csv")

# now do separately for fls
repro_outliers_fl <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  select(Plant_ID,
         Sp_Code,
         Scientific_Name,
         Population,
         Flower,
         Flower_Mass_g,
         Veg_Mass) %>% 
  filter(!is.na(Flower_Mass_g))

# export dataset for intern/undergraduate practice
# write_csv(repro_outliers, "repro_intern.csv")

# Group by spp and population and then find averages and standard deviations
str(repro_outliers_fl)

repro_avgs_fl <- repro_outliers_fl %>% 
  group_by(Sp_Code,
           Population) %>% 
  mutate(fl_avg = mean(Flower_Mass_g, na.rm = TRUE),
            fl_sd = sd(Flower_Mass_g, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Population,
         Flower,
         Flower_Mass_g,
         fl_avg,
         fl_sd)

# remove any rows with NAs
repro_avgs_fl <- repro_avgs_fl[complete.cases(repro_avgs_fl),]

# get minimums and maximums using 3 standard deviations
repro_avgs_fl <- repro_avgs_fl %>% 
  mutate(fl_max = fl_avg + 3*fl_sd,
         fl_min = fl_avg - 3*fl_sd)

# Make a dataframe with rows of observations of flower masses that are either larger than the maximum or smaller than the minimum
fl_outliers <- repro_avgs_fl[which(repro_avgs_fl$Flower_Mass_g > repro_avgs_fl$fl_max | repro_avgs_fl$Flower_Mass_g < repro_avgs_fl$fl_min),c("Plant_ID","Sp_Code","Population", "Flower", "Flower_Mass_g","fl_avg","fl_sd","fl_min","fl_max")]

# make csvs to look deeper
# write_csv(fl_outliers, "fl_outliers.csv")

# find difference between flower mass and maximum
fl_outliers$dif <- fl_outliers$Flower_Mass_g-fl_outliers$fl_max

# how bad is it? anything bigger than 1 standard deviation difference is too big
fl_hi <- fl_outliers[which(fl_outliers$dif >= fl_outliers$fl_sd),]
# flower masses in fl_hi are way too big to include

# Make those flower masses NAs and replace in repro_full

# make repro_full long and then rbind and remove duplicates
fl_hi$Flower_Mass_g <- NA
fl_hi <- fl_hi %>% 
  select(
    Plant_ID,
    Sp_Code,
    Population,
    Flower,
    Flower_Mass_g) %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# make original dataset long
repro_full_long <- repro_full %>% 
  pivot_longer(cols = c(Flower_1_g, 
                        Flower_2_g, 
                        Flower_3_g,
                        Flower_4_g,
                        Flower_5_g,
                        Flower_6_g,
                        Flower_7_g,
                        Flower_8_g,
                        Flower_9_g,
                        Flower_10_g),
               names_to = "Flower",
               values_to = "Flower_Mass_g") %>% 
  mutate(
    ID_no = paste0(Plant_ID,Flower)
  )

# create list of Plant_ID and flower number combos that need to become NA
nays <- fl_hi$ID_no

# create that dataset of 
repro_long_filt <- repro_full_long %>% 
  filter(ID_no %in% c(nays)) %>% 
  mutate(
    Flower_Mass_g = NA
  )
# 27 outliers converted to NA

# check your work, should return an NA
repro_long_filt$Flower_Mass_g[repro_long_filt$ID_no == sample(nays,1)]

# Now rbind, order, and remove duplicates
# 5620 in repro fill long, and 27 in repro long filt
# should have 5647
repro_fullfilt_long <- rbind(repro_long_filt,repro_full_long)

repro_long <- repro_fullfilt_long[-which(duplicated(repro_fullfilt_long$ID_no)),]

repro_full_2 <- repro_long %>% 
  pivot_wider(id_cols = Plant_ID,
              names_from = Flower,
              values_from = Flower_Mass_g)

# bind with original
repro_full <- repro_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Recorder,
    Collection_Date,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Flower_Count,
    Veg_Mass
  )

repro_full_3 <- left_join(repro_full_2,repro_full,by=join_by(Plant_ID))

repro_full <- repro_full_3 %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Growth_2,
    Rating,
    Rating_2,
    Source,
    Plot,
    Block, 
    Shelter,
    Plot_number,
    Population,
    Treatment,
    Recorder,
    Collection_Date,
    Flower_1_g,
    Flower_2_g,
    Flower_3_g,
    Flower_4_g,
    Flower_5_g,
    Flower_6_g,
    Flower_7_g,
    Flower_8_g,
    Flower_9_g,
    Flower_10_g,
    Flower_Count,
    Veg_Mass
  )

rm(repro_outliers_fl,
   repro_outliers_vg, 
   repro_avgs_fl,
   veg_avgs, 
   fl_outliers, 
   vg_outliers, 
   fl_hi, 
   repro_full_long, 
   nays, 
   repro_long_filt, 
   repro_fullfilt_long, 
   repro_long,
   repro_full_2,
   repro_full_3)
```

Now flower outliers have been removed, veg outliers have been checked.

Now fill in average flower mass for for each Species x population x treatment and fill in for individuals with no flower masses. 

```{r}
#| eval: false
# Average flower masses for each individual
firstfl = which(colnames(repro_full)=="Flower_1_g")
lastfl = which(colnames(repro_full)=="Flower_10_g")
repro_full[,c(firstfl:lastfl)]

repro_full$fl_average <- rowMeans(repro_full[,c(firstfl:lastfl)], na.rm = TRUE)

# Find the individuals with no flower mass
no_fls <- repro_full[which(repro_full$fl_average == "NaN"),]

# include everything except fl-Average
firstcol = which(colnames(no_fls)=="Plant_ID")
lastcol = which(colnames(no_fls)=="Veg_Mass")
no_fls <- no_fls[,c(firstcol:lastcol)]

# remove the individuals with no flower masses
repro_full <- repro_full[which(repro_full$fl_average != "NaN"),]

# detach(package:plyr)
# when plyr is loaded after dplyr, group_by doesn't work
avg_fls <- repro_full %>% 
  group_by(Sp_Code, Population, Treatment) %>% 
  summarise(fl_average = mean(fl_average, na.rm = TRUE))

# mast avg fls onto no_fls and then re-merge no-fls with larger dataset
rep_fls <- left_join(no_fls,avg_fls,by=join_by(Sp_Code,Population,Treatment))
repro_fill <- rbind(rep_fls,repro_full)

# check your work
repro_fill[which(is.na(repro_fill$fl_average)),]
# There's one without averages to fill in? Calmen in E6...investigate

repro_fill$total_fl_mass <- repro_fill$Flower_Count*repro_fill$fl_average

# There's an entry for ESCCAL which hasn't been done yet -- look into
# Later you can remove any species with only one entry

rm(repro_full,
   no_fls,
   avg_fls)

```
Now remove species with low data, less than three observations

```{r}
#| eval: false
unique(repro_fill$Sp_Code)

# Who are the species with low n?
summary <- repro_fill %>% 
  group_by(Sp_Code) %>% 
  summarise(n = n()) %>% 
  filter(n < 3)
# ACMPAR, CAUCOU, ESCCAL, LUPSUC

Sp_List_small <- summary$Sp_Code

# Remove species with less than 3 entries
repro_fill_hi_n <- repro_fill[-(which(repro_fill$Sp_Code %in% c(Sp_List_small))),]
unique(repro_fill_hi_n$Sp_Code)

summary(repro_fill_hi_n)

repro_fill_hi_n_narm <- repro_fill_hi_n[-(which(repro_fill_hi_n$total_fl_mass %in% c(NA, -Inf))),]

repro_fill_hi_n_narm_1 <- repro_fill_hi_n_narm

# Lastly add in precipitation
precip <- read_csv(here::here("data/Site_Precip.csv"))
repro_fill_hi_n_narm <- left_join(repro_fill_hi_n_narm_1,precip, by=join_by("Population"))

rm(summary,
   Sp_List_small,
   repro_fill,
   repro_fill_hi_n,
   precip,
   repro_fill_hi_n_narm_1
   )
```

Find and add in SPEI

```{r}
#| eval: false
install.packages("devtools")
require(devtools)
devtools::install_github('seschaub/getSpei')
require(getSpei)

workdir <- getwd()
setwd(workdir)

# create data frame:
Population  <- c("ANGELO", "MCL", "BORR","SEDG","WWP")
longitude    <- c(-123.6302, -121.72841, -122.42114, -120.0545, -119.16882)
latitude     <- c(39.75141, 37.37914, 38.87508, 34.70243, 34.91771)
locations_df <- data.frame(Population, longitude, latitude)

# specify arguments and run function:
d1 <- spec_spei(spei_files = "spei48", start_y = 2015, end_y = 2015, 
                locations = locations_df)

write.csv(d1, "SPEI_Pop.csv")
```

```{r}
#| eval: false
spei <- read_csv(here::here("data/SPEI_Pop.csv"))
spei <- spei %>% 
  filter(month > 2 & month < 6)


spei_avg <- spei %>% 
  group_by(location_id) %>% 
  summarise(spei_avg = mean(spei48)) %>% 
  mutate(Population = location_id)

spei_avg <- spei_avg[,c("Population","spei_avg")]

repro_fill_hi_n_narm_1 <- left_join(repro_fill_hi_n_narm,spei_avg, by=join_by("Population"))

repro_fill_hi_n_narm <- repro_fill_hi_n_narm_1

repro_fill_hi_n_narm$Veg_Mass_ln <- log(repro_fill_hi_n_narm$Veg_Mass)
repro_fill_hi_n_narm$total_fl_mass_ln <- log(repro_fill_hi_n_narm$total_fl_mass)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm[-(which(repro_fill_hi_n_narm$total_fl_mass_ln %in% c(NA, -Inf))),]

# Create ratio of fl to v
repro_fill_hi_n_narm$RV <- repro_fill_hi_n_narm$total_fl_mass_ln/repro_fill_hi_n_narm$Veg_Mass_ln

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  mutate(rating.n = (
    case_when(
      Rating == "Native" ~ 1
      , Rating == "Non-native" ~ 2
      , Rating == "Limited" ~ 3
      , Rating == "Moderate" ~ 4
      , Rating == "High" ~ 5
    )
  ))

rm(
  rep_fls
  , repro_fill_hi_n_narm_1
  , spei
  , spei_avg
)

write.csv(repro_fill_hi_n_narm, here::here("data/repro_fill_hi_n_narm.csv"))

```

We did it! Now graph and analyze. Use "repro_fill_hi_n_narm" for reproductive allocation.

Read in cleaned dataset to reduce computing power
```{r}
repro_fill_hi_n_narm <- read_csv(here::here("data/repro_fill_hi_n_narm.csv"))
```

## Q1: 1:1

```{r}
anova(lm(total_fl_mass_ln ~ Veg_Mass_ln*Scientific_Name, data = repro_fill_hi_n_narm))

mod <- lm(total_fl_mass_ln ~ Veg_Mass_ln * Scientific_Name, 
          data = repro_fill_hi_n_narm)


slopes <- emtrends(mod, ~ Scientific_Name, var = "Veg_Mass_ln")
slopes

slopes.df <- summary(slopes) %>% 
  arrange(desc(Veg_Mass_ln.trend))

test_slopes_vs1 <- summary(slopes, infer = TRUE, null = 1)
test_slopes_vs1[which(test_slopes_vs1$p.value<0.05),]

level_order <- unique(as.factor(slopes.df$Scientific_Name))


# Graph a bar plot of each species in order of slope with standard error bars and line at 1

#("ANGELO" = "#440154FF", "MCL" = "#3B528BFF", "BORR" = "#21908CFF", "SEDG" = "#5DC863FF", "WWP" = "#FDE725FF")) + 

# Make species names italic and 45 degrees
# Maybe make bar colors if significant

sigs <- test_slopes_vs1[which(test_slopes_vs1$p.value<0.05),1]

slopes.df$columnColor <- "N.S."

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[1])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[2])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[3])] <- "p-value < 0.05"

slopes.df$columnColor[which(slopes.df$Scientific_Name == sigs[4])] <- "p-value < 0.05"

ggplot(slopes.df, aes(x = factor(Scientific_Name, level=c(level_order)), y = Veg_Mass_ln.trend
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.05" = "#440154FF")) +
  theme(
    axis.text.x = element_text(face = "italic", angle = 45, vjust = 1, hjust = 1),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp.jpg"))


ggplot(slopes.df, aes(x = factor(Scientific_Name, level=c(level_order)), y = Veg_Mass_ln.trend
                       , fill = columnColor
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, linewidth=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  scale_fill_manual(values = c("N.S." = "#21908CFF", "p-value < 0.05" = "#440154FF")) +
  theme(
    axis.text.x = element_blank(),
    legend.text =element_text(size=12, colour="black"),
    axis.text.y=element_text(size=12, colour="black"),
    axis.title=element_text(size=12,face="bold"),
    legend.position = c(0.75, 0.95),
    #legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),
    plot.margin = unit(c(0.25,0.25,0.25,0.75), "cm")
  ) + 
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "Species",
       y = "RV Slope",
       fill = "")

ggsave(here::here("figs/Q1_slopes_spp_prez.jpg"), width = 7,height = 3.5)

rm(
  mod
  , slopes
  , slopes.df
  , test_slopes_vs1
)

```

It's interesting that the three most widespread species are significant and low. And the arguable invasive native is the other significant.

## Q2: Growth Native Population

#### Model: RV
Model looking at what creates the ratio of log r and log v. Use species as random effect, and use that to determine best graph

```{r}
str(repro_fill_hi_n_narm)

repro_fill_hi_n_narm <- repro_fill_hi_n_narm %>% 
  mutate(rating.n = (
    case_when(
      Rating == "Native" ~ 1
      , Rating == "Non-native" ~ 2
      , Rating == "Limited" ~ 3
      , Rating == "Moderate" ~ 4
      , Rating == "High" ~ 5
    )
  ))

m0 <- glm(RV ~ 1, family = gaussian, data = repro_fill_hi_n_narm)

m1 <- glmer(RV ~ 1 + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m2 <- glmer(RV ~ 1 + (1|Block/Shelter/Plot), family = gaussian, data = repro_fill_hi_n_narm)

m3 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment), family = gaussian, data = repro_fill_hi_n_narm)

m4 <- glmer(RV ~ 1 + (1|Block/Shelter/Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m5 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m6 <- glmer(RV ~ 1 + (1|Block) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

ICtab(m0,m1,m2,m3,m4,m5,m6)

m5 <- glmer(RV ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

m_ratingxpop <- glmer(RV ~ Rating*Population + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm)

ICtab(m5
      ,m_ratingxpop)
lrtest(m_ratingxpop, m5)
Anova(m_ratingxpop)
summary(m_ratingxpop)

# try additive terms

repro_sub <- repro_fill_hi_n_narm %>% 
  select(
    RV,
    Native_Status,
    Rating,
    Rating_2,
    rating.n,
    Growth_Habit,
    Growth_2,
    Population,
    Precip_mm,
    precip_dif,
    spei_avg
  )

# Initialize an empty model
forward_model <- lm(RV ~ ., data = repro_sub)

# Forward stepwise regression
forward_model <- step(forward_model, direction = "forward", scope = formula(~ .))

# Initialize a model with all predictors
backward_model <- lm(RV ~ ., data = repro_sub)

# Backward stepwise regression
backward_model <- step(backward_model, direction = "backward")

# Initialize a model with all predictors
both_model <- lm(RV ~ ., data = repro_sub)

# Both-direction stepwise regression
both_model <- step(both_model, direction = "both")

# Check for overdispersion
# deviance/residual df


### Come back to this
### Kind of a mess

```

#### Model: Fl Mass

This is kind of a mess. The best models are speip, speip_growthp, growthp, and speip_growthx.
Growthp gets the closest to being "normal". repro_fill_hi_n_narm_i has three influential observations removed

```{r}
str(repro_fill_hi_n_narm)

repro_fill_hi_n_narm_i <- repro_fill_hi_n_narm

m0 <- glm(total_fl_mass_ln ~ 1, family = gaussian, data = repro_fill_hi_n_narm_i)

m1 <- glmer(total_fl_mass_ln ~ 1 + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m2 <- glmer(total_fl_mass_ln ~ 1 + (1|Plot), family = gaussian, data = repro_fill_hi_n_narm_i)

m3 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment), family = gaussian, data = repro_fill_hi_n_narm_i)

m4 <- glmer(total_fl_mass_ln ~ 1 + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m5 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m6 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m7 <- glmer(total_fl_mass_ln ~ 1 + (1|Block) + (1|Treatment) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m8 <- glmer(total_fl_mass_ln ~ 1 + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

ICtab(m0,m1,m2,m3,m4,m5,m6,m7,m8)

mV <- glmer(total_fl_mass_ln ~ Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_native <- glmer(total_fl_mass_ln ~ Native_Status*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating <- glmer(total_fl_mass_ln ~ Rating*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)
  
m_growth <- glmer(total_fl_mass_ln ~ Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating.n <- glmer(total_fl_mass_ln ~ rating.n*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_growth2 <- glmer(total_fl_mass_ln ~ Growth_2*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating2 <- glmer(total_fl_mass_ln ~ Growth_2*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_nativexgrowth <- glmer(total_fl_mass_ln ~ Native_Status*Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_ratingxgrowth <- glmer(total_fl_mass_ln ~ Rating*Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating.nxgrowth <- glmer(total_fl_mass_ln ~ rating.n*Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating2xgrowth <- glmer(total_fl_mass_ln ~ Rating_2*Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_rating2xgrowth2 <- glmer(total_fl_mass_ln ~ Rating_2*Growth_2*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_nativexgrowth2 <- glmer(total_fl_mass_ln ~ Native_Status*Growth_2*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_nativepgrowth <- glmer(total_fl_mass_ln ~ Native_Status + Growth_Habit + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_nativep <- glmer(total_fl_mass_ln ~ Native_Status + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_growthp <- glmer(total_fl_mass_ln ~ Growth_Habit + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name) + (1|Population), family = gaussian, data = repro_fill_hi_n_narm_i)

m_popp <- glmer(total_fl_mass_ln ~ Population + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_popx <- glmer(total_fl_mass_ln ~ Population*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_precipp <- glmer(total_fl_mass_ln ~ Precip_mm + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_precipx <- glmer(total_fl_mass_ln ~ Precip_mm*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_pdifp <- glmer(total_fl_mass_ln ~ precip_dif + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_pdifx <- glmer(total_fl_mass_ln ~ precip_dif*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_speip <- glmer(total_fl_mass_ln ~ spei_avg + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_speix <- glmer(total_fl_mass_ln ~ spei_avg*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_speip_growthx <- glmer(total_fl_mass_ln ~ spei_avg + Growth_Habit*Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

m_speip_growthp <- glmer(total_fl_mass_ln ~ spei_avg + Growth_Habit + Veg_Mass_ln + (1|Plot) + (1|Scientific_Name), family = gaussian, data = repro_fill_hi_n_narm_i)

ICtab(m8
      ,mV
      ,m_native
      ,m_rating
      ,m_growth
      ,m_rating.n
      ,m_growth2
      ,m_rating2
      ,m_nativexgrowth
      ,m_ratingxgrowth
      ,m_rating.nxgrowth
      ,m_rating2xgrowth
      ,m_rating2xgrowth2
      ,m_nativexgrowth2
      ,m_nativepgrowth
      ,m_nativep
      ,m_growthp
      ,m_popp
      ,m_popx
      ,m_precipp
      ,m_precipx
      ,m_pdifp
      ,m_pdifx
      ,m_speip
      ,m_speix
      ,m_speip_growthx
      ,m_speip_growthp)
anova(m_speip_growthp, mV)
anova(m_speip, mV)
anova(m_native, mV)
Anova(m_speip_growthx)
summary(m_speip_growthx)

#mall

# Check assumptions
# 1 Correct mean structure

# 2 Normally distributed residuals
res <- resid(m_speip_growthp)
qqnorm(res)
qqline(res)

library(DHARMa)

sim_res <- simulateResiduals(m_speip_growthp)
plot(sim_res)
testUniformity(sim_res)


plot(fitted(m_speip_growthx), resid(m_speip_growthx))
abline(h = 0, lty = 2)

# kind of a strong s shape...

# 3 Homescedascity
plot(fitted(m_speip_growthx), resid(m_speip_growthx))
abline(h = 0)

boxplot(resid(m_speip_growthx) ~ spei_avg,
        data = repro_fill_hi_n_narm)

boxplot(resid(m_speip_growthx) ~ Growth_Habit,
        data = repro_fill_hi_n_narm)
# No funnel shape but forbs have a larger spread

# 4 independence
boxplot(resid(m_speip_growthx) ~ Plot,
        data = repro_fill_hi_n_narm,
        las = 2,
        outline = FALSE)
abline(h = 0, lty = 2)
# no clear structure

# 5 Random effects are normally distributed
ranef_vals <- ranef(m_speip, condVar = TRUE)
qqnorm(ranef_vals$Scientific_Name[,1])
qqline(ranef_vals$Scientific_Name[,1])
# quite skewy

#qqnorm(ranef_vals$Block[,1])
#qqline(ranef_vals$Block[,1])

#qqnorm(ranef_vals$Shelter[,1])
#qqline(ranef_vals$Shelter[,1])
# a bit skewy?

qqnorm(ranef_vals$Plot[,1])
qqline(ranef_vals$Plot[,1])

#qqnorm(ranef_vals$Population[,1])
#qqline(ranef_vals$Population[,1])
# skewy

# 6 Random effects independent of fixed effects
## species aren't systematically related to veg mass ln (I think they are)
# May need random slopes?

# 7 No singularity
isSingular(m_speip_growthx)
# it's true because one or more variance components = 0
summary(m_speip_growthx)
# Plot:Shelter:Block is 0

# 8 no severe multicollinearity
car::vif(lm(total_fl_mass_ln ~ spei_avg + Growth_Habit + Veg_Mass_ln, data = repro_fill_hi_n_narm_i))
# all good because all less than 3

# 9 no influential points
library(influence.ME)
#influence.ME::influence(m_speip_growthp, obs = TRUE)
#infl <- influence(m_speip, obs = TRUE)
#cd <- cooks.distance(infl)
#cd_df <- data.frame(
#  row = seq_along(cd),
#  cooks_d = cd
#)
#p <- length(fixef(m_speip))   # number of fixed effects
#n <- nobs(m_speip)

#cd_thresh <- 4 / (n - p)

#high_infl <- cd_df[cd_df$cooks_d > cd_thresh, ]
#high_infl_2 <- cd_df[order(-cd_df$cooks_d), ][1:20, ]

#high_infl_m <- merge(
#  high_infl,
#  repro_fill_hi_n_narm,
 # by.x = "row",
#  by.y = "row.names"
#)
# This gives us 14 plants with high influence
# Try removing plants with high influence

#high_infl_m <- high_infl_m %>% 
#  filter(cooks_d > 0.099)

#repro_fill_hi_n_narm_i <- repro_fill_hi_n_narm %>% 
#  filter(!Plant_ID %in% high_infl_m$Plant_ID)

slopes <- emtrends(m_growth, ~ Growth_Habit, var = "Veg_Mass_ln")
slopes

slopes.df <- summary(slopes)

ggplot(slopes.df, aes(x = Growth_Habit, y = Veg_Mass_ln.trend
                      )) + 
  geom_bar(position = position_dodge(), stat = "identity", fill = "#21908CFF") +
  # geom_hline(yintercept = 1, linetype = "dashed") +
  geom_errorbar(aes(ymin=Veg_Mass_ln.trend-SE, ymax=Veg_Mass_ln.trend+SE), width=0.3, colour="#FDE725FF", alpha=0.9, size=1.3, position = position_dodge(0.9)) +
  theme_classic() +
  labs(x = "Growth Habit",
       y = "RV Slope",
       fill = "")

#ggsave(here::here("figs/slopes_growth_pred.jpg"))

```

Best to stop here and come back to it. Now work on bringing in functional traits and cleaning those up then woorking on PCA.

Maybe I could divide it up between non-native grasses and forbs, the lack of variation in the groups could be effecting things. I wonder if native status would hold any weight for forbs. and grasses maybe degree of invasiveness would play a bigger role?

## Q3: Traits

First read in all the traits we need and clean them up

-   Relative Growth Rate
-   Photosynthetic Rate
-   Leaf Mass Area
-   Leaf Area
-   WUE Height
-   Leaf Nitrogen
-   Date of first flower
-   Mid season biomass partial
-   Root biomass partial
-   Root:shoot ratio partial
-   Root diameter partial
-   Specific root length partial
-   Root nitrogen partial

## Relative Growth Rate

### Clean up dataset

```{r}
#| eval: false
rgr <- read_csv(here::here("data/Relative Growth Rate 2023.csv"))
plots <- read_csv(here::here("data/Plot_info.csv"))
str(rgr)
summary(rgr)
head(rgr)
unique(rgr$Month)

# All IDs from May and June are missing ".1" at the end
# Extract species code from ID
# Extract plot, shelter letter from ID, add in block and population
rgr2 <- rgr %>% 
  mutate(Plant_ID = if_else(Month == "May" | Month == "June", true = paste0(Plant_ID, ".1"), false = Plant_ID),
         Sp_Code = substring(Plant_ID, 4,9),
         Plot = substring(Plant_ID, 1,2),) %>% 
  left_join(plots,by = join_by(Plot))
str(rgr2)  

unique(rgr2$Sp_Code)
unique(rgr2$Shelter)
unique(rgr2$Plot)

# Find gaps, investigate
# Fixed, entering errors besides A1.ACMPAR in May which is missing leaf count

incomplete_rows <- rgr2[,c("Plant_ID","Month","Height_cm","Leaf_number","canopy_x","canopy_y")]
incomplete_rows_2 <- incomplete_rows[!complete.cases(incomplete_rows),]
incomplete_rows_2 <- incomplete_rows_2[rowSums(is.na(incomplete_rows_2)) < 4,] 

# Find and investigate outliers
# Good job for interns
# Find average and standard deviation for each species, each growth variable (make area first), per population per month
str(rgr2)
rgr2$area <- as.numeric(rgr2$canopy_x)*as.numeric(rgr2$canopy_y)

rgr_avgs <- rgr2 %>% 
  group_by(Sp_Code,
           Month,
           Population) %>% 
  mutate(ht_avg = mean(Height_cm, na.rm = TRUE),
            ht_sd = sd(Height_cm, na.rm = TRUE),
            lf_avg = mean(Leaf_number, na.rm = TRUE),
            lf_sd = sd(Leaf_number, na.rm = TRUE),
            area_avg = mean(area, na.rm = TRUE),
            area_sd = sd(area, na.rm = TRUE),
            x_avg = mean(canopy_x, na.rm = TRUE),
            x_sd = sd(canopy_x, na.rm = TRUE),
            y_avg = mean(canopy_y, na.rm = TRUE),
            y_sd = sd(canopy_y, na.rm = TRUE)) %>% 
  select(Plant_ID,
         Sp_Code,
         Month,
         Population,
         Height_cm,
         ht_avg,
         ht_sd,
         Leaf_number,
         lf_avg,
         lf_sd,
         area,
         area_avg,
         area_sd,
         canopy_x,
         canopy_y,
         x_avg,
         x_sd,
         y_avg,
         y_sd)

rgr_avgs <- rgr_avgs[complete.cases(rgr_avgs),]

rgr_avgs <- rgr_avgs %>% 
  mutate(ht_max = ht_avg + 3*ht_sd,
         ht_min = ht_avg - 3*ht_sd,
         lf_max = lf_avg + 3*lf_sd,
         lf_min = lf_avg - 3*lf_sd,
         area_max = area_avg + 3*area_sd,
         area_min = area_avg - 3*area_sd,
         x_max = x_avg + 3*x_sd,
         x_min = x_avg - 3*x_sd,
         y_max = y_avg + 3*y_sd,
         y_min = y_avg - 3*y_sd)

ht_outliers <- rgr_avgs[which(rgr_avgs$Height_cm > rgr_avgs$ht_max | rgr_avgs$Height_cm < rgr_avgs$ht_min),c("Plant_ID","Sp_Code","Month","Population", "Height_cm","ht_avg","ht_sd","ht_min","ht_max")]

lf_outliers <- rgr_avgs[which(rgr_avgs$Leaf_number > rgr_avgs$lf_max | rgr_avgs$Leaf_number < rgr_avgs$lf_min),c("Plant_ID","Sp_Code","Month","Population", "Leaf_number","lf_avg","lf_sd","lf_min","lf_max")]

area_outliers <- rgr_avgs[which(rgr_avgs$area > rgr_avgs$area_max | rgr_avgs$area < rgr_avgs$area_min),c("Plant_ID","Sp_Code","Month","Population", "area","area_avg","area_sd","area_min","area_max")]

x_outliers <- rgr_avgs[which(rgr_avgs$canopy_x > rgr_avgs$x_max | rgr_avgs$canopy_x < rgr_avgs$x_min),c("Plant_ID","Sp_Code","Month","Population", "canopy_x","x_avg","x_sd","x_min","x_max")]

y_outliers <- rgr_avgs[which(rgr_avgs$canopy_y > rgr_avgs$y_max | rgr_avgs$canopy_y < rgr_avgs$y_min),c("Plant_ID","Sp_Code","Month","Population", "canopy_y","y_avg","y_sd","y_min","y_max")]

## Now go and investigate outliers using original data
```

### Create variables

```{r}
#| eval: false

# Calculate canopy volume (height x canopy area)
head(rgr2)
rgr2$volume <- rgr2$Height_cm*rgr2$canopy_x*rgr2$canopy_y
rgr2$Date <- as.Date(rgr2$Date, tryFormats = "%m/%d/%y")
str(rgr2)

# Separate by metric

rgr_h_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         Height_cm) %>% 
  filter(!is.na(Height_cm)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, Height_cm)) %>% 
  mutate(rgr_h_FM = (log(Height_cm_March)-log(Height_cm_February))/as.numeric((Date_March-Date_February)),
         rgr_h_MA = (log(Height_cm_April)-log(Height_cm_March))/as.numeric((Date_April-Date_March)),
         rgr_h_AM = (log(Height_cm_May)-log(Height_cm_April))/as.numeric((Date_May-Date_April)),
         rgr_h_MJ = (log(Height_cm_June)-log(Height_cm_May))/as.numeric((Date_June-Date_May)))

rgr_l_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         Leaf_number) %>% 
  filter(!is.na(Leaf_number)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, Leaf_number)) %>% 
  mutate(rgr_l_FM = (log(Leaf_number_March)-log(Leaf_number_February))/as.numeric((Date_March-Date_February)),
         rgr_l_MA = (log(Leaf_number_April)-log(Leaf_number_March))/as.numeric((Date_April-Date_March)),
         rgr_l_AM = (log(Leaf_number_May)-log(Leaf_number_April))/as.numeric((Date_May-Date_April)),
         rgr_l_MJ = (log(Leaf_number_June)-log(Leaf_number_May))/as.numeric((Date_June-Date_May)))

rgr_v_wide <- rgr2 %>% 
  select(Plant_ID,
         Date,
         Month,
         volume) %>% 
  filter(!is.na(volume)) %>%
  pivot_wider(id_cols = Plant_ID,
              names_from = Month,
              values_from = c(Date, volume)) %>% 
  mutate(rgr_v_FM = (log(volume_March)-log(volume_February))/as.numeric((Date_March-Date_February)),
         rgr_v_MA = (log(volume_April)-log(volume_March))/as.numeric((Date_April-Date_March)),
         rgr_v_AM = (log(volume_May)-log(volume_April))/as.numeric((Date_May-Date_April)),
         rgr_v_MJ = (log(volume_June)-log(volume_May))/as.numeric((Date_June-Date_May)))
  
# Join all three together but only keep rgrs, full join by Plant_ID

rgr_h_l <- full_join(rgr_h_wide,rgr_l_wide,by = join_by(Plant_ID))

rgr_hlv <- full_join(rgr_h_l,rgr_v_wide,by = join_by(Plant_ID))

rgr_hlv2 <- rgr_hlv %>% 
  select(
    Plant_ID,
    rgr_h_FM,
    rgr_h_MA,
    rgr_h_AM,
    rgr_h_MJ,
    rgr_l_FM,
    rgr_l_MA,
    rgr_l_AM,
    rgr_l_MJ,
    rgr_v_FM,
    rgr_v_MA,
    rgr_v_AM,
    rgr_v_MJ
  )

## Wooooo we did it!
## Some values have -Inf, investigate that at some point. For now you're done!

write_csv(rgr_hlv2,here::here("data/RGR_clean.csv"))
```

### Exploratory Graphs
We have height, leaf number, area, and volume

```{r}
rgr_hlv2 <- read_csv(here::here("data/RGR_clean.csv"))

rgr_hlv2_long <- rgr_hlv2 %>% 
  pivot_longer(cols = c("rgr_h_FM":"rgr_v_MJ"),
               names_to = "var_mo",
               values_to = "rate") %>%
  filter(!is.na(rate))

rgr_sum <- rgr_hlv2_long %>%
  filter(!rate %in% c("-Inf","NaN","Inf")) %>% 
  group_by(var_mo) %>% 
  summarise(avg = mean(rate),
            stdev = sd(rate)) %>% 
  arrange(desc(avg))
```

Highest growth rate for volume is Feb-Mar (0.0957), then Mar-Apr (0.0827). For height, highest growth rate is Mar-Apr (0.0405), then Feb-Mar (0.0367). For leaves, highest growth rate is 0.0291 in Feb-Mar, then May-Jun 0.0198 (seems wrong). 

Remember we didn't know who each grass was in February. So might need to throw those out OR separate by growth habit.

### Select Reference times
One predictive variable of allometric slope might be relative growth rate timing and/or maximum relative growth rate. 

### Explore RGR

```{r}
#| eval: false

# parse out plant ID for species, plot, Population
plots <- read_csv(here::here("data/Plot_info.csv"))
str(plots)
summary(plots)
head(plots)

spp <- read_csv(here::here("data/Species Codes.csv"))
unique(spp$Sp_Code)

# need to pull out first two characters in string of plant ID to create a "plot" column

rgr_hlv <- rgr_hlv2 %>% 
  mutate(Plot = substr(Plant_ID, start = 1, stop = 2),
         Species = substr(Plant_ID, start = 4, stop = 9))
unique(rgr_hlv$Plot)
unique(rgr_hlv$Species)

rgr_hlv_full <- left_join(rgr_hlv,plots,by = join_by(Plot))

rgr_hlv_full$Sp_Code <- rgr_hlv$Species

rgr_hlv_full <- left_join(rgr_hlv_full,spp,by = join_by(Sp_Code))

rgr_h <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_h_FM,
    rgr_h_MA,
    rgr_h_AM,
    rgr_h_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_h_FM":"rgr_h_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_h,here::here("data/RGR_height.csv"))

rgr_l <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_l_FM,
    rgr_l_MA,
    rgr_l_AM,
    rgr_l_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_l_FM":"rgr_l_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_l,here::here("data/RGR_leaves.csv"))

rgr_v <- rgr_hlv_full %>% 
  select(
    Plant_ID,
    Sp_Code,
    Scientific_Name,
    Native_Status,
    Growth_Habit,
    Rating,
    Source,
    Plot,
    Block,
    Shelter,
    Plot_number,
    Population,
    Treatment,
    rgr_v_FM,
    rgr_v_MA,
    rgr_v_AM,
    rgr_v_MJ
  ) %>% 
  pivot_longer(cols = c("rgr_v_FM":"rgr_v_MJ"),
               names_to = "var_mo",
               values_to = "rate")

write_csv(rgr_v,here::here("data/RGR_volume.csv"))

```

Start graphing, what do I want to see? growth rate by rating or native status across months
```{r}
# Order Invasive rating: Native, Non-native, Limited, Moderate, High
# Order Month rates
rgr_h <- read_csv(here::here("data/RGR_height.csv"))
rgr_l <- read_csv(here::here("data/RGR_leaves.csv"))
rgr_v <- read_csv(here::here("data/RGR_volume.csv"))

rgr_h$Rating<- factor(rgr_h$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_h$var_mo<- factor(rgr_h$var_mo, levels = c("rgr_h_FM", "rgr_h_MA","rgr_h_AM","rgr_h_MJ"))

# Change violin plot colors by groups
ggplot(rgr_h, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()


rgr_l$Rating<- factor(rgr_l$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_l$var_mo<- factor(rgr_l$var_mo, levels = c("rgr_l_FM", "rgr_l_MA","rgr_l_AM","rgr_l_MJ"))

ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

rgr_v$Rating<- factor(rgr_v$Rating, levels = c("Native", "Non-native","Limited","Moderate","High"))
rgr_v$var_mo<- factor(rgr_v$var_mo, levels = c("rgr_v_FM", "rgr_v_MA","rgr_v_AM","rgr_v_MJ"))

ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()


```
Should almost look at each of these graphs by population, so it's limited to community. It's possible an invasive from Angelo has the same growth rate as a native from santa barbara.

```{r}
rgr_h_ACR <- rgr_h %>% 
  filter(Population == "ANGELO")

rgr_h_MCL <- rgr_h %>% 
  filter(Population == "MCL")

rgr_h_BORR <- rgr_h %>% 
  filter(Population == "BORR")

rgr_h_SEDG <- rgr_h %>% 
  filter(Population == "SEDG")

# Change violin plot colors by groups
ggplot(rgr_h_ACR, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_ACR, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_MCL, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_MCL, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_BORR, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_BORR, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_h_SEDG, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_h_SEDG, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()

ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_l, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()



ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Rating)) +
  geom_violin()
ggplot(rgr_v, aes(x=var_mo, y=rate, fill=Native_Status)) +
  geom_violin()
```
These graphs should give us clues about invasive coverage and niche dynamics and competition within each plot.

Now curious how these relate to reproductive allocation/scaling

Remember we didn't know who each grass was in February. So might need to throw those out OR separate by growth habit.

## Photosynthetic Rate

```{r}
#| eval: false
#| warning: false
#| message: false
# how to tell it the header and which rows to skip
# how to read through the filenames in a certain folder and make a list and then read those in 1 by 1
getwd()
licor_files <- list.files("/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/licor")
licor_files[4]
licor_files_path <- paste0("data/licor/",licor_files)
licor_files_path[4]

trial <- read_csv(here::here(licor_files_path[4]), skip = 1, col_select = "obs":"gtc")

lic <- list() # creates a list
for (k in 1:length(licor_files_path)){
 lic[[k]] <- read_csv(here::here(licor_files_path[k]), skip = 1, col_select = "obs":"gtc")
}

lic <- lapply(lic, function(tbl) {
  tbl$WUE <- as.numeric(tbl$WUE)
  return(tbl)
})

licor <- bind_rows(lic)
licor <- licor %>% 
  filter(!is.na(obs))

write_csv(licor,here::here("data/licor_clean.csv"))

```

Now I need to bring in Licor leaf area csv and join. Also need to double check numbers for each species.

GALPAR, Holvir, miccal not working -- fixed, just needed to adjust minimum size.

```{r}
#| eval: false
acmpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ACMPAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
acmpar<-acmpar$summary

aircar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AIRCAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
aircar<-aircar$summary

amsmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AMSMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
amsmen<-amsmen$summary

avebar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/AVEBAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
avebar<-avebar$summary

broare = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROARE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
broare<-broare$summary

brodia = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BRODIA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brodia<-brodia$summary

brohor = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROHOR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brohor<-brohor$summary

bromad = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/BROMAD",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
bromad<-bromad$summary

calmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CALMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
calmen<-calmen$summary

caucou = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CAUCOU",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
caucou<-caucou$summary

cenmel = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CENMEL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
cenmel<-cenmel$summary

clapur = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CLAPUR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
clapur<-clapur$summary

corfil = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/CORFIL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
corfil<-corfil$summary

daupus = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/DAUPUS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
daupus<-daupus$summary

elycap = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ELYCAP",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
elycap<-elycap$summary

erobot = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/EROBOT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
erobot<-erobot$summary

erocic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/EROCIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erocic<-erocic$summary

esccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/ESCCAL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
esccal<-esccal$summary


galpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/GALPAR", 
                distance.pixel = 355, 
                known.distance = 3,
                low.size = 0.005,
                log = TRUE)
          #      save.image = TRUE)
galpar<-galpar$summary

gerdis = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/GERDIS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
gerdis<-gerdis$summary

holvir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/HOLVIR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
                # save.image = TRUE)
holvir<-holvir$summary

hypgla = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/HYPGLA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
hypgla<-hypgla$summary

lupbic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPBIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
lupbic<-lupbic$summary

lupmic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPMIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupmic<-lupmic$summary

lupsuc = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/LUPSUC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupsuc<-lupsuc$summary

madgra = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/MADGRA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
madgra<-madgra$summary

miccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/MICCAL", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
miccal<-miccal$summary

phatan = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/PHATAN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
phatan<-phatan$summary

plaere = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/PLAERE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
plaere<-plaere$summary

stipul = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/STIPUL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
stipul<-stipul$summary

torarv = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/TORARV",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
torarv<-torarv$summary

trihir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/TRIHIR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
trihir<-trihir$summary

vicsat = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/Licor Leaf Area/VICSAT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
vicsat<-vicsat$summary

# missing galpar, miccal, holvir
leaf_licor_area <- rbind(acmpar,
                         aircar,
                         amsmen,
                         avebar,
                         broare,
                         brodia,
                         brohor,
                         bromad,
                         calmen,
                         caucou,
                         cenmel,
                         clapur,
                         corfil,
                         daupus,
                         elycap,
                         erobot,
                         erocic,
                         esccal,
                         galpar,
                         gerdis,
                         holvir,
                         hypgla,
                         lupbic,
                         lupmic,
                         lupsuc,
                         madgra,
                         miccal,
                         phatan,
                         plaere,
                         stipul,
                         torarv,
                         trihir,
                         vicsat)

write_csv(leaf_licor_area, here::here("data/leaf_licor_area.csv"))
```

Double check that the numbers of values per species aligns with number of images per species folder. -- All good

Then merge with Licor dataset by plant ID.

```{r}
# Separate date from plant ID, make in date format
leaf_licor_area <- read_csv(here::here("data/leaf_licor_area.csv"))
head(leaf_licor_area)
licor <- read_csv(here::here("data/licor_clean.csv"))
head(licor)

licor$Date <- as.Date(substr(licor$date, start=1, stop=10))
str(licor$Date)

leaf_licor_area$Date <- as.Date(substr(leaf_licor_area$sample, start=11, stop=18), "%Y%m%d")
str(licor$Date)

# separate plant ID from same and add periods to plant id
leaf_licor_area$Plant_ID <- substr(leaf_licor_area$sample, start=1,stop=9)
leaf_licor_area$Plant_ID <- paste0(
  substr(leaf_licor_area$Plant_ID, 1, 2),
  ".",
  substr(leaf_licor_area$Plant_ID, 3, nchar(leaf_licor_area$Plant_ID))
)
leaf_licor_area$Plant_ID

leaf_licor_area$Plant_ID <- paste0(
  substr(leaf_licor_area$Plant_ID, 1, 9),
  ".",
  substr(leaf_licor_area$Plant_ID, 10, nchar(leaf_licor_area$Plant_ID))
)
leaf_licor_area$Plant_ID

# capitalize Licor plant id
licor$Plant_ID <- toupper(licor$`plant id`)

# NAs to fill in
# C6.PHATAN.1 is missing, no photo
# H1.AVEBAR.1 is missing, no photo
# G8.VICSAT.2 is missing, no photo
# A2.BROMAD.2 is missing, no photo
# H1.BROMAD.2 is missing, no photo
# C7.ESCCAL.2 is missing, no photo
# D4.TRIHIR.2 is missing, no photo
# C4.TRIHIR.1 is missing, no photo
# A6.STIPUL.2 is missing, no photo
# D7.CORFIL.2 is missing, no photo

# Now I need to interpolate leaf area for the missing ones
leaf_licor_area$Species <- substr(leaf_licor_area$Plant_ID,4,9)
lf_area_avg <- leaf_licor_area %>% 
  group_by(Species) %>% 
  summarise(lf_avg = mean(total.leaf.area))

lf_NAs <- data.frame(
  Plant_ID = c(
    "C6.PHATAN.1",
    "H1.AVEBAR.1",
    "G8.VICSAT.2",
    "A2.BROMAD.2",
    "H1.BROMAD.2",
    "C7.ESCCAL.2",
    "D4.TRIHIR.2",
    "C4.TRIHIR.1",
    "A6.STIPUL.2",
    "D7.CORFIL.2"
  ))

lf_NAs$Species <- substr(lf_NAs$Plant_ID,4,9)
lf_NAs_fill <- left_join(lf_NAs,lf_area_avg,by = c("Species" = "Species"))

str(leaf_licor_area)
str(lf_NAs_fill)

lf_NAs_fill$Date <- NA
lf_NAs_fill$sample <- NA
lf_NAs_fill <- lf_NAs_fill %>% 
  mutate(total.leaf.area = lf_avg) %>% 
  select(
    sample,
    total.leaf.area,
    Date,
    Plant_ID,
    Species
  )

leaf_licor_area <- rbind(leaf_licor_area, lf_NAs_fill)
### Should I have actually done this by species x population


# merge with licor leaf area
licor2 <- left_join(licor,leaf_licor_area,by=c("Plant_ID" = "Plant_ID"))

# Come back and investigate NAs for licor2 -- why are some leaf areas missing
licor.na <- licor2 %>% 
 filter(is.na(total.leaf.area)) %>% 
  select(
    date
    , A
    , `plant id`
    , Plant_ID
  )

# multiply focal data by leaf area fraction
licor2$leaf_ratio <- 6/licor2$total.leaf.area
licor2 <- licor2 %>% 
  filter(!is.na(total.leaf.area)) %>% # filter NAs for now
  mutate(WUE_adj = leaf_ratio*WUE,
         E_adj = leaf_ratio*as.numeric(E),
         A_adj = leaf_ratio*as.numeric(A),
         Ci_adj = leaf_ratio*as.numeric(Ci),
         gsw_adj = leaf_ratio*as.numeric(gsw))
str(licor2)

# remove licor data with gsw below threshold
# licor2 2315
licor2_b <- licor2[-which(licor2$gsw_adj < 0.03),]

# test for licor outliers by species, population, and treatment
# Maybe I do this later though? Because what if an outlier in photosynthetic rate goes with a reproductive outlier?

# average licor data by plant_id
licor3 <- licor2_b %>% 
  group_by(Plant_ID) %>% 
  summarise(A_adj_av = mean(A_adj, na.rm=TRUE),
            WUE_adj_av= mean(WUE_adj, na.rm = TRUE),
            E_adj_av=mean(E_adj, na.rm = TRUE),
            Ci_adj_av = mean(Ci_adj, na.rm = TRUE),
            gsw_adj_av = mean(gsw_adj, na.rm = TRUE))

# extract species and plot from plant id
licor3$Sp_Code <- substr(licor3$Plant_ID,4,9)
licor3$Plot <- substr(licor3$Plant_ID,1,2)

# merge with plot and species info
plots <- read_csv(here::here("data/Plot_info.csv"))
licor4 <- left_join(licor3,plots, by = join_by(Plot))
spp <- read_csv(here::here("data/Species Codes.csv"))
licor5 <- left_join(licor4,spp, by = join_by(Sp_Code))
```



### Explore licor
```{r}
summary(licor5)
# okay some things are way off with WUE, and others -- will need to deal with outliers at some point

p <- ggplot(licor5, aes(x=Sp_Code, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

p <- ggplot(licor5, aes(x=Rating, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

p <- ggplot(licor5, aes(x=Native_Status, y=A_adj_av)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )
p

```

At some point I need to do a correlation table with all the traits.

## Leaf Mass Area
```{r}
#| eval: false
acmpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/ACMPAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
acmpar<-acmpar$summary

aircar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AIRCAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
aircar<-aircar$summary

amsmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AMSMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
amsmen<-amsmen$summary

avebar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/AVEBAR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
avebar<-avebar$summary

broare = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROARE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
broare<-broare$summary

brodia = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BRODIA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brodia<-brodia$summary

brohor = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROHOR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
brohor<-brohor$summary

bromad = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/BROMAD",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
bromad<-bromad$summary

calmen = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CALMEN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
calmen<-calmen$summary

caucou = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CAUCOU",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
caucou<-caucou$summary

cenmel = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CENMEL",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
cenmel<-cenmel$summary

clapur = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/CLAPUR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
clapur<-clapur$summary

daupus = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/DAUPUS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
daupus<-daupus$summary

elycap = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/ELYCAP",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
elycap<-elycap$summary

erobot = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/EROBOT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erobot<-erobot$summary

erocic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/EROCIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
erocic<-erocic$summary

galpar = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/GALPAR", 
                distance.pixel = 355, 
                known.distance = 3,
                low.size = 0.005,
                log = TRUE)
          #      save.image = TRUE)
galpar<-galpar$summary

gerdis = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/GERDIS",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
gerdis<-gerdis$summary

holvir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/HOLVIR", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
                # save.image = TRUE)
holvir<-holvir$summary

hypgla = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/HYPGLA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
hypgla<-hypgla$summary

lupbic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPBIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
lupbic<-lupbic$summary

lupmic = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPMIC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupmic<-lupmic$summary

lupsuc = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/LUPSUC",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
lupsuc<-lupsuc$summary

madgra = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/MADGRA",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
madgra<-madgra$summary

miccal = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/MICCAL", 
                distance.pixel = 355, 
                known.distance = 3, 
                low.size = 0.005,
                log = TRUE)
             #   save.image = TRUE)
miccal<-miccal$summary

phatan = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/PHATAN",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
phatan<-phatan$summary

plaere = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/PLAERE",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
plaere<-plaere$summary

torarv = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/TORARV",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
torarv<-torarv$summary

trihir = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/TRIHIR",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE,
              low.size = 0.005)
trihir<-trihir$summary

vicsat = run.ij(set.directory = "/Users/brookewainwright/Documents/GitHub_Projects/repro-grow/data/LMA Leaf Images/VICSAT",
              distance.pixel = 355,
              known.distance = 3,
              log = TRUE)
vicsat<-vicsat$summary

# missing galpar, miccal, holvir
lma_lf_area <- rbind(acmpar,
                         aircar,
                         amsmen,
                         avebar,
                         broare,
                         brodia,
                         brohor,
                         bromad,
                         calmen,
                         caucou,
                         cenmel,
                         clapur,
                         daupus,
                         elycap,
                         erobot,
                         erocic,
                         galpar,
                         gerdis,
                         holvir,
                         hypgla,
                         lupbic,
                         lupmic,
                         lupsuc,
                         madgra,
                         miccal,
                         phatan,
                         plaere,
                         torarv,
                         trihir,
                         vicsat)

write_csv(lma_lf_area, here::here("data/LMA_lf_area.csv"))


```


```{r}
lf_area <- read_csv(here::here("data/LMA_lf_area.csv"))
# There may be errors in this -- not sure how to check

## Extract plant id, species, plot
lf_area$Date <- as.Date(substr(lf_area$sample, start=11, stop=18), "%Y%m%d")

# separate plant ID from same and add periods to plant id
lf_area$Plant_ID <- substr(lf_area$sample, start=1,stop=9)
lf_area$Plant_ID <- paste0(
  substr(lf_area$Plant_ID, 1, 2),
  ".",
  substr(lf_area$Plant_ID, 3, nchar(lf_area$Plant_ID))
)
lf_area$Plant_ID

lf_area$Plant_ID <- paste0(
  substr(lf_area$Plant_ID, 1, 9),
  ".",
  substr(lf_area$Plant_ID, 10, nchar(lf_area$Plant_ID))
)
lf_area$Plant_ID

lf_area$Sp_Code <- substr(lf_area$Plant_ID,4,9)


# merge with plot and species info
#lf_area2 <- left_join(lf_area,plots, by = join_by(Plot))
#lf_area3 <- left_join(lf_area2,spp, by = join_by(Sp_Code))

# Now merge with leaf mass and calculate LMA
mass <- read_csv(here::here("data/Leaf Mass_2023_20231009.csv"))
head(mass)

# Remove perennials
mass_a <- mass %>% 
  filter(!Sp_Code %in% c("CORFIL","ESCCAL","STIPUL","DANCAL"))

# 696 masses
# 691 leaf areas

# I want to know how many observations per species for each dataset, start there

library(dplyr)
ct_area <- lf_area %>% count(Sp_Code)
ct_mass <- mass_a %>% count(Sp_Code)

join_check <- full_join(lf_area,mass_a,by=join_by("Plant_ID"))
join_check %>% filter(is.na(total.leaf.area))
join_check %>% filter(is.na(Dry_leaf_mass_g))
# Missing:
# H1.AVEBAR.2 area
# H3.BRODIA.2 Mass
# G8.VICSAT.2 area

str(lf_area)
str(mass_a)

LMA <- full_join(lf_area,mass_a, by = join_by(Plant_ID, Sp_Code == Sp_Code))
# LMA is measured in g/m^2

LMA$area_m2 <- LMA$total.leaf.area/10000
LMA$lma_g_m2 <- LMA$area_m2/LMA$Dry_leaf_mass_g

summary(LMA)
LMA[which(is.na(LMA$Dry_leaf_mass_g)),"Plant_ID"]
LMA[which(is.na(LMA$total.leaf.area)),"Plant_ID"]
# interpolate LMA based on spp x population averages
# First make sure they are NA so they don't effect averages
LMA$lma_g_m2[which(LMA$Plant_ID == "H1.AVEBAR.2")] <- NA
LMA$lma_g_m2[which(LMA$Plant_ID == "H3.BRODIA.2")] <- NA
LMA$lma_g_m2[which(LMA$Plant_ID == "G8.VICSAT.2")] <- NA
# H1.AVEBAR.2 area
# H3.BRODIA.2 Mass
# G8.VICSAT.2 area
LMA[which(is.na(LMA$lma_g_m2)),"Plant_ID"]

# merge with plot and species info
LMA$Plot <- substr(LMA$Plant_ID,1,2)
LMA2 <- left_join(LMA,plots, by = join_by(Plot))
LMA <- left_join(LMA2,spp, by = join_by(Sp_Code))

pop_spp_avg <- LMA %>% 
  group_by(Sp_Code,Population) %>% 
  summarise(avg_lma = mean(lma_g_m2,na.rm = TRUE))

pop <- LMA[which(LMA$Plant_ID == "H1.AVEBAR.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "AVEBAR" & pop_spp_avg$Population == "SEDG"),"avg_lma"]
LMA[which(LMA$Plant_ID == "H1.AVEBAR.2"),"lma_g_m2"] <- avg

pop <- LMA[which(LMA$Plant_ID == "H3.BRODIA.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "BRODIA" & pop_spp_avg$Population == "WWP"),"avg_lma"]
LMA[which(LMA$Plant_ID == "H3.BRODIA.2"),"lma_g_m2"] <- avg

pop <- LMA[which(LMA$Plant_ID == "G8.VICSAT.2"),"Population"]
avg <- pop_spp_avg[which(pop_spp_avg$Sp_Code == "VICSAT" & pop_spp_avg$Population == "BORR"),"avg_lma"]
LMA[which(LMA$Plant_ID == "G8.VICSAT.2"),"lma_g_m2"] <- avg

LMA[which(is.na(LMA$lma_g_m2)),"Plant_ID"]

# Check for duplicates
LMA[which(duplicated(LMA$Plant_ID)),]


```

### Explore LMA
```{r}
summary(LMA)
# okay some things are way off with WUE, and others -- will need to deal with outliers at some point

ggplot(LMA, aes(x=Sp_Code, y=lma_g_m2)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )


ggplot(LMA, aes(x=Rating, y=lma_g_m2)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

```

## Leaf Area
```{r}
#| eval: false
### do not evaluate yet
## Need to input the single leaf entries for MICCAL, GALPAR, etc.


summary(lf_area)

ggplot(lf_area, aes(x=Sp_Code, y=total.leaf.area)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

ggplot(lf_area, aes(x=Rating, y=total.leaf.area)) +
  geom_violin() + stat_summary(fun.data=mean_sdl, mult=1, 
                 geom="pointrange", color="red") + theme(axis.text.x = element_text(angle=90,hjust = 1, size=14, color="black") )

```


